### 2025-02-07 Research_report.tex
Initial Overleaf Import

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\begin{document}

\title{Quantitative, Multispecies Monitoring at a Continental Scale}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Gledis Guri}
\author[b]{Owen Liu}
\author[a]{Ryan P. Kelly}
\author[a]{Megan Shaffer}
\author[b]{Kim Parsons}
\author[b]{Krista M. Nichols}
\author[a]{Pedro F. P. Brandão-Dias}
\author[b]{Andrew Olaf Shelton}

\affil[a]{University of Washington}
\affil[b]{NOAA fisheries science}

% Please give the surname of the lead author for the running footer
\leadauthor{Guri}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership. The significance statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
AOS, KMN, RPK, MS, and KP designed the research;
MS and PB did the laboratory and the bioinformatic work;
GG OL RPK AOS analysed the data;
GG OL RPK AOS analysed interpreted the results;
GG OL RPK AOS analysed wrote the manuscript}
\authordeclaration{The authors declare no competing interest.}
\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: gguri@uw.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{eDNA $|$ Bayesian $|$ bio-monitoring $|$ quantification $|$ fish}

\begin{abstract}
Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural resources management at a broad geographic scale. These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously. Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological research. However, unlike single-species techniques, metabarcoding by itself can reflect only the proportions of target-species DNA present, not their absolute quantities. Here, we combine qPCR and metabarcoding data derived from the same samples to map quantities of many fish species along the US West Coast in three dimensions, demonstrating a technique of practical importance for both management and ecology. We derive spatially explicit estimates of eDNA abundance for 12 common species of commercial and management importance and point the way to quantitative surveys of wild species using molecules alone. We find that species distribution maps derived from eDNA effectively capture species niches and spatial distributions. Notably, our analyses identified biodiversity hotspots that align with previously documented regions of ecological significance, such as the Columbia River plume and the Heceta Bank.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\firstpage[4]{4}
Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}. However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2024}). While the use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g. there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2024}, and coastal oceans \cite{maes2023, shelton2019}). However, uses of quantitative eDNA observations in management applications have significantly lagged behind their use in occurrence and biodiversity applications.

Quantitative estimates of DNA are more common with species-specific assays (e.g. qPCR and ddPCR) because these approaches directly measure DNA concentration. In contrast, multispecies analysis (metabarcoding) reflects information on many species of interest simultaneously, but at the costs of producing compositional estimates of proportional abundance \cite{gloor2017a, mclaren2019} that may require statistical correction for species-specific biases due to PCR amplification \cite{shelton2023}. Consequently, metabarcoding data alone cannot yield information about the abundance of DNA from environmental samples (e.g. \cite{guri2024a}).

While trawling and other traditional sampling methods have been the fundamental tools for directly measuring fish abundance and biomass, they are often impractical for species inhabiting inaccessible habitats, occurring at low densities, or being fragile and easily injured during capture. By contrast, eDNA sampling captures information on a broader range of taxa, including those that are difficult or impossible to study with traditional gear. As such, eDNA can meaningfully complement traditional stock assessment methods for commercially important species while at the same time revealing the abundance and distribution of elusive or understudied species.

Here, we develop quantitative estimates of eDNA abundance for a diverse group of fish species spanning different depth niches across 10 degrees of latitude along the US West Coast entirely from eDNA. We integrate single- and multi-species (using MiFish-U universal primers \cite{miya2015}) techniques to create three-dimensional distribution maps, leveraging models established in previous studies \cite{shelton2022, guri2024a, shelton2023, allan2023}. By calibrating metabarcoding data -- correcting raw species observations to their initial proportions using a mock community \cite{shelton2023} -- and using the absolute eDNA quantity of a reference species (here Pacific Hake (\textit{Merluccius productus})), we expand initial estimated proportions into absolute estimates of the abundance of each species’ eDNA \cite{guri2024a}. We subsequently smooth the estimated eDNA quantities to generate species distribution maps \cite{liu2023}. We present results for 12 fish species to illustrate the approach, but note that it immediately generalizes to any ecological community of interest.

The eDNA-derived estimates we provide here speak directly to the requirements of natural-resources management at large scales. For example, indices of abundance have a central role in determining stock status for fisheries applications (i.e., stock assessments in the U.S. under the Magnuson-Stevens Act) as well as informing conservation status (e.g., under the U.S. Endangered Species Act). Thus eDNA-derived abundance indices can fill a variety of existing requirements for management. While these indices are still in their infancy and will require additional technical developments to ensure their proper usage, they are conceptually and practically little different from abundance estimates derived from more traditional ocean survey methods (i.e., net-based or acoustic surveys). Indeed, eDNA methods may provide some notable improvements because single- and multi-species eDNA approaches avoid tradeoffs involved with physically capturing many species simultaneously using a single net. Molecular surveys capture many species with the same "gear" using methods that are easily replicated in space and time. As U.S. agencies begin to implement the new National Aquatic eDNA Strategy \cite{kelly2024} -- the first goal of which is to fold eDNA data into federal decisionmaking -- this and other large-scale examples point the way to maximizing the benefit of this information-rich datastream for natural-resources management at a continental scale.

\section*{Results}

We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig. \ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI.pdf}{SI Appendix, Fig. S1}, Pacific Herring and Jack Mackerel). For other high-biomass species otherwise lacking well-described distributions, we provide the first species distribution models (Fig. \ref{fig:fig1}, \href{SI.pdf}{SI Appendix, Fig. S1}). We find that the estimated eDNA concentrations are correlated among species with similar traits and habitat preferences (Fig. \ref{fig:fig2}). Emergent patterns of species richness and estimated eDNA aggregation correspond with known ecological drivers, evidence that eDNA traces can reliably capture ecologically meaningful patterns (Fig. \ref{fig:fig3}). Particularly for forage fish species – which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012LittleFB} – molecular methods offer a tractable and quantitative index of abundance over large spatial scales.

Below we highlight results for six species (Fig. \ref{fig:fig1}) representing a range of habitat associations and general species information; distribution maps for all the remaining species are presented in \href{SI.pdf}{SI Appendix, Fig. S1}.

\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=16cm]{plots/1_Figure_1.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Estimated species eDNA concentration (A) across 0, 50, 150, 300, and 500m depth samples and known depth distribution of those species from literature (B).}
\label{fig:fig1}
\end{figure*}

\subsection*{Surface}
Northern anchovy (\textit{Engraulis mordax}) and Pacific sardine (\textit{Sardinops sagax}) are the two iconic coastal pelagic species of the eastern North Pacfic \cite{steinbeck1945} and play a vital ecological role linking lower and upper trophic levels (e.g., Chinook salmon diets, sea lions, \cite{kaplan2019}); historically these supported large commercial fisheries particularly in California. Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2°N; Fig. \ref{fig:fig1}). Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24°C for sardine vs. 10–14°C for anchovy); Pacific sardine also typically inhabits slightly greater depths (25; Fig. \ref{fig:fig1}). eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI.pdf}{SI Appendix, Fig. S7}) such that eDNA captured subtle yet meaningful differences in habitat preferences. Furthermore, eDNA abundance patterns aligned with stock assessments in indicating a higher abundance of northern anchovy; however, the magnitude of difference varied, with eDNA suggesting a 2.5-fold difference, while stock assessments indicated a 35-fold difference in abundance. [CITE stock assessment OOO]

\subsection*{Mid-water}
Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} [ADD SCUBA REF OOO]. Moreover, because it is a common bycatch species in the hake (\textit{Merluccius productus}) fishery, limits on bycatch of widow rockfish can also limit the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}). Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat. We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI.pdf}{SI Appendix, Fig. S7}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.

The eDNA patterns for northern lampfish (Stenobrachius leucopsarus) -- an abundant myctophid that migrates from 400-700m up to 20-200m at night \cite{moku2000,suntsov2008} -- are again consistent with the ecology of the species, having notably lower concentrations in surface samples. Owing to the diel vertical migration of the northern lampfish, the signal of eDNA for this species is observed across depths.

\subsection*{Benthic}
Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common yet little studied, such as the dragonfish (\textit{Tactostoma macropus}). Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}. The eDNA concentrations appeared to reflect this activity (\href{SI.pdf}{SI Appendix, Fig. S7}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig. \ref{fig:fig1}), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude. However, no other quantitative estimates of abundance or detailed distribution patterns for dragonfish exist in the current literature.

Also consistent with known habitats are commercially caught flatfish such as Dover sole (\textit{Microstomus pacificus}), which lives at 200m to 1200m along the Pacific coast \cite{drazen2007,brodziak2000,drazen2012}; individuals move progressively deeper as they grow \cite{vetter1994,hunter1990}. The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI.pdf}{SI Appendix, Fig. S7}\textit{A}) \cite{ono2016}.

\subsection*{Species Associations}

Pairwise correlations of DNA concentration showed interesting patterns among species associated with depth (Fig. \ref{fig:fig2}). Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$). Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$). Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$) with both more common in the more southern (and somewhat warmer) waters (Fig. \ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig. S7}). Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig. \ref{fig:fig2}) consistent with more northerly distribution (\href{SI.pdf}{SI Appendix, Fig. S7}). Together these finding suggest that eDNA is capturing both depth and latitudinal habitats associations. Deep-water species are distinct (Fig. \ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig. S7}), not strongly correlated with those from other zones (Fig. \ref{fig:fig2}). Deeper water tends to feature lower and more diffuse fish eDNA concentrations (Fig. \ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig. S7}). Lower overall eDNA concentrations and fewer deep water samples may make similarities among species distributions more difficult to detect at 300m and deeper.

\begin{figure}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Correlation plot describing species co-abundance and co-occurance. Spearman’s rho correlation coefficient are expressed in numbers and colors ranging from -1 (red) negatively correlated to 1 (blue) positively correlated. Species are ordered in known ecological groups (surface, midwater, and deepwater) to visually indicate the gradual change of species co-occurance.}
\label{fig:fig2}
\end{figure}

%\begin{SCfigure*}[\sidecaptionrelwidth][t!]
%\centering
%\includegraphics[width=11.4cm,height=11.4cm]{frog.pdf}
%\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
%\end{SCfigure*}

\subsection*{Species richness and total eDNA accumulation}
In addition to providing species-specific patterns, eDNA metrics of abundance from individual species can be combined to reveal important aggregate ecological patterns. For the 12 species considered here, we show how species richness (see \href{SI.pdf}{SI Appendix, Fig. S2} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig. \ref{fig:fig3}\textit{A}). In part, this pattern is likely to be an artifact of the focal species included in this study -- chosen for their relevance to fisheries management as well as for their frequency in the observations (see Methods) -- but the richness-depth gradient also reflects the expectation that habitat and nutrient availability is greatest near the surface \cite{smith2002,hickey2005}. Surface waters also host the eggs, larvae, and juvenile stages of many meso- and bathypelagic species \cite{parnel2008}, some of which we detect, and hence we both expect and observe higher richness in the upper layer of the water column \cite{kim2011}.

Species richness also corresponds strongly with the influence of Columbia River plume (Fig. \ref{fig:fig3}\textit{A}), which extends well beyond the coastline \cite{hickey2003}. The nitrate-rich river water interacts with ocean currents and intensifies coastal upwelling, enhancing both primary and secondary productivity \cite{hickey2010} and likely influencing the patterns of species richness we observe (Fig. \ref{fig:fig3}\textit{A}) \cite{tolimieri2015}. The total fish eDNA accumulation (Fig. \ref{fig:fig3}\textit{B}) suggests the highest aggregation of forage fish DNA near Heceta Bank (44 degrees N; overlapping anchovy and sardine distributions are the two largest contributors to eDNA concentration) in addition to other commercialy important mesopelagic fish (\textit{Sebastes entomelas} and \textit{Tarletonbeania crenularis}) a center of abundance for commercial fisheries \cite{bosley2014,Tissot2008TwentyYO}, and elsewhere nearshore in areas of coastal upwelling.

\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=14cm]{plots/3_Figure_3.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m). (B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI.pdf}{SI Appendix, Fig. S2} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
\label{fig:fig3}
\end{figure*}

%\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
%\centering
%\includegraphics[width=17cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
%\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater). Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A). Probability density functions (PDF) of estimated eDNA concentrations across latitudinal gradients (B; square root transformed of x-axis to enhance low concentration peaks). Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
%\label{fig:fig4}
%\end{figure*}

\section*{Discussion}
We report species-specific DNA concentrations smoothed over ten degrees of latitude and across the continental shelf to 500m. We reveal distributional pattern for 12 species ranging from the iconic coastal pelagic species of the northeast Pacific to understudied-but-abundant deepwater species, and our results suggest clear pathways for the broader use of eDNA in both fisheries and ecosystem management.

\subsection*{eDNA as Management Tool}
By mapping three-dimensional spatial patterns, eDNA provides information on species distributions, habitat preferences, and ecological processes that is often otherwise unavailable. A single water sample can yield quantitative information on tens to hundreds of species – including for species for which traditional fisheries surveys are not currently conducted – without the need to develop species-specific assays. We find species-level eDNA concentrations align closely the preferred habitats physiological constraints (e.g., temperature tolerance or food availability) described for each species in the available literature. The observed latitudinal differences of eDNA concentration among species – such as those between anchovy and sardine – highlights species’ thermal ranges and largely coincide with other surveys \cite{zwolinski2024}; moreover, eDNA patterns appear to reflect diel vertical migrations as observed in myctophids, where eDNA concentrations are consistent with the species’ ontogenetic movement between surface and mid- to deep-water habitats. We also identify aggregations of richness and total eDNA near high-productivity regions of ecological and commercial value. Regions like Heceta Bank host dense aggregations of forage fish and are critical for both fishery management and for broader marine food-web stability.

For well-studied taxa, existing information helps contextualize eDNA signals, providing insights into habitat preferences, seasonal patterns, or temperature tolerances. For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa, revealing unexpected habitats. Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys [NWFSC trawl survey ref OOO]. This species has been of special interest, as its targeted commercial harvest resumed in the late 2010s after a decade of closure (harvest was less than 500 mt annually between 2003 and 2013 but has exceeded 10,000 mt annually since 2018) and no existing survey adequately captures its patterns of abundance \cite{adams2019a}.

\subsection*{Environmental Smoothing}
Our findings underline eDNA as a transformative tool for fisheries management, biodiversity conservation, and the study of environmental change. However, the observable eDNA signal at any given location and time is an integrated product of DNA production and loss across time and space \cite{gold2023}: the vast majority of DNA detectable in a given sample is of recent origin (within the previous hours to days \cite{strickler2015,collins2018,machler2018}), and it may be subject to a degree of physical transport within its window of detectability. Accordingly, the vast majority of DNA will be relatively proximate to its source in both space and time, and the distribution of DNA from a given species at any given time will be smoother and more extensive when compared to the distribution of discrete source individuals. We refer to this as environmental smoothing of eDNA signal relative to the source organisms themselves.

The majority of eDNA shed into the environment dissipates from surface waters within hours (Brasseale et al., in review RRR). Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most. Given that our sampling stations are spaced more than 30 km apart, the spatial patterns we report are likely to reflect genuine differences in fish abundance rather than artifacts of eDNA transport dynamics \cite{guri2023}. Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; \textit{Sebastes entomelas} occurs almost exclusively at the 50m depth bin; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig. \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station. Notably, the comparison between sample-specific eDNA estimation (\href{SI.pdf}{SI Appendix, Fig. S6}) and spatially smoothed maps (Fig. \ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig. S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound. This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.[Ryan check the framing of the last two sentences RRR]

Large-scale sampling combined with statistical smoothing mitigates many challenges associated with eDNA transport and degradation -- which decouple the eDNA observations from their particular source organisms -- allowing reliable detection of broad patterns in species distributions. Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature \cite{jo2019b}.

%(see \href{SI.pdf}{SI Appendix, Fig. S5}) posterior predictive

\subsection*{Integrating DNA with traditional surveys}
Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes. Here, we follow \cite{shelton2022} and Liu et al. (submitted) RRR, leveraging the same water samples to provide information on an additional 12 species of commercial importance. Elsewhere, the same set of samples has yielded information on the distribution of marine mammals (Valdivia et al, in revision) RRR. The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa. The use of these samples, from a targeted fisheries survey for hake, highlights the efficiency for which eDNA can effectively provide survey data for tens or hundreds of additional species for which traditional surveys do not exist.

This capacity for retrospective analysis allows historical samples to be re-examined at any point, insofar as the sample is not exhausted, thereby offering opportunities to address emerging priorities and to reconstruct past ecological conditions. For example, the ability to retrospectively pinpoint the origins and pathways of invasive species can inform targeted interventions, helping managers understand not only where an incursion began but also how it spread \cite{gilbey2021}. Similarly, the capacity to trace the initial onset of harmful algal blooms or the earliest presence of pathogenic microbes enables researchers and policymakers to identify the underlying drivers of these events and mitigate future outbreaks \cite{shaw2019}.

Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts. For example, \cite{diaz2020a} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material. Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time. We therefore see eDNA as an important datastream that maximizes the information value of existing research cruises, rather than a means of replacing those cruises themselves.

We have shown an example of the substantial value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease). Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional trawl-based surveys.

Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}. In reality, fish can move in between sampling location within the study area, which may bias abundance estimates  \cite{rota2009}. By contrast, fish continuously shed DNA into their surroundings, creating a dynamic signal of presence. As the shedded DNA decays, disperses, and degrades over time, fresher DNA remains at higher concentrations while older DNA persists at lower concentrations \cite{thomsen2015,jerde2011}. This cumulative, time-integrated signal allows eDNA surveys to circumvent the closure assumption and provide a more accurate representation of fish presence and activity.

Perhaps most notably, management decisions happen in units of mass or abundance, not in units of eDNA concentration. Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that are likely to vary among species. We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.

\matmethods{We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI.pdf}{SI Appendix}.

\subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
Samples were collected in 10L Niskin bottles deployed on a CTD rosette at five depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship Bell Shimada during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey. 2.5L water samples were collected from the Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}. A subset of the samples collected in \cite{shelton2022}, was used for this study.

\subsection*{Metabarcoding: Environmental Samples and Mock Community}
In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit). Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI.pdf}{SI Appendix} for details on metabarcoding methods).

Sequences were de-multiplexed and the adapters removed by Illumina sequencing. Primers were removed with Cutadapt v4.9 \cite{martin2011}. We then used DADA2 (default parameters, \cite{callahan2016}) to denoise sequences, remove chimeras, and generate amplicon sequence variants (ASVs). The sequences were then BLASTed against the NCBI nucleotide (nt) database (access: August 2024) using BLASTn algorithm with a cut off at 97\% identity, assigning the least common ancestor of the top hit, resulting in species, genus, family, or higher rank classifications.  Samples with fewer total number of reads than 1000 were filtered as indicator of poor reliability. Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI.pdf}{SI Appendix, Fig. S3}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI.pdf}{SI Appendix}). Details on bioinformatic pipeline are available in \href{SI.pdf}{SI Appendix}.

\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=17.4cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species. Model compartment indicated in the same color are run jointly. Latent parameters, state variables, data and subscripts are indicated in gray. See \href{SI.pdf}{SI Appendix} for full model breakdown.}
\label{fig:DAG}
\end{figure*}

To calibrate metabarcoding observations and account for amplification bias \cite{shelton2023,gold2023}, we generated multiple mock communities comprising a total of 39 fish species. Vouchered DNA extracts or tissues were obtained from either the University of Washington Fish Collection at the Burke Museum or the NOAA Northwest Fisheries Science Center collections. After quantifying the concentration of mitochondrial DNA template in each DNA extract, we constructed a total of eight mock communities of varying species compositions (see \href{SI.pdf}{SI Appendix, Table S1}). We selected 12 species (additional to Pacific hake) of commercial and ecological importance that co-occurred in the environmental samples and in the mock community for downstream analysis: Pacific herring (\textit{Clupea pallasii}), northern anchovy (\textit{Engraulis mordax}), northern smoothtongue (\textit{Leuroglossus stilbius}), Dover sole (\textit{Microstomus pacificus}), Pacific sardine (\textit{Sardinops sagax}), Pacific chub mackerel (\textit{Scomber japonicus}), widow rockfish (\textit{Sebastes entomelas}), northern lampfish (\textit{Stenobrachius leucopsarus}), California barracudina (\textit{Tactostoma macropus}), plainfin midshipman (\textit{Tarletonbeania crenularis}), eulachon (\textit{Thaleichthys pacificus}), Pacific jack mackerel (\textit{Trachurus symmetricus}). Because we used Pacific hake (\textit{Merluccius productus}) as a comparative reference species throughout (see below), we did not include the spatial patterns of that species in our analysis. After amplification and sequencing using the protocols outlined above, we used the mock communities to estimate species-specific amplification efficiencies and correct for amplification bias following \cite{shelton2022}.

\subsection*{Statistical analysis}
To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples. The full statistical details are described in \href{SI.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig. \ref{fig:DAG}). We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions). This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig. \ref{fig:DAG}). This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification) as well as the effect of covariates (e.g., the depth of the ocean bottom) on hake DNA concentration (Fig. \ref{fig:DAG}). This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions coming from DNA metabarcoding (Fig. \ref{fig:DAG}).

Unlike single-species observation methods, metabarcoding data is compositional data. This means that each sample only provides information about relative species abundances rather than absolute DNA concentrations. Furthermore, metabarcoding does not yield a consistent, species-specific detection probability (i.e., sensitivity). Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification, as well as on the total sequencing depth for that sample \cite{kelly2019}. Higher sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.

We follow \cite{shelton2022} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR. We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species). In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample. But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI.pdf}{SI Appendix Fig. S3}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}. Our model generates estimates of DNA concentration (copies/L) for each species. Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species; we refer to hake as our “reference species”. Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species. To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration. However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI.pdf}{SI Appendix Fig. S4}) [Ryan check the framing of the last two sentences RRR]. Samples lacking hake detections were sporadically distributed across the study area, and their omission appears not to have induced substantial bias in our estimates.

We focused on species that were relatively common in the raw dataset – those well above the lower detection limits under our sampling and sequencing conditions – ensuring greater confidence in their consistent detectability. Although many more species were detected in the raw data, the majority were much rarer and thus less reliably observed.

The above statistical model (Fig. \ref{fig:DAG}) provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m). We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA (Fig. \ref{fig:DAG}, green box). Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI.pdf}{SI Appendix}).

All statistical analysis, figure preparation, and data curation were performed in R \cite{rcoreteam2024}.

}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank Linda Park for her foresight and perseverance, and Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis. This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719). The authors gratefully acknowledge the survey expertise and support of NOAA's Pacific hake survey team and the personnel of the NOAA Ship Bell M. Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. Government.}

\showacknow{} % Display the acknowledgments section

\bibsplit[11]
%Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2025-02-07 SI.tex
Initial Overleaf Import

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\title{Quantitative, Multispecies Monitoring at a Continental Scale}
\author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Krista M. Nichols, Pedro F. P. Brandão-Dias and Andrew Olaf Shelton}
\correspondingauthor{Gledis Guri.\\E-mail: gguri@uw.edu}

\begin{document}

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section*{Extended Methods}
\subsection*{qPCR}
Environmental samples were analysed in triplicate using the multiplexed assay targeting Pacific hake, Eulachon and Pacific lamprey as described in \cite{ramon-laca2021} on a QuanStudio 6 (Applied Biosystems). Only Pacific hake was used in this study due to the remaining targets having little positive amplification. The samples underwent real-time thermocycler protocol including an initial denaturation step at 95°C 10 min followed by 40 cycles of 15 s at 95°C and 1 min at 58°C. All thermocycler reactions were run in 20 $\mu$L volume consisting of 10 $\mu$L of TaqMan Environmental Master Mix 2.0, 1 $\mu$L of each primer (forward: AAATGTTTAAACTAGAGCCGAATAGC and reverse: TCGTGGAGTCAAAGTGGGGTAGA, 10 nM concentration each), 0.5 $\mu$L of probe (AAATGTTTAAACTAGAGCCGAATAGC; 10nM concentration), 3 $\mu$L of dH20 and 2 $\mu$L of DNA template. To monitor quality control, all runs included an internal positive control (IPC) to detect PCR inhibition and a negative template control (NTC) containing water instead of DNA template. Any IPC delay exceeding 0.5 cycles in the NTC was considered inhibition.

Alongside environmental samples standard samples were analyzed constructed from a 120 bp synthetic DNA fragment (gBlock; IDT) representing the 12S region of Pacific hake (\textit{Merluccius productus}), which encompasses the 101 bp 12S qPCR target (sequence information available in Ramón-Laca et al. 2021). The synthetic DNA fragment was diluted to create a series of standards with final concentrations ranging from $10^{0}$ to $10^{5}$ copies/$\mu$L.

\subsection*{Metabarcoding}
In total, 568 samples, including 554 environmental samples, 7 PCR blanks and 7 positive controls (holding only kangaroo DNA) were amplified were amplified using MiFish-U universal primers \cite{miya2015} with Illumina tails (forward TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGGCCGGTAAAACTCGTGCCAGC; reverse GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCATAGTGGGGTATCTAATCCCAGTTTG) by using a two-step PCR protocol. The DNA was amplified in the first PCR reaction (PCR 1) in a 20 µL reaction consisting of: 10 µL of Phusion Master Mix (2X), 0.4 µL of the forward primer (10 µM); 0.4 µL of the reverse primer (10 µM), 0.6 µL of 100\% DMSO, 0.5 µL of rAlbumin (20 µg/µL), 4.4 µL of nuclease-free water and 2 µL of DNA template. Reactions were run with the following cycling conditions: an initial denaturation of 98°C for 30 sec; followed by 35 cycles of 98°C for 10 sec, 60°C for 30 sec, and of 72°C for 3 sec; with a final extension of 72°C for 10 min.

PCR product was cleaned using Ampure Beads (1.2x) and then indexed in a second PCR reaction (PCR 2) consiting of 8 cycles. Resulting products were visualized on a 2\% agarose gel and quantified using Quant-iT™ dsDNA Assay Kit (Thermo Fisher Scientific, USA) with Fluoroskan™ Microplate Fluorometer (Thermo Fisher Scientific, USA). Indexed products were pooled into libraries for sequencing, and then size- selected to extract only the target fish band using the E-Gel™ SizeSelect™ II Agarose Gels (Thermo Fisher Scientific, USA).  Subsequently the libraries were sequenced on Illumina MiSeq platform using the v3 600 cycle kit.

%A total of 554 environmental samples of metabarcoding samples sequenced are 554 of which 535 had higher sequencing depth than 1000 reads (indicating good quality samples). From those we indicated that only 371 samples had Pacific hake (\textit{Merluccius productus}) present jointly with qPCR positive detection of the same species. All libraries produced a total of 56,279,345 reads with 54,285,713 reads (96.5\%) belonging to class \textit{Actinopteri}. A total of 222 unique species was identified from which 165 belonged to fish.

\subsection*{Mock community}

We quantified the total genomic DNA in each extract using Qubit HS assay. We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly match the template and give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system). Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water. The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read. Droplets were determined to be positive after drawing a threshold based on NTCs.

\subsection*{Bioinformatics}
PEDRO

\subsection*{Bayesian model}
OLE

\subsection*{Model convergance}
OLE + GLED

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.99\textwidth]{plots/5. Supplementary Figure 1.jpg}
\caption{First figure}
\end{figure}

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.99\textwidth]{plots/6. Supplementary Figure 2.jpg}
\caption{First figure}
\end{figure}

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.99\textwidth]{plots/7. Supplementary Figure 3.jpg}
\caption{First figure}
\end{figure}

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.99\textwidth]{plots/8. Supplementary Figure 4.jpg}
\caption{First figure}
\end{figure}

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.80\textwidth]{plots/9. Supplementary Figure 5.jpg}
\caption{First figure}
\end{figure}

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=0.89\textwidth]{plots/10. Supplementary Figure 6.jpg}
\caption{First figure}
\end{figure}

\begin{table}\centering
\caption{Initial relative abundances of the selected species across different mock communities.}
    \begin{tabular}{lcccccccc}
        \toprule
        {Species} & \multicolumn{2}{c}{Mock1} & \multicolumn{2}{c}{Mock2} & \multicolumn{2}{c}{Mock3} & \multicolumn{2}{c}{Mock4} \\
        & Even & Skew & Even & Skew & Even & Skew & Even & Skew \\
        \midrule
        \textit{Clupea pallasii}           & 0.437 & 0.325  & 0.042  & 0.009     & NA         & NA          & 0.025      & 0.038 \\
        \textit{Engraulis mordax}          & 0.122 & 0.034  & 0.15 & 0.15       & 0.146      & 0.16       & 0.308      & 0.478 \\
        \textit{Leuroglossus stilbius}     & 0.108 & 0.29   & NA & NA         & NA         & NA         & NA         & NA \\    
        \textit{Merluccius productus}      & 0.079 & 0.144  & 0.12 & 0.06       & 0.117      & 0.075      & 0.247      & 0.287 \\
        \textit{Microstomus pacificus}     & NA    & NA     & 0.057  & 0.215     & NA         & NA         & NA         & NA \\    
        \textit{Sardinops sagax}           & 0.028 & 0.033  & 0.14 & 0.035      & 0.137      & 0.025      & 0.289      & 0.131 \\
        \textit{Scomber japonicus}         & NA    & NA     & 0.122  & 0.053      & 0.119      & 0.027     & NA         & NA \\    
        \textit{Sebastes entomelas}        & NA    & NA     & NA & NA          & 0.068      & 0.218     & NA         & NA \\    
        \textit{Stenobrachius leucopsarus} & NA    & NA     & NA & NA          & 0.356      & 0.487     & NA         & NA \\    
        \textit{Tactostoma macropus}       & NA    & NA     & 0.13 & 0.324     & NA         & NA         & NA         & NA \\    
        \textit{Tarletonbeania crenularis} & NA    & NA     & 0.083  & 0.124     & NA         & NA         & NA         & NA \\    
        \textit{Thaleichthys pacificus}    & 0.123 & 0.113  & 0.099  & 0.012     & NA         & NA          & 0.013      & 0.021 \\
        \textit{Trachurus symmetricus}     & 0.103 & 0.06   & 0.057  & 0.018      & 0.056      & 0.008      & 0.118      & 0.04 \\
        \bottomrule
    \end{tabular}
\end{table}

\bibliography{pnas-sample}

\end{document}

### 2025-02-07 Main_text.tex
Uploaded SI_Appendix

--- 2025-02-07 Research_report.tex
@@
 We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
-\ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}, Pacific Herring and Jack Mackerel).
 For other high-biomass species otherwise lacking well-described distributions, we provide the first species distribution models (Fig.
-\ref{fig:fig1}, \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}).
@@
 Below we highlight results for six species (Fig.
-\ref{fig:fig1}) representing a range of habitat associations and general species information; distribution maps for all the remaining species are presented in \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1}) representing a range of habitat associations and general species information; distribution maps for all the remaining species are presented in \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}.
@@
 \ref{fig:fig1}).
-eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI.pdf}{SI Appendix, Fig.
+eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
@@
 Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat.
-We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI.pdf}{SI Appendix, Fig.
+We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
@@
 Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
-The eDNA concentrations appeared to reflect this activity (\href{SI.pdf}{SI Appendix, Fig.
+The eDNA concentrations appeared to reflect this activity (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
@@
 Also consistent with known habitats are commercially caught flatfish such as Dover sole (\textit{Microstomus pacificus}), which lives at 200m to 1200m along the Pacific coast \cite{drazen2007,brodziak2000,drazen2012}; individuals move progressively deeper as they grow \cite{vetter1994,hunter1990}.
-The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI.pdf}{SI Appendix, Fig.
+The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}\textit{A}) \cite{ono2016}.
@@
 Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$) with both more common in the more southern (and somewhat warmer) waters (Fig.
-\ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}).
 Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig.
-\ref{fig:fig2}) consistent with more northerly distribution (\href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig2}) consistent with more northerly distribution (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}).
@@
 Deep-water species are distinct (Fig.
-\ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}), not strongly correlated with those from other zones (Fig.
@@
 Deeper water tends to feature lower and more diffuse fish eDNA concentrations (Fig.
-\ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}).
@@
 In addition to providing species-specific patterns, eDNA metrics of abundance from individual species can be combined to reveal important aggregate ecological patterns.
-For the 12 species considered here, we show how species richness (see \href{SI.pdf}{SI Appendix, Fig.
+For the 12 species considered here, we show how species richness (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S2} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
@@
 \caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m).
-(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI.pdf}{SI Appendix, Fig.
+(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S2} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
@@
 \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station.
-Notably, the comparison between sample-specific eDNA estimation (\href{SI.pdf}{SI Appendix, Fig.
+Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S6}) and spatially smoothed maps (Fig.
-\ref{fig:fig1} and \href{SI.pdf}{SI Appendix, Fig.
+\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound.
@@
 
-%(see \href{SI.pdf}{SI Appendix, Fig. S5}) posterior predictive
+%(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S5}) posterior predictive
 
@@
 
-\matmethods{We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI.pdf}{SI Appendix}.
+\matmethods{We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
 
@@
 In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
-Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI.pdf}{SI Appendix} for details on metabarcoding methods).
+Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix} for details on metabarcoding methods).
 
@@
 Samples with fewer total number of reads than 1000 were filtered as indicator of poor reliability.
-Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI.pdf}{SI Appendix, Fig.
-S3}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI.pdf}{SI Appendix}).
-Details on bioinformatic pipeline are available in \href{SI.pdf}{SI Appendix}.
+Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S3}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI_Appendix.pdf}{SI Appendix}).
+Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
 
@@
 Latent parameters, state variables, data and subscripts are indicated in gray.
-See \href{SI.pdf}{SI Appendix} for full model breakdown.}
+See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
 \label{fig:DAG}
@@
 Vouchered DNA extracts or tissues were obtained from either the University of Washington Fish Collection at the Burke Museum or the NOAA Northwest Fisheries Science Center collections.
-After quantifying the concentration of mitochondrial DNA template in each DNA extract, we constructed a total of eight mock communities of varying species compositions (see \href{SI.pdf}{SI Appendix, Table S1}).
+After quantifying the concentration of mitochondrial DNA template in each DNA extract, we constructed a total of eight mock communities of varying species compositions (see \href{SI_Appendix.pdf}{SI Appendix, Table S1}).
 We selected 12 species (additional to Pacific hake) of commercial and ecological importance that co-occurred in the environmental samples and in the mock community for downstream analysis: Pacific herring (\textit{Clupea pallasii}), northern anchovy (\textit{Engraulis mordax}), northern smoothtongue (\textit{Leuroglossus stilbius}), Dover sole (\textit{Microstomus pacificus}), Pacific sardine (\textit{Sardinops sagax}), Pacific chub mackerel (\textit{Scomber japonicus}), widow rockfish (\textit{Sebastes entomelas}), northern lampfish (\textit{Stenobrachius leucopsarus}), California barracudina (\textit{Tactostoma macropus}), plainfin midshipman (\textit{Tarletonbeania crenularis}), eulachon (\textit{Thaleichthys pacificus}), Pacific jack mackerel (\textit{Trachurus symmetricus}).
@@
 To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples.
-The full statistical details are described in \href{SI.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig.
+The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig.
 \ref{fig:DAG}).
@@
 In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample.
-But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI.pdf}{SI Appendix Fig.
+But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig.
 S3}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
@@
 To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI.pdf}{SI Appendix Fig.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
 S4}) [Ryan check the framing of the last two sentences RRR].
@@
 \ref{fig:DAG}, green box).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI.pdf}{SI Appendix}).
+Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
 

### 2025-02-07 SI_Appendix.tex
Uploaded SI_Appendix

--- 2025-02-07 SI.tex
@@
-\documentclass[9pt,twoside,lineno]{pnas-new}
+\documentclass[9pt,twoside,lineno]{pnas-SI}
 % Use the lineno option to display guide line numbers if required.
@@
 \subsection*{qPCR}
-Environmental samples were analysed in triplicate using the multiplexed assay targeting Pacific hake, Eulachon and Pacific lamprey as described in \cite{ramon-laca2021} on a QuanStudio 6 (Applied Biosystems).
-Only Pacific hake was used in this study due to the remaining targets having little positive amplification.
+Environmental samples were analyzed in triplicate using the multiplexed assay targeting Pacific hake, Eulachon and Pacific lamprey as described in \cite{ramon-laca2021} on a QuanStudio 6 (Applied Biosystems).
+Only Pacific hake qPCR data was used in this study due to the remaining qPCR targets having little positive amplification.
 The samples underwent real-time thermocycler protocol including an initial denaturation step at 95°C 10 min followed by 40 cycles of 15 s at 95°C and 1 min at 58°C.
@@
 To monitor quality control, all runs included an internal positive control (IPC) to detect PCR inhibition and a negative template control (NTC) containing water instead of DNA template.
-Any IPC delay exceeding 0.5 cycles in the NTC was considered inhibition.
+Any IPC delay exceeding 0.5 cycles in the NTC was considered inhibition and inhibited samples were diluted and re-analyzed.
 
@@
 \subsection*{Metabarcoding}
-In total, 568 samples, including 554 environmental samples, 7 PCR blanks and 7 positive controls (holding only kangaroo DNA) were amplified were amplified using MiFish-U universal primers \cite{miya2015} with Illumina tails (forward TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGGCCGGTAAAACTCGTGCCAGC; reverse GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCATAGTGGGGTATCTAATCCCAGTTTG) by using a two-step PCR protocol.
-The DNA was amplified in the first PCR reaction (PCR 1) in a 20 µL reaction consisting of: 10 µL of Phusion Master Mix (2X), 0.4 µL of the forward primer (10 µM); 0.4 µL of the reverse primer (10 µM), 0.6 µL of 100\% DMSO, 0.5 µL of rAlbumin (20 µg/µL), 4.4 µL of nuclease-free water and 2 µL of DNA template.
-Reactions were run with the following cycling conditions: an initial denaturation of 98°C for 30 sec; followed by 35 cycles of 98°C for 10 sec, 60°C for 30 sec, and of 72°C for 3 sec; with a final extension of 72°C for 10 min.
+In total, 568 samples, including 554 environmental samples, 7 PCR blanks and 7 positive controls (holding only kangaroo DNA) were amplified using MiFish-U universal primers \cite{miya2015} with Illumina tails (forward TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGGCCGGTAAAACTCGTGCCAGC; reverse GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCATAGTGGGGTATCTAATCCCAGTTTG) by using a two-step PCR protocol.
+The DNA was amplified in the first PCR reaction (PCR1) in a 20 µL reaction consisting of: 10 µL of Phusion Master Mix (2X), 0.4 µL of the forward primer (10 µM); 0.4 µL of the reverse primer (10 µM), 0.6 µL of 100\% DMSO, 0.5 µL of rAlbumin (20 µg/µL), 4.4 µL of nuclease-free water and 2 µL of DNA template.
+Reactions were run with the following cycling conditions: an initial denaturation of 98°C for 30 sec; followed by 35 cycles of 98°C for 10 sec, 60°C for 30 sec, and of 72°C for 3 sec; with a final extension of 72°C for 10 min and hold at 4°C.
 
-PCR product was cleaned using Ampure Beads (1.2x) and then indexed in a second PCR reaction (PCR 2) consiting of 8 cycles.
+PCR product was cleaned using Ampure Beads (1.2x) and then indexed in a second PCR reaction (PCR2), with the following recipe: 12.5 $\mu$L of KAPA HiFi HotStart ReadyMix (Roche Diagnostics), 1.25 $\mu$L of an index from IDT for Illumina DNA/RNA UD Indexes (Sets A-D), 5 $\mu$L of PCR1 product, and 6.25 $\mu$L of nuclease free water.
+Cycling conditions included an initial denaturation of 95°C for 5 min; 8 cycles of: 98°C for 20 sec, 56°C for 30 sec, 72°C for 1 min; and a final extension of 72°C for 5 min.
 Resulting products were visualized on a 2\% agarose gel and quantified using Quant-iT™ dsDNA Assay Kit (Thermo Fisher Scientific, USA) with Fluoroskan™ Microplate Fluorometer (Thermo Fisher Scientific, USA).
-Indexed products were pooled into libraries for sequencing, and then size- selected to extract only the target fish band using the E-Gel™ SizeSelect™ II Agarose Gels (Thermo Fisher Scientific, USA).
+Indexed products were normalized by concentration, pooled into libraries for sequencing, and then size- selected to extract only the target fish band using the E-Gel™ SizeSelect™ II Agarose Gels (Thermo Fisher Scientific, USA).
 Subsequently the libraries were sequenced on Illumina MiSeq platform using the v3 600 cycle kit.
@@
 
-We quantified the total genomic DNA in each extract using Qubit HS assay.
+We constructed mock communities with even and skewed species total genomic DNA using Qubit HS assay quantification method.
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly match the template and give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
@@
 \subsection*{Bioinformatics}
-PEDRO
+Starting with Illumina-demultiplexed sequencing files, we first removed primer and adapter sequences using Cutadapt v4.9, applying the default mismatch tolerance parameter (-e 0.1) \cite{martin2011}.
+Next, we performed quality filtering using the filterAndTrim function from the dada2 R package, retaining its default parameters while specifying the optional truncLen argument to match the minimum MiFish amplicon length.
+After filtering, we denoised the sequences using dada2 with default settings \cite{callahan2016}, which resulted in a list of Amplicon Sequence Variants (ASVs) across all samples.
+Each resulting ASV was then assigned a unique 28-character hash code using the sha-1 algorithm with the R function map\_chr.
+This allowed us to create a common database of ASVs by hash, enabling us to track which sequences have already been annotated for consistency and to avoid redundant computational efforts.
+
+For taxonomic assignment of ASVs, we used standalone BLAST \cite{altschul1990} to compare each unique ASV against the NCBI nr eukaryote database as of January 2025.
+BLAST arguments were as follows: "-word\_size 30 -evalue "1e-30" -max\_target\_seqs 50".
+To streamline annotation, we also used the "-negative\_taxids" argument, including a list of target organisms closely related to those present in the sampled region, but known to be absent.
+The complete list of taxids that were excluded is available alongside the pipeline code.
+Finally, we performed a least common ancestor (LCA) analysis on the top BLAST hits using TaxonKit v0.17 \cite{shen2021}.
+Sequences with a 100 percent identity match to any database sequence were prioritized for LCA.
+In the absence of perfect matches, we conducted LCA analysis on the top hits with identity scores above 99.3, , 98, or 96 percent respectively.
+In the absence of any hits above 96 identity, taxonomic assignment was truncated.
+Finally, for mock communities were identify of species was known, we manually curated annotations before analysis.
 
@@
 OLE
+
+The above statistical model provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
+We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as:
+$$\ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)$$
+where x,y,z index longitude, latitude, and depth, respectively, $\tau$ is scalar intercept, $\eta(s)$ is a spatial smooth shared among depths, and $\varepsilon_d(s)$ is a depth-specific spatial smooth.
+Both $\eta(s)$ and $\varepsilon_d(s)$ are a zero-centered Gaussian random field estimated using the stochastic partial differntial equation (SPDE) approximation (see \cite{anderson2022}).
+Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model.
 
@@
 \centering
-\includegraphics[width=0.99\textwidth]{plots/5. Supplementary Figure 1.jpg}
+\includegraphics[width=0.99\textwidth]{plots/5_Supplementary_Figure_1.jpg}
 \caption{First figure}
@@
 \centering
-\includegraphics[width=0.99\textwidth]{plots/6. Supplementary Figure 2.jpg}
+\includegraphics[width=0.99\textwidth]{plots/6_Supplementary_Figure_2.jpg}
 \caption{First figure}
@@
 \centering
-\includegraphics[width=0.99\textwidth]{plots/7. Supplementary Figure 3.jpg}
+\includegraphics[width=0.99\textwidth]{plots/7_Supplementary_Figure_3.jpg}
 \caption{First figure}
@@
 \centering
-\includegraphics[width=0.99\textwidth]{plots/8. Supplementary Figure 4.jpg}
+\includegraphics[width=0.99\textwidth]{plots/8_Supplementary_Figure_4.jpg}
 \caption{First figure}
@@
 \centering
-\includegraphics[width=0.80\textwidth]{plots/9. Supplementary Figure 5.jpg}
+\includegraphics[width=0.80\textwidth]{plots/9_Supplementary_Figure_5.jpg}
 \caption{First figure}
@@
 \centering
-\includegraphics[width=0.89\textwidth]{plots/10. Supplementary Figure 6.jpg}
+\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
 \caption{First figure}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.89\textwidth]{plots/4_Figure_4.jpg}
+\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
+Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A).
+Probability density functions (PDF) of estimated eDNA concentrations across latitudinal gradients (B; square root transformed of x-axis to enhance low concentration peaks).
+Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
 \end{figure}

### 2025-02-07 DAG.tex
Updates from Overleaf

\documentclass[landscape]{article}
\usepackage[paperwidth=292mm, paperheight=129mm, margin=0cm]{geometry}
\usepackage{tikz}
\usepackage{xcolor}
\definecolor{myBlue}{RGB}{173, 216, 230}
\definecolor{myGrey}{RGB}{200, 200, 200}
\usetikzlibrary{positioning, shapes.geometric, arrows}

% Define styles
\tikzstyle{node} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=1cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=myBlue] at (0,0) {
    $\begin{array}{c}
        Z_{ik} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot K_{ik} \cdot \phi) \right) \\
        Y_{ik} \sim \mathrm{Normal} (\mu_{ik}, \sigma_{ik}) \quad \textnormal{if } Z_{ik} = 1 \\
        \mu_{ik} = \beta0_{ip} + \beta1_{ip} \cdot \ln K_{ik} \\
        \sigma_{ik} = \exp(\gamma0 + \gamma1 \cdot \ln K_{ik})
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standard model};
\node (qPCR_en) [node, right=of qPCR_st, fill=myBlue]  at (4,0) {
    $\begin{array}{c}
        Z_{ijdbtr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot C_{ijdbt} \cdot \phi) \right) \\
        Y_{ijdbtr} \sim \mathrm{Normal} (\mu_{ijdbt}, \sigma_{ijdbt}) \quad \textnormal{if } Z_{ijr} = 1 \\
        \mu_{ijdbt} = \beta0_{ip} + \beta1_{ip} \cdot \ln C_{ijdbt} \\
        \sigma_{ijdbt} = \exp(\gamma0 + \gamma1 \cdot \ln C_{ijdbt})
    \end{array}$
};
\node [above=of qPCR_en, yshift=-1.5cm, font=\large\bfseries] {qPCR Unknown model};
\node (covariates) [node, below=of qPCR_en, fill=myBlue] { 
% + \delta0_{ijdbt} \cdot
% + \delta1_{ijdb}
    $\begin{array}{c}
        \ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        \ln D_{ijdb} = \psi_i + \ln D_{ijd} + \rho_{ijdbt}
    \end{array}$
};
\node [above=of covariates, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {Covariates};
\node (metabarcoding) [node, below=of covariates, fill=myBlue] {
    $\begin{array}{c}
        R_{ijdb} \sim \mathrm{Multinomial} \left( \pi_{ijdb}, T_{jdb} \right) \\[8pt]
        \pi_{ijdb} = \frac{e^{\nu_{ijdb}}}{\sum_{i=1}^{I} e^{\nu_{ijdb}}} \\[8pt]
        \nu_{ijdb} = (\ln D_{ijdb} - \ln D_{jdb, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of metabarcoding, yshift=-1.5cm, xshift=-2.5cm, font=\large\bfseries] {Metabarcoding model};
\node (mock) [node, right=of metabarcoding, fill=yellow!50] {
    $\begin{array}{c}
        R_{im} \sim \mathrm{Multinomial} \left( \pi_{im}, T_{m} \right) \\[8pt]
        \pi_{im} = \frac{e^{\nu_{im}}}{\sum_{i=1}^{I} e^{\nu_{im}}} \\[8pt]
        \nu_{im} = (\ln D_{im} - \ln D_{m, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock metabarcoding model};
\node (spatial) [node, below=of metabarcoding, fill=green!40] {
    $\begin{array}{c}
        \ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)
    \end{array}$
};
\node [above=of spatial, yshift=-1.5cm, xshift=-1.6cm, font=\large\bfseries] {Spatial model};

        % Arrows
        \draw [arrow] (qPCR_st) -- node[midway, above] {
        $\begin{array}{c}
            \phi \\ \beta0, \beta1 \\ \gamma0, \gamma1
        \end{array}$} (qPCR_en);
        \draw [arrow] (qPCR_en) -- node[midway, right] {$\ln C_{ijdbt}$} (covariates);
        \draw [arrow] (covariates) -- node[midway, right] {$\ln D_{ijdb}$} (metabarcoding);
        \draw [arrow] (mock) -- node[midway, above] {$\alpha_i$} (metabarcoding);
        \draw [arrow] (metabarcoding) -- node[midway, right] {$\ln D_{ijdb}$} (spatial);

% Legend: Latent Parameters
\node (parameters) [rectangle, draw, align=left, right=of qPCR_en, text width=12cm, fill=myGrey] at (11.9,-0.8){
    \textbf{Latent parameters} \\
    $\psi$ = mean hake eDNA concentration \\
    %$\delta0, \delta1$ = offset from mean hake \\
    $\rho$ = biological replicate random effect \\
    $\omega$ = ethanol wash effect \\
    $\xi$ = volume and dilution offset \\
    $\mu$ = mean Ct values of qPCR \\
    $\phi$ = probability of positive qPCR amplification \\
    $\sigma$ = standard deviation of qPCR Ct values \\
    $\beta0, \beta1$ = intercept and slope between $\mu$ and eDNA concentration (K and C) \\
    $\gamma0, \gamma1$ = intercept and slope between $\sigma$ and eDNA concentration (K and C) \\
    $\alpha$ = amplification efficiency relative to reference species \\
    $\pi$ = sequence proportions at the end of PCR [0,1] \\
    $\nu$ = sequence proportions at the end of PCR \\
    $\tau$ = scalar intercept\\
    $\eta$ = spatial field \\
    $\varepsilon$ =  depth specific spatial field
};
% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_st, text width=7cm, fill=myGrey] at (0.4,0) {
    \textbf{Subscripts} \\
    $i$ = species \\
    $j$ = station \\
    $d$ = depth \\
    $b$ = biological replicate \\
    $t$ = aliquot replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community aliquote sample \\
    $k$ = qPCR standard aliquote sample
};
% Legend: Data
\node (data) [rectangle, draw, align=left, below=of subscripts, text width=7cm, fill=myGrey] at (0.4,-5.6){
    \textbf{Data} \\
    $Z$ = qPCR amplification yes/no (0,1) \\
    $Y$ = qPCR Ct values \\
    $K$ = Known concentration in copies/$\mu$L \\
    $R$ = metabarcoding sequencing reads \\
    $T$ = total number sequencing reads
};
% Legend: State variables
\node (state) [rectangle, draw, align=left, right=of qPCR_en, text width=6cm, fill=myGrey] at (11.9,-9){
    \textbf{State variables} \\
    $C$ = unknown eDNA concentration \\
    $D$ = unknown eDNA concentration
};
    \end{tikzpicture}
    %\caption{Multifish DAG}
\end{figure}

\end{document}

### 2025-02-07 Main_text.tex
Updates from Overleaf

@@
 % Use the lineno option to display guide line numbers if required.
+\usepackage{soul} % Delete this line before submission XXX
 
@@
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
 \author[a,1]{Gledis Guri}
@@
 
-\affil[a]{University of Washington}
-\affil[b]{NOAA fisheries science}
+\affil[a]{School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA}
+\affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}
 
@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
-The significance statement appears in the paper itself and is required for all research papers.}
+\significancestatement{Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales.
+We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States.
+The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to broadly useful across natural-resources management and the life sciences more generally.}
 
@@
 MS and PB did the laboratory and the bioinformatic work;
-GG OL RPK AOS analysed the data;
-GG OL RPK AOS analysed interpreted the results;
-GG OL RPK AOS analysed wrote the manuscript}
+GG, OL, RPK, and AOS analysed the data;
+GG, OL, RPK, and AOS analysed interpreted the results;
+GG, OL, RPK, and AOS analysed wrote the manuscript
+\hl{[Please check that this is correct and add/edit as needed XXX]}
+}
 \authordeclaration{The authors declare no competing interest.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: gguri@uw.edu}
+
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: gguri@uw.edu}
 
@@
 Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}.
-However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2024}).
+However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2022a}).
 While the use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
@@
 \ref{fig:fig3}).
-Particularly for forage fish species – which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012LittleFB} – molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
+Particularly for forage fish species – which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} – molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
 
@@
 eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
+S2}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
 Furthermore, eDNA abundance patterns aligned with stock assessments in indicating a higher abundance of northern anchovy; however, the magnitude of difference varied, with eDNA suggesting a 2.5-fold difference, while stock assessments indicated a 35-fold difference in abundance.
-[CITE stock assessment OOO]
+\hl{[CITE stock assessment here; OLE XXX]}
 
 \subsection*{Mid-water}
-Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} [ADD SCUBA REF OOO].
+Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} \hl{[ADD SCUBA REF; OLE XXX]}.
 Moreover, because it is a common bycatch species in the hake (\textit{Merluccius productus}) fishery, limits on bycatch of widow rockfish can also limit the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
@@
 We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
+S2}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
 
@@
 The eDNA concentrations appeared to reflect this activity (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
+S2}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
 \ref{fig:fig1}), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
@@
 The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}\textit{A}) \cite{ono2016}.
+S2}\textit{A}) \cite{ono2016}.
 
@@
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}).
+S2}).
 Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig.
 \ref{fig:fig2}) consistent with more northerly distribution (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}).
+S2}).
 Together these finding suggest that eDNA is capturing both depth and latitudinal habitats associations.
@@
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}), not strongly correlated with those from other zones (Fig.
+S2}), not strongly correlated with those from other zones (Fig.
 \ref{fig:fig2}).
@@
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}).
+S2}).
 Lower overall eDNA concentrations and fewer deep water samples may make similarities among species distributions more difficult to detect at 300m and deeper.
@@
 \end{figure}
-
-%\begin{SCfigure*}[\sidecaptionrelwidth][t!]
-%\centering
-%\includegraphics[width=11.4cm,height=11.4cm]{frog.pdf}
-%\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
-%\end{SCfigure*}
 
@@
 For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa, revealing unexpected habitats.
-Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys [NWFSC trawl survey ref OOO].
+Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \hl{[NWFSC trawl survey ref; OLE XXX]}.
 This species has been of special interest, as its targeted commercial harvest resumed in the late 2010s after a decade of closure (harvest was less than 500 mt annually between 2003 and 2013 but has exceeded 10,000 mt annually since 2018) and no existing survey adequately captures its patterns of abundance \cite{adams2019a}.
@@
 
-The majority of eDNA shed into the environment dissipates from surface waters within hours (Brasseale et al., in review RRR).
+The majority of eDNA shed into the environment dissipates from surface waters within hours \hl[(Brasseale et al., in review); add this reference; RYAN or OLE XXX].
 Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most.
@@
 Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S6}) and spatially smoothed maps (Fig.
+S7}) and spatially smoothed maps (Fig.
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound.
-This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.[Ryan check the framing of the last two sentences RRR]
+This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.
+\hl{[Check the framing of the last two sentences XXX]}
 
@@
 
-%(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S5}) posterior predictive
+%(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S6}) posterior predictive
 
@@
 Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes.
-Here, we follow \cite{shelton2022} and Liu et al.
-(submitted) RRR, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
-Elsewhere, the same set of samples has yielded information on the distribution of marine mammals (Valdivia et al, in revision) RRR.
+Here, we follow \cite{shelton2022} and \hl{[Liu et al.
+(submitted); add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
+Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision); add this reference; RYAN or OLE XXX]}.
 The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa.
@@
 Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S3}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI_Appendix.pdf}{SI Appendix}).
+S4}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI_Appendix.pdf}{SI Appendix}).
 Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
@@
 \centering
-\includegraphics[width=17.4cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
 \caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
@@
 But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S3}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
+S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
 Our model generates estimates of DNA concentration (copies/L) for each species.
@@
 However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}) [Ryan check the framing of the last two sentences RRR].
-Samples lacking hake detections were sporadically distributed across the study area, and their omission appears not to have induced substantial bias in our estimates.
+S5}).
+\hl{[Check
+the framing of the last two sentences XXX]}Samples lacking hake detections were sporadically distributed across the study area, and their omission appears not to have induced substantial bias in our estimates.
 
@@
 
-\acknow{We thank Linda Park for her foresight and perseverance, and Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis.
+\acknow{We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
 This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
@@
 
-\bibsplit[11]
+\bibsplit[9]
 %Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)

### 2025-02-07 SI_Appendix.tex
Updates from Overleaf

@@
 We constructed mock communities with even and skewed species total genomic DNA using Qubit HS assay quantification method.
-We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly match the template and give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
+We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
 Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water.
@@
 \centering
-\includegraphics[width=0.99\textwidth]{plots/5_Supplementary_Figure_1.jpg}
-\caption{First figure}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.99\textwidth]{plots/6_Supplementary_Figure_2.jpg}
-\caption{First figure}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.99\textwidth]{plots/7_Supplementary_Figure_3.jpg}
-\caption{First figure}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.99\textwidth]{plots/8_Supplementary_Figure_4.jpg}
-\caption{First figure}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.80\textwidth]{plots/9_Supplementary_Figure_5.jpg}
-\caption{First figure}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
-\caption{First figure}
+\includegraphics[width=0.98\textwidth]{plots/5_Supplementary_Figure_1.jpg}
+\caption{Continuation of Fig.
+1.
+Estimated species eDNA for \textit{Clupea pallasi}, \textit{Trachurus symmetricus}, \textit{Scomber japonicus}, \textit{Thaleichthys pacificus}, \textit{Leuroglossus stilbius}, and \textit{Tarletonbaeania crenularis} concentration (A) across 0, 50, 150, 300, and 500m depth samples and known depth distribution of those species from literature (B).}
 \end{figure}
@@
 
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.99\textwidth]{plots/6_Supplementary_Figure_2.jpg}
+\caption{Frequency distribution of eDNA concentration for each (n = 12) species included in this study.
+The dashed red line indicates the threshold model estimated concentration (translated copies/L) for defining presence and absence of sequenced (through metabarcoding) taxa}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.99\textwidth]{plots/7_Supplementary_Figure_3.jpg}
+\caption{Three-dimentional distribution of samples collected that were run through qPCR (ligh blue), metabarcoding (dark blue), and were used jointly in the Bayesian model (Fig. 4; red)}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.99\textwidth]{plots/8_Supplementary_Figure_4.jpg}
+\caption{Probability of species co-occurrence with Pacific hake, where blue bars represent non-occurrence and orange bars indicate co-occurrence.
+\textit{Engraulis mordax} exhibits the most inverse habitat overlap with hake, while \textit{Stenobrachius leucopsarus} shows the highest habitat overlap.}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.80\textwidth]{plots/9_Supplementary_Figure_5.jpg}
+\caption{Posterior predictes of three model compartments, qPCR continuous model (top part), qPCR occupancy model (middle part), and metabarcoding model (bottom part)}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
+\caption{Un-smoothed estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples.}
+\end{figure}
+
 \begin{table}\centering
-\caption{Initial relative abundances of the selected species across different mock communities.}
+\caption{Initial relative abundances (measured by amplifying the 12S rRNA gene using MarVer1 primers \cite{valsecchi2020} and ddPCR QX200 Droplet Digital PCR system (Bio-Rad, Inc.)) of the selected species (n=12) across different mock communities (n=8).
+NA represent no DNA template included.}
     \begin{tabular}{lcccccccc}

### 2025-02-08 DAG.tex
Updates from Overleaf

@@
 \node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standard model};
+\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_st.south east) {\textbf{(2)}};
+
 \node (qPCR_en) [node, right=of qPCR_st, fill=myBlue]  at (4,0) {
@@
 \node [above=of qPCR_en, yshift=-1.5cm, font=\large\bfseries] {qPCR Unknown model};
+\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_en.south east) {\textbf{(3)}};
+
 \node (covariates) [node, below=of qPCR_en, fill=myBlue] { 
@@
 \node [above=of covariates, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {Covariates};
+\node[anchor=south east] at ([xshift=2pt, yshift=-15pt] covariates.south east) {\textbf{(4)}};
+
 \node (metabarcoding) [node, below=of covariates, fill=myBlue] {
@@
 \node [above=of metabarcoding, yshift=-1.5cm, xshift=-2.5cm, font=\large\bfseries] {Metabarcoding model};
+\node[anchor=south east] at ([xshift=4pt, yshift=-15pt] metabarcoding.south east) {\textbf{(5)}};
+
 \node (mock) [node, right=of metabarcoding, fill=yellow!50] {
@@
 };
-\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock metabarcoding model};
+\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock-community metabarcoding model};
+\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] mock.south east) {\textbf{(1)}};
+
 \node (spatial) [node, below=of metabarcoding, fill=green!40] {
@@
 \node [above=of spatial, yshift=-1.5cm, xshift=-1.6cm, font=\large\bfseries] {Spatial model};
+\node[anchor=south east] at ([xshift=19pt, yshift=-4pt] spatial.south east) {\textbf{(6)}};
 
@@
     $\omega$ = ethanol wash effect \\
-    $\xi$ = volume and dilution offset \\
+    $\xi$ = volume filtered and aliquote dilution offset \\
     $\mu$ = mean Ct values of qPCR \\
@@
     $\alpha$ = amplification efficiency relative to reference species \\
-    $\pi$ = sequence proportions at the end of PCR [0,1] \\
-    $\nu$ = sequence proportions at the end of PCR \\
+    $\pi$ = sequence proportions at the end of PCR (proportion [0,1]) \\
+    $\nu$ = sequence proportions at the end of PCR (additive log-ratio)\\
     $\tau$ = scalar intercept\\
@@
 };
+
 % Legend: Subscripts
-\node (subscripts) [rectangle, draw, align=left, below=of qPCR_st, text width=7cm, fill=myGrey] at (0.4,0) {
+\node (subscripts) [rectangle, draw, align=left, below=of qPCR_st, text width=7cm, fill=myGrey] at (0.4,-0.7) {
     \textbf{Subscripts} \\
@@
     \textbf{Data} \\
-    $Z$ = qPCR amplification yes/no (0,1) \\
+    $Z$ = qPCR amplification (yes=1; no=0) \\
     $Y$ = qPCR Ct values \\
@@
 % Legend: State variables
-\node (state) [rectangle, draw, align=left, right=of qPCR_en, text width=6cm, fill=myGrey] at (11.9,-9){
+\node (state) [rectangle, draw, align=left, right=of qPCR_en, text width=7.5cm, fill=myGrey] at (11.9,-9){
     \textbf{State variables} \\
-    $C$ = unknown eDNA concentration \\
-    $D$ = unknown eDNA concentration
+    $C$ = unknown eDNA concentration (copies/$\mu$L) \\
+    $D$ = unknown eDNA concentration (copies/L)
 };

### 2025-02-08 Main_text.tex
Updates from Overleaf

@@
 \author[a]{Ryan P. Kelly}
-\author[a]{Megan Shaffer}
+\author[a]{Megan R. Shaffer}
 \author[b]{Kim Parsons}
+\author[a,b,c]{Ana Ram\'on-Laca}
 \author[b]{Krista M. Nichols}
@@
 \affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}
+\affil[c]{Museo Nacional de Ciencias Naturales, Madrid, Spain}
 
@@
 We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States.
-The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to broadly useful across natural-resources management and the life sciences more generally.}
+The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to be broadly useful in natural-resources management and in the life sciences more generally.}
 
@@
 AOS, KMN, RPK, MS, and KP designed the research;
-MS and PB did the laboratory and the bioinformatic work;
-GG, OL, RPK, and AOS analysed the data;
-GG, OL, RPK, and AOS analysed interpreted the results;
-GG, OL, RPK, and AOS analysed wrote the manuscript
+MRS and AR-L did the laboratory work; 
+PB, RPK, and GG did the bioinformatics;
+GG, OL, RPK, and AOS analyzed the data;
+GG, OL, RPK, and AOS interpreted the results;
+GG, OL, RPK, and AOS wrote the manuscript
 \hl{[Please check that this is correct and add/edit as needed XXX]}
@@
 \begin{abstract}
-Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural resources management at a broad geographic scale.
+Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural-resources management at a broad geographic scale.
 These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously.
@@
 
-The majority of eDNA shed into the environment dissipates from surface waters within hours \hl[(Brasseale et al., in review); add this reference; RYAN or OLE XXX].
+The majority of eDNA shed into the environment dissipates from surface waters within hours \hl{(Brasseale et al., in review)[add this reference; RYAN or OLE XXX]}.
 Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most.
@@
 Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes.
-Here, we follow \cite{shelton2022} and \hl{[Liu et al.
-(submitted); add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
-Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision); add this reference; RYAN or OLE XXX]}.
+Here, we follow \cite{shelton2022} and \hl{Liu et al.
+(submitted)[add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
+Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision)[add this reference; RYAN or OLE XXX]}.
 The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa.
@@
 \caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
-Model compartment indicated in the same color are run jointly.
+Model compartment indicated in the same color (2-5) are run jointly.
 Latent parameters, state variables, data and subscripts are indicated in gray.
@@
 To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurance, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
 S5}).
-\hl{[Check
-the framing of the last two sentences XXX]}Samples lacking hake detections were sporadically distributed across the study area, and their omission appears not to have induced substantial bias in our estimates.
+Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
+S4}), and their omission appears not to have induced substantial bias in our estimates.
 

### 2025-02-08 SI_Appendix.tex
Updates from Overleaf

@@
 \title{Quantitative, Multispecies Monitoring at a Continental Scale}
-\author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Krista M. Nichols, Pedro F. P. Brandão-Dias and Andrew Olaf Shelton}
+\author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Ana Ram\'on-Laca,\\
+Krista M. Nichols, Pedro F. P. Brandão-Dias and Andrew Olaf Shelton}
 \correspondingauthor{Gledis Guri.\\E-mail: gguri@uw.edu}
@@
 Subsequently the libraries were sequenced on Illumina MiSeq platform using the v3 600 cycle kit.
-
-%A total of 554 environmental samples of metabarcoding samples sequenced are 554 of which 535 had higher sequencing depth than 1000 reads (indicating good quality samples).
-From those we indicated that only 371 samples had Pacific hake (\textit{Merluccius productus}) present jointly with qPCR positive detection of the same species.
-All libraries produced a total of 56,279,345 reads with 54,285,713 reads (96.5\%) belonging to class \textit{Actinopteri}.
-A total of 222 unique species was identified from which 165 belonged to fish.
 
@@
 \includegraphics[width=0.99\textwidth]{plots/6_Supplementary_Figure_2.jpg}
-\caption{Frequency distribution of eDNA concentration for each (n = 12) species included in this study.
-The dashed red line indicates the threshold model estimated concentration (translated copies/L) for defining presence and absence of sequenced (through metabarcoding) taxa}
+\caption{Frequency distribution of eDNA concentration for each species included in this study (n = 12).
+The dashed red line indicates the threshold model estimated concentration (copies/L) for defining presence and absence of sequenced (through metabarcoding) taxa}
 \end{figure}
@@
 \includegraphics[width=0.99\textwidth]{plots/7_Supplementary_Figure_3.jpg}
-\caption{Three-dimentional distribution of samples collected that were run through qPCR (ligh blue), metabarcoding (dark blue), and were used jointly in the Bayesian model (Fig. 4; red)}
+\caption{Three-dimensional distribution of samples collected that were run through qPCR (light blue), metabarcoding (dark blue), and were used jointly in the Bayesian model (Fig.
+4; red).
+Jointly modeled samples are those that detected Pacific hake -- used as a reference species for quantification -- in both qPCR and metabarcoding assays.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.80\textwidth]{plots/9_Supplementary_Figure_5.jpg}
-\caption{Posterior predictes of three model compartments, qPCR continuous model (top part), qPCR occupancy model (middle part), and metabarcoding model (bottom part)}
+\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
+\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
 \end{figure}
@@
 \includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
-\caption{Un-smoothed estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples.}
+\caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
 \end{figure}
@@
 \begin{table}\centering
-\caption{Initial relative abundances (measured by amplifying the 12S rRNA gene using MarVer1 primers \cite{valsecchi2020} and ddPCR QX200 Droplet Digital PCR system (Bio-Rad, Inc.)) of the selected species (n=12) across different mock communities (n=8).
+\caption{Initial proportional abundances (measured by amplifying the 12S rRNA gene using MarVer1 primers \cite{valsecchi2020} and ddPCR QX200 Droplet Digital PCR system (Bio-Rad, Inc.)) of the selected species (n=12) across different mock communities (n=8).
 NA represent no DNA template included.}
@@
         \midrule
-        \textit{Clupea pallasii}           & 0.437 & 0.325  & 0.042  & 0.009     & NA         & NA          & 0.025      & 0.038 \\
+        \textit{Clupea pallasii}           & 0.437 & 0.325  & 0.042  & 0.009     & -         & -          & 0.025      & 0.038 \\
         \textit{Engraulis mordax}          & 0.122 & 0.034  & 0.15 & 0.15       & 0.146      & 0.16       & 0.308      & 0.478 \\
-        \textit{Leuroglossus stilbius}     & 0.108 & 0.29   & NA & NA         & NA         & NA         & NA         & NA \\    
+        \textit{Leuroglossus stilbius}     & 0.108 & 0.29   & - & -         & -         & -         & -         & - \\    
         \textit{Merluccius productus}      & 0.079 & 0.144  & 0.12 & 0.06       & 0.117      & 0.075      & 0.247      & 0.287 \\
-        \textit{Microstomus pacificus}     & NA    & NA     & 0.057  & 0.215     & NA         & NA         & NA         & NA \\    
+        \textit{Microstomus pacificus}     & -    & -     & 0.057  & 0.215     & -         & -         & -         & - \\    
         \textit{Sardinops sagax}           & 0.028 & 0.033  & 0.14 & 0.035      & 0.137      & 0.025      & 0.289      & 0.131 \\
-        \textit{Scomber japonicus}         & NA    & NA     & 0.122  & 0.053      & 0.119      & 0.027     & NA         & NA \\    
-        \textit{Sebastes entomelas}        & NA    & NA     & NA & NA          & 0.068      & 0.218     & NA         & NA \\    
-        \textit{Stenobrachius leucopsarus} & NA    & NA     & NA & NA          & 0.356      & 0.487     & NA         & NA \\    
-        \textit{Tactostoma macropus}       & NA    & NA     & 0.13 & 0.324     & NA         & NA         & NA         & NA \\    
-        \textit{Tarletonbeania crenularis} & NA    & NA     & 0.083  & 0.124     & NA         & NA         & NA         & NA \\    
-        \textit{Thaleichthys pacificus}    & 0.123 & 0.113  & 0.099  & 0.012     & NA         & NA          & 0.013      & 0.021 \\
+        \textit{Scomber japonicus}         & -    & -     & 0.122  & 0.053      & 0.119      & 0.027     & -         & - \\    
+        \textit{Sebastes entomelas}        & -    & -     & - & -          & 0.068      & 0.218     & -         & - \\    
+        \textit{Stenobrachius leucopsarus} & -    & -     & - & -          & 0.356      & 0.487     & -         & - \\    
+        \textit{Tactostoma macropus}       & -    & -     & 0.13 & 0.324     & -         & -         & -         & - \\    
+        \textit{Tarletonbeania crenularis} & -    & -     & 0.083  & 0.124     & -         & -         & -         & - \\    
+        \textit{Thaleichthys pacificus}    & 0.123 & 0.113  & 0.099  & 0.012     & -         & -          & 0.013      & 0.021 \\
         \textit{Trachurus symmetricus}     & 0.103 & 0.06   & 0.057  & 0.018      & 0.056      & 0.008      & 0.118      & 0.04 \\

### 2025-02-08 DAG.tex
Updates from Overleaf

@@
 \documentclass[landscape]{article}
-\usepackage[paperwidth=292mm, paperheight=129mm, margin=0cm]{geometry}
+\usepackage[paperwidth=292mm, paperheight=123mm, margin=0cm]{geometry}
 \usepackage{tikz}
@@
     $\begin{array}{c}
-        \ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
+        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
+        \ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
         \ln D_{ijdb} = \psi_i + \ln D_{ijd} + \rho_{ijdbt}
@@
 % Legend: Latent Parameters
-\node (parameters) [rectangle, draw, align=left, right=of qPCR_en, text width=12cm, fill=myGrey] at (11.9,-0.8){
+\node (parameters) [rectangle, draw, align=left, right=of qPCR_en, text width=12cm, fill=myGrey] at (11.9,-1.3){
     \textbf{Latent parameters} \\
@@
     $\rho$ = biological replicate random effect \\
-    $\omega$ = ethanol wash effect \\
+    %$\omega$ = ethanol wash effect \\
     $\xi$ = volume filtered and aliquote dilution offset \\
@@
 % Legend: Data
-\node (data) [rectangle, draw, align=left, below=of subscripts, text width=7cm, fill=myGrey] at (0.4,-5.6){
+\node (data) [rectangle, draw, align=left, below=of subscripts, text width=7cm, fill=myGrey] at (0.4,-5.4){
     \textbf{Data} \\
@@
     $R$ = metabarcoding sequencing reads \\
-    $T$ = total number sequencing reads
+    $T$ = total number sequencing reads \\
+    $W$ = ethanol wash effect
 };

### 2025-02-11 Main_text.tex
Updates from Overleaf

@@
 \usepackage{soul} % Delete this line before submission XXX
-
 \templatetype{pnasresearcharticle} % Choose template
@@
 \author[a]{Pedro F. P. Brandão-Dias}
+\author[b]{Abigail Wells}
 \author[b]{Andrew Olaf Shelton}
@@
 AOS, KMN, RPK, MS, and KP designed the research;
-MRS and AR-L did the laboratory work; 
+MRS, AR-L and AW did the laboratory work; 
 PB, RPK, and GG did the bioinformatics;
@@
 However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2022a}).
-While the use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
+The use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
 there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2024}, and coastal oceans \cite{maes2023, shelton2019}).
@@
 
-While trawling and other traditional sampling methods have been the fundamental tools for directly measuring fish abundance and biomass, they are often impractical for species inhabiting inaccessible habitats, occurring at low densities, or being fragile and easily injured during capture.
+While trawling and other traditional sampling methods have been the fundamental tools for directly measuring fish abundance and biomass, they are often impractical for species inhabiting inaccessible habitats, occurring at low densities, or fragile species easily injured during capture.
 By contrast, eDNA sampling captures information on a broader range of taxa, including those that are difficult or impossible to study with traditional gear.
@@
 \subsection*{Benthic}
-Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common yet little studied, such as the dragonfish (\textit{Tactostoma macropus}).
+Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common, yet under studied, such as the dragonfish (\textit{Tactostoma macropus}).
 Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
@@
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}), not strongly correlated with those from other zones (Fig.
+S2}) and not strongly correlated with those from other zones (Fig.
 \ref{fig:fig2}).
@@
 By mapping three-dimensional spatial patterns, eDNA provides information on species distributions, habitat preferences, and ecological processes that is often otherwise unavailable.
-A single water sample can yield quantitative information on tens to hundreds of species – including for species for which traditional fisheries surveys are not currently conducted – without the need to develop species-specific assays.
-We find species-level eDNA concentrations align closely the preferred habitats physiological constraints (e.g., temperature tolerance or food availability) described for each species in the available literature.
+A single water sample can yield quantitative information on tens to hundreds of species -- including for species for which traditional fisheries surveys are not currently conducted -- without the need to develop species-specific assays.
+We find species-level eDNA concentrations align closely with the preferred habitats' physiological constraints (e.g., temperature tolerance or food availability) described for each species in the available literature.
 The observed latitudinal differences of eDNA concentration among species – such as those between anchovy and sardine – highlights species’ thermal ranges and largely coincide with other surveys \cite{zwolinski2024}; moreover, eDNA patterns appear to reflect diel vertical migrations as observed in myctophids, where eDNA concentrations are consistent with the species’ ontogenetic movement between surface and mid- to deep-water habitats.
@@
 \subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
-Samples were collected in 10L Niskin bottles deployed on a CTD rosette at five depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship Bell Shimada during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey.
+Samples were collected in 10L Niskin bottles deployed on a CTD rosette at five depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
+Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey.
 2.5L water samples were collected from the Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.

### 2025-02-11 Main_text.tex
Updates from Overleaf

@@
 PB, RPK, and GG did the bioinformatics;
-GG, OL, RPK, and AOS analyzed the data;
+GG, OL, RPK, and eAOS analyzed the data;
 GG, OL, RPK, and AOS interpreted the results;

### 2025-02-11 SI_Appendix.tex
Updates from Overleaf

@@
 \author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Ana Ram\'on-Laca,\\
-Krista M. Nichols, Pedro F. P. Brandão-Dias and Andrew Olaf Shelton}
+Krista M. Nichols, Pedro F. P. Brandão-Dias, Abigail Wells and Andrew Olaf Shelton}
 \correspondingauthor{Gledis Guri.\\E-mail: gguri@uw.edu}
@@
 Only Pacific hake qPCR data was used in this study due to the remaining qPCR targets having little positive amplification.
-The samples underwent real-time thermocycler protocol including an initial denaturation step at 95°C 10 min followed by 40 cycles of 15 s at 95°C and 1 min at 58°C.
-All thermocycler reactions were run in 20 $\mu$L volume consisting of 10 $\mu$L of TaqMan Environmental Master Mix 2.0, 1 $\mu$L of each primer (forward: AAATGTTTAAACTAGAGCCGAATAGC and reverse: TCGTGGAGTCAAAGTGGGGTAGA, 10 nM concentration each), 0.5 $\mu$L of probe (AAATGTTTAAACTAGAGCCGAATAGC; 10nM concentration), 3 $\mu$L of dH20 and 2 $\mu$L of DNA template.
-To monitor quality control, all runs included an internal positive control (IPC) to detect PCR inhibition and a negative template control (NTC) containing water instead of DNA template.
-Any IPC delay exceeding 0.5 cycles in the NTC was considered inhibition and inhibited samples were diluted and re-analyzed.
 
-Alongside environmental samples standard samples were analyzed constructed from a 120 bp synthetic DNA fragment (gBlock; IDT) representing the 12S region of Pacific hake (\textit{Merluccius productus}), which encompasses the 101 bp 12S qPCR target (sequence information available in Ramón-Laca et al.
-2021).
+The samples underwent real-time thermocycler protocol including an initial denaturation step at 95°C 10 min followed by 45 cycles of 15 s at 95°C and 1 min at 60°C.
+All samples were run in triplicate in 10 $\mu$L volume consisting of 1 x TaqPath ProAmp Multiplex Master Mix, 0.9 $\mu$M of each primer (forward: AAATGTTTAAACTAGAGCCGAATAGC and reverse: TCGTGGAGTCAAAGTGGGGTAGA), 0.2 $\mu$M of probe (6FAM-CACTCGAGGCCACGAAGTACAATT-(MGB)NFQ), and 2 $\mu$L of DNA template or water for the non-template control.
+All reactions included 100 copies of internal positive control (IPC) to detect PCR inhibition.
+Any IPC delay exceeding 0.5 cycles in the non-template control was considered inhibition and inhibited samples were diluted and re-analyzed.
+
+Alongside environmental samples standard samples were analyzed constructed from a 130 bp synthetic DNA fragment (gBlock; IDT) representing the 12S region of Pacific hake (\textit{Merluccius productus}), which encompasses the 101 bp 12S qPCR target (sequence information available in \cite{ramon-laca2021}).
 The synthetic DNA fragment was diluted to create a series of standards with final concentrations ranging from $10^{0}$ to $10^{5}$ copies/$\mu$L.

### 2025-02-11 SI_Appendix.tex
Updates from Overleaf

@@
 The samples underwent real-time thermocycler protocol including an initial denaturation step at 95°C 10 min followed by 45 cycles of 15 s at 95°C and 1 min at 60°C.
-All samples were run in triplicate in 10 $\mu$L volume consisting of 1 x TaqPath ProAmp Multiplex Master Mix, 0.9 $\mu$M of each primer (forward: AAATGTTTAAACTAGAGCCGAATAGC and reverse: TCGTGGAGTCAAAGTGGGGTAGA), 0.2 $\mu$M of probe (6FAM-CACTCGAGGCCACGAAGTACAATT-(MGB)NFQ), and 2 $\mu$L of DNA template or water for the non-template control.
+All samples were run in 10 $\mu$L volume consisting of 1 x TaqPath ProAmp Multiplex Master Mix, 0.9 $\mu$M of each primer (forward: AAATGTTTAAACTAGAGCCGAATAGC and reverse: TCGTGGAGTCAAAGTGGGGTAGA), 0.2 $\mu$M of probe (6FAM-CACTCGAGGCCACGAAGTACAATT-(MGB)NFQ), and 2 $\mu$L of DNA template or water for the non-template control.
 All reactions included 100 copies of internal positive control (IPC) to detect PCR inhibition.

### 2025-02-12 Manuscript_plain.tex
Updates from Overleaf

--- 2025-02-11 Main_text.tex
@@
-\documentclass[9pt,twocolumn,twoside]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-\usepackage{soul} % Delete this line before submission XXX
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} %= Template for a one-column mathematics article
-% {pnasinvited} %= Template for a PNAS invited submission
+\documentclass{article}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage{soul}
+\bibliographystyle{plain}
+
+\title{Quantitative, Multispecies Monitoring at a Continental Scale}
+
+\author{Gledis Guri$^1$\textbf{*} \and
+Owen Liu$^2$ \and
+Ryan P. Kelly$^1$ \and
+Megan R. Shaffer$^1$\and
+Kim Parsons$^2$ \and
+Ana Ram\'on-Laca$^{1,2,3}$ \and
+Krista M. Nichols$^2$ \and
+Pedro F. P. Brandão-Dias$^1$ \and
+Abigail Wells$^2$ \and
+Andrew Olaf Shelton$^2$ \and}
+
+\date{\today}
 
@@
 
-\title{Quantitative, Multispecies Monitoring at a Continental Scale}
+\maketitle
 
-\author[a,1]{Gledis Guri}
-\author[b]{Owen Liu}
-\author[a]{Ryan P. Kelly}
-\author[a]{Megan R. Shaffer}
-\author[b]{Kim Parsons}
-\author[a,b,c]{Ana Ram\'on-Laca}
-\author[b]{Krista M. Nichols}
-\author[a]{Pedro F. P. Brandão-Dias}
-\author[b]{Abigail Wells}
-\author[b]{Andrew Olaf Shelton}
+\section*{}
 
-\affil[a]{School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA}
-\affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}
-\affil[c]{Museo Nacional de Ciencias Naturales, Madrid, Spain}
+\begin{center}
+\begin{tabular}{ll}
+1 & School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA \\
+2 & Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA \\
+3 & Museo Nacional de Ciencias Naturales, Madrid, Spain \\
+\hline
+\textbf{*} & \textbf{Gledis Guri}, lead and corresponding author \textbf{gguri@uw.edu}
+\end{tabular}
+\end{center}
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{Guri}
+\section*{Significance statement}
+Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales.
+We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States.
+The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to be broadly useful in natural-resources management and in the life sciences more generally.
 
-% Please add a significance statement to explain the relevance of your work
-\significancestatement{Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales.
-We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States.
-The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to be broadly useful in natural-resources management and in the life sciences more generally.}
+\section*{Author Contributions}
+AOS, KMN, RPK, MS, and KP designed the research;\\
+MRS, AR-L and AW did the laboratory work; \\
+PB, RPK, and GG did the bioinformatics;\\
+GG, OL, RPK, and eAOS analyzed the data;\\
+GG, OL, RPK, and AOS interpreted the results;\\
+GG, OL, RPK, and AOS wrote the manuscript
 
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{
-AOS, KMN, RPK, MS, and KP designed the research;
-MRS, AR-L and AW did the laboratory work; 
-PB, RPK, and GG did the bioinformatics;
-GG, OL, RPK, and AOS analyzed the data;
-GG, OL, RPK, and AOS interpreted the results;
-GG, OL, RPK, and AOS wrote the manuscript
-\hl{[Please check that this is correct and add/edit as needed XXX]}
-}
-\authordeclaration{The authors declare no competing interest.}
+[Please check that this is correct and add/edit as needed XXX]
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: gguri@uw.edu}
+\section*{Author Declaration}
+The authors declare no competing interest.
 
-% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{eDNA $|$ Bayesian $|$ bio-monitoring $|$ quantification $|$ fish}
+\section*{Keywords}
+eDNA $|$ Bayesian $|$ bio-monitoring $|$ quantification $|$ fish
 
-\begin{abstract}
+\section*{Abstract}
 Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural-resources management at a broad geographic scale.
@@
 Notably, our analyses identified biodiversity hotspots that align with previously documented regions of ecological significance, such as the Columbia River plume and the Heceta Bank.
-\end{abstract}
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-\firstpage[4]{4}
+\section{Introduction}
 Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}.
@@
 
-\section*{Results}
-
+\section{Results}
 We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
@@
 
-\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
+\begin{figure*}[tbhp]
 \centering
@@
 
-\begin{figure}[tbhp]  % Use figure* for a column-wide figure
+\begin{figure}[tbhp] 
 \centering
-\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}  
 \caption{Correlation plot describing species co-abundance and co-occurance.
@@
 
-\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
+\begin{figure*}[tbhp]  
 \centering
-\includegraphics[width=14cm]{plots/3_Figure_3.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=13cm]{plots/3_Figure_3.jpg}  
 \caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m).
@@
 
-%\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
+%\begin{figure*}[tbhp]  
 %\centering
@@
 
-\section*{Discussion}
+\section{Discussion}
 We report species-specific DNA concentrations smoothed over ten degrees of latitude and across the continental shelf to 500m.
@@
 
-\matmethods{We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
+\section{Methods}
+We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
 
@@
 A subset of the samples collected in \cite{shelton2022}, was used for this study.
+
+\begin{figure*}[tbhp] 
+\centering
+\includegraphics[width=16cm]{plots/ppt_figures/DAG.pdf} 
+\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
+Model compartment indicated in the same color (2-5) are run jointly.
+Latent parameters, state variables, data and subscripts are indicated in gray.
+See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
+\label{fig:DAG}
+\end{figure*}
 
@@
 Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
-
-\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
-\centering
-\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
-\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
-Model compartment indicated in the same color (2-5) are run jointly.
-Latent parameters, state variables, data and subscripts are indicated in gray.
-See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
-\label{fig:DAG}
-\end{figure*}
 
@@
 
-}
-
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
-This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
-The authors gratefully acknowledge the survey expertise and support of NOAA's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
-Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.}
-
-\showacknow{} % Display the acknowledgments section
-
-\bibsplit[9]
-%Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
-% Bibliography
-\bibliography{pnas-sample}
+\bibliography{pnas-sample.bib}
 

### 2025-02-12 Appendix_plain.tex
Updates from Overleaf

--- 2025-02-11 SI_Appendix.tex
@@
-\documentclass[9pt,twoside,lineno]{pnas-SI}
-% Use the lineno option to display guide line numbers if required.
+\documentclass{article}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage{soul}
+\usepackage{booktabs}
+\bibliographystyle{plain}
 
-\templatetype{pnassupportinginfo}
+\title{Appendix: Quantitative, Multispecies Monitoring at a Continental Scale}
 
-\title{Quantitative, Multispecies Monitoring at a Continental Scale}
-\author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Ana Ram\'on-Laca,\\
-Krista M. Nichols, Pedro F. P. Brandão-Dias, Abigail Wells and Andrew Olaf Shelton}
-\correspondingauthor{Gledis Guri.\\E-mail: gguri@uw.edu}
+\author{Gledis Guri$^1$\textbf{*} \and
+Owen Liu$^2$ \and
+Ryan P. Kelly$^1$ \and
+Megan R. Shaffer$^1$\and
+Kim Parsons$^2$ \and
+Ana Ram\'on-Laca$^{1,2,3}$ \and
+Krista M. Nichols$^2$ \and
+Pedro F. P. Brandão-Dias$^1$ \and
+Abigail Wells$^2$ \and
+Andrew Olaf Shelton$^2$ \and}
+
+\date{\today}
 
@@
 
-%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
-\SItext
+\section*{}
+
+\begin{center}
+\begin{tabular}{ll}
+1 & School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA \\
+2 & Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA \\
+3 & Museo Nacional de Ciencias Naturales, Madrid, Spain \\
+\hline
+\textbf{*} & \textbf{Gledis Guri}, lead and corresponding author \textbf{gguri@uw.edu}
+\end{tabular}
+\end{center}
 
@@
 \subsection*{Metabarcoding}
-In total, 568 samples, including 554 environmental samples, 7 PCR blanks and 7 positive controls (holding only kangaroo DNA) were amplified using MiFish-U universal primers \cite{miya2015} with Illumina tails (forward TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGGCCGGTAAAACTCGTGCCAGC; reverse GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCATAGTGGGGTATCTAATCCCAGTTTG) by using a two-step PCR protocol.
+In total, 568 samples, including 554 environmental samples, 7 PCR blanks and 7 positive controls (holding only kangaroo DNA) were amplified using MiFish-U universal primers \cite{miya2015} with Illumina tails (forward TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGGCCGGTAAAACTCGTGCCAGC; reverse GTCT- CGTGGGCTCGGAGATGTGTATAAGAGACAGCATAGTGGGGTATCTAATCCCAGTTTG) by using a two-step PCR protocol.
 The DNA was amplified in the first PCR reaction (PCR1) in a 20 µL reaction consisting of: 10 µL of Phusion Master Mix (2X), 0.4 µL of the forward primer (10 µM); 0.4 µL of the reverse primer (10 µM), 0.6 µL of 100\% DMSO, 0.5 µL of rAlbumin (20 µg/µL), 4.4 µL of nuclease-free water and 2 µL of DNA template.
@@
 \subsection*{Bayesian model}
-OLE
+OLE ...
 
@@
 \centering
-\includegraphics[width=0.98\textwidth]{plots/5_Supplementary_Figure_1.jpg}
+\includegraphics[width=0.90\textwidth]{plots/5_Supplementary_Figure_1.jpg}
 \caption{Continuation of Fig.
@@
 
+\clearpage
 \begin{table}\centering
@@
 \end{table}
-
+\clearpage
 \bibliography{pnas-sample}
-
 \end{document}

### 2025-02-14 Appendix_plain.tex
bib
Add aos edits

@@
 \usepackage{booktabs}
+\usepackage{amsmath}
 \bibliographystyle{plain}
@@
 
-\subsection*{Bayesian model}
-OLE ...
+\subsection*{Statistical model}
+
+\subsubsection*{Overview}
+
+We developed a joint model for modeling DNA concentration in the coastal ocean using single-species qPCR data Pacific hake and metabarcoding data from the 12S MiFish Universal primer.
+This work builds off previous single-species eDNA models constructed for Pacific hake (Shelton et al.
+2022) and models for joining qPCR and metabarcoding data (Guri et al.
+2024).
+Here qPCR observations provide information about DNA concentration for one species (Pacific hake) and the metabarcoding data provide information about species composition within a sample (i.e.
+proportions that sum to one).
+In combination, these two data sources allow for estimation of DNA concentrations for all species observed by metabarcoding.
+
+\subsubsection*{qPCR model}
+
+We developed a state-space framework for modeling hake DNA concentration in the coastal ocean building and modifying the work of Shelton et al.
+(2022).
+State-space models separate the true biological process from the methods used to observe the process.
+Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA $copies$ $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
+Let $\alpha_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
+We write these as additive on the log-scale,
+
+\begin{align}
+  D_H(x,y,d) &= \psi_0 + \epsilon(x,y,d) \\
+  \epsilon(x,y,d) &\sim N(0,\sigma_D)
+\end{align}
+
+This model describes perhaps the simplest possible random effect model and notably does not leverage any information about the spatial arrangement of samples (i.e.
+the random effect does not include a spatial correlation component).
+This is intentional.
+The focus of this model is on the other species DNA concentration (not Pacific hake) and therefore we allow the estimated hake concentrations to follow the qPCR observations as much as possible.
+
+Unfortunately, qPCR does not directly measure DNA concentration.
+Instead qPCR measures the PCR cycle at which a 2$\mu L$ aliquot of a extacted DNA was detected to fluoresce and compares that observation against the flourescence pattern of samples of known DNA concentration ("the standard curve").
+This provides an estimate of DNA concentration for each unknown sample.
+Each water sample is analyzed using at least 3 independent qPCR replicates to account for laboratory and machine variability.
+In addition to qPCR variability among replicates, we also have to account for modifications that affect each water sample and occurred during water sampling and processing.
+First, we have three offsets that modify the true DNA concentration to affect what we observed in the qPCR.
+For water sample $i$: 1) $V_i$ is the proportion of 2.5 L filtered from Niskin $i$ (occasionally some seawater was spilled or not filtered; $V_i =1$ for the vast majority of samples); 2) $I_i$ is the known dilution used to on sample $i$ to eliminate PCR inhibition.
+PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see Shelton et al.
+2022 Supp S1 for additional description).
+
+Finally, we add a random effect for the individual bottles sampled at each year-station-depth combination ($\delta_i(x,y,d)$).
+This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_i(x,y,d) \sim N(0,\tau)$, with $\tau$ indicating a depth-specific standard deviation among bottles.
+Because we have only two replicate bottles at each location, we impose a sum to zero constraint on the bottles collected from a single location (i.e.
+for a given location, $\sum_i \delta_i(x,y,d) = 0$).
+Then, the log-concentration of hake DNA in a sample analyzed by qPCR, $E_{Hi}$ is,
+
+\begin{align}
+  E_{Hi} &= D_H({x,y,d}) +  \log{V_i} + \log{I_i} + \mathbf{I}\zeta + \delta_i(x,y,d)
+\end{align}
+
+We connect the estimated concentrations to observations from qPCR using two likelihoods connected to the concentration $E$.
+First we determine whether amplification was detected in each qPCR replicate $r$ run on each plate $j$, $Z_{Hijr}$ and if amplification was observed we model the PCR cycle at which amplification was detected $Y_{Hijr}$; $Y_{Hijr}$ is observed as a continuous, positive value.
+DNA concentrations have a log-linear relationships with $Y_{Hijr}$; smaller values of $Y_{Hijr}$ are associated with higher DNA concentrations.
+
+Because we are modeling the discrete number of DNA molecules that are present in the assayed sample, we know that the Poisson distribution provides an appropriate observation distribution for the number of molecules present in a given qPCR reaction.
+Assuming a Poisson distribution, conditional on a true mean number of DNA copies in the sample $X$, the probability of having exactly zero DNA copies in a qPCR reaction is $e^{-X}$ and the probability of having non-zero DNA copies is the complement, $1-e^{-X}$.
+If there are exactly zero molecules in a qPCR reaction, amplification and therefore detection of amplification will not occur.
+However, there are other factors (e.g.
+PCR inhibition) that may reduce amplification further and therefore we expect the probability of amplification to be at most $1-e^{-X}$.
+As a result, we estimate an additional term, $\phi$ representing the fractional reduction in amplification efficiency due to other factors on qPCR plate $j$ ($0<\phi<1$).
+Then we have a pair of observation models for hake DNA, both of which are a function of DNA concentration, $E_{Hi}$
+
+\begin{align}
+  Z_{Hijr} &\sim Bernoulli\left(1 - \exp(-2 e^{E_{Hij}}\phi) \right) \\ 
+  Y_{Hijr} &\sim Normal \left(\beta_{0j}+\beta_{1j}{E_{Hij}},\sigma_Y(E_{Hij}) \right) \qquad  if \: Z_{Hijr} = 1 
+\end{align}
+
+The 2 is present in the first line of the equation because we use 2 $\mu L$ of sample in each qPCR reaction and so the expected DNA copies in reaction is $2e^{E_{ij}}$.
+We allow $\sigma_C(E_{ij})$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_C({E_{ij}}) = exp(\gamma_{0}+\gamma_{1}E_{ij})$ (see also Fig.
+XX).
+
+By itself, the above model is unidentifiable because field samples do not provide information about the parameters that define relationship between the number of DNA copies and PCR cycle ($\beta_{0j}$, $\beta_{1j}$, and $\phi$).
+As is standard with qPCR analyses, we include standards of known concentration to estimate these parameters.
+Each qPCR plate has replicate samples with a known number of DNA copies.
+These standards span six orders of magnitude (1 to 100,000 copies $\mu L^{-1}$, each of 2$\mu L$) and determine the relationship between copy number and PCR cycle of detection.
+Let $K_{jp}$ be the known copy number for sample $j$ in PCR plate $p$, then,
+
+\begin{align}
+  Z_{jp} &\sim Bernoulli \left(1-exp(-2 K_{jp} \phi) \right)\\
+  Y_{jp} &\sim Normal \left(\beta_{0j}+\beta_{1j}\log{K_{jp}},\sigma_Y(K_{jp})\right) \quad  if \: Z_{jp} =1
+\end{align}
+
+Note that there are different intercept $\beta_{0p}$ and slope $\beta_{1p}$ parameters for each PCR plate to allow for among-plate variation in amplification.
+We model each calibration paratemeter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
+$\beta_{0p} \sim N(\mu_{\beta_{0}},\sigma_{\beta_{0}})$).
+
+\subsubsection*{Metabarcoding model}
+
+In addition to the qPCR observations, we used the MiFish primer and high throughput sequencing to detect and enumerate the sequences of other fish species in the same samples that were examined for hake DNA.
+Again, we are interested in determining the DNA concentration of each focal species in each sample.
+Unfortunately, metabarcoding data is compositional data (see *Glor et al.
+*, *Shelton et al.
+2023*), and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
+Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
+Here, we follow and extend the metabarcoding model developed by Shelton et al.
+(2023) and show how it links with the qPCR model above.
+
+As above, we let $D_s(x,y,d)$ be the log-scale eDNA concentration but now for species $s$.
+We let $E_{si}(x,y,d)$ be the log-DNA concentration for each species in each bottle.
+Importantly one of the $s$ species is Pacific hake and so connects to the qPCR model above.
+Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; *Aitchison XXXX*).
+We define a reference species in each sample, $R$, with the species used as a reference varying among sample.
+Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see Shelton et al.
+2023).
+We can then calculate $\nu_{si}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see *Mock Communities* below),
+
+\begin{align} 
+  \nu_{si} &= \left( E_{si}(x,y,d) - E_{Ri}(x,y,d)\right) + N_{PCR} \left(\alpha_s - \alpha_R \right) 
+\end{align}
+
+To maintain identifiability, we assign one $\alpha_s = 0$.
+We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\mu_{si}$ using the softmax transformation and to the observed reads via a Dirichlet-Multinomial likelihood,
+
+\begin{align} 
+  \pi_{si} &= \dfrac{e^{\nu_{si}}}{\sum_{s=1}^S e^{\nu_{si}}} \\
+  \boldsymbol{R}_{i} & \sim Multinomial\left(\boldsymbol{\pi}_{i},T_{i}\right) 
+\end{align}
+
+where $T_{i}$ is the total number of reads observed across all species for that bottle.
+
+As with the qPCR analyses, in the absence of additional information estimates of $\alpha$ and $E$ cannot be statistically identified.
+We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but for the moe
+
+\subsubsection*{Smoothing predictions}
+
+\textbf{OWEN.  COULD YOU HELP HERE?}
 
@@
 \clearpage
-\bibliography{pnas-sample}
+\bibliography{Bib.bib}
 \end{document}

### 2025-02-14 Manuscript_plain_AOS.tex
Add aos edits

--- 2025-02-12 Manuscript_plain.tex
@@
 \usepackage{soul}
+\usepackage{setspace} 
+\doublespacing 
+\usepackage{lineno} 
+\linenumbers
 \bibliographystyle{plain}
@@
 GG, OL, RPK, and AOS interpreted the results;\\
-GG, OL, RPK, and AOS wrote the manuscript
+All authors provided edits to the manuscript
 
@@
 Here, we combine qPCR and metabarcoding data derived from the same samples to map quantities of many fish species along the US West Coast in three dimensions, demonstrating a technique of practical importance for both management and ecology.
-We derive spatially explicit estimates of eDNA abundance for 12 common species of commercial and management importance and point the way to quantitative surveys of wild species using molecules alone.
+We derive spatially explicit estimates of eDNA abundance for 12 common species of ecological and management importance and point the way to quantitative surveys of wild species using molecules alone.
 We find that species distribution maps derived from eDNA effectively capture species niches and spatial distributions.
@@
 Indeed, eDNA methods may provide some notable improvements because single- and multi-species eDNA approaches avoid tradeoffs involved with physically capturing many species simultaneously using a single net.
-Molecular surveys capture many species with the same "gear" using methods that are easily replicated in space and time.
+Molecular surveys capture many species with the same ``gear" using methods that are easily replicated in space and time.
 As U.S.
@@
 Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24°C for sardine vs.
-10–14°C for anchovy); Pacific sardine also typically inhabits slightly greater depths (25; Fig.
+10–14°C for anchovy).
+We found more Pacific sardine DNA in sub-surface waters as well (Fig.
 \ref{fig:fig1}).
@@
 S2}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
-Furthermore, eDNA abundance patterns aligned with stock assessments in indicating a higher abundance of northern anchovy; however, the magnitude of difference varied, with eDNA suggesting a 2.5-fold difference, while stock assessments indicated a 35-fold difference in abundance.
-\hl{[CITE stock assessment here; OLE XXX]}
+Furthermore, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
+However, due to differences in stock definitions and survey areas (the anchovy stock assessment includes southern California waters (32.5°–36° N) not covered by the eDNA survey) these magnitudes are not directly comparable.
 
 \subsection*{Mid-water}
-Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} \hl{[ADD SCUBA REF; OLE XXX]}.
-Moreover, because it is a common bycatch species in the hake (\textit{Merluccius productus}) fishery, limits on bycatch of widow rockfish can also limit the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
-Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat.
+Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024, laidig2007} .
+Moreover, because it is a common bycatch species in the Pacific hake (\textit{Merluccius productus}) fishery, concerns about bycatch of widow rockfish can also influence the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
+Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat well.
 We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
@@
 Also consistent with known habitats are commercially caught flatfish such as Dover sole (\textit{Microstomus pacificus}), which lives at 200m to 1200m along the Pacific coast \cite{drazen2007,brodziak2000,drazen2012}; individuals move progressively deeper as they grow \cite{vetter1994,hunter1990}.
-The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S2}\textit{A}) \cite{ono2016}.
@@
 For well-studied taxa, existing information helps contextualize eDNA signals, providing insights into habitat preferences, seasonal patterns, or temperature tolerances.
-For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa, revealing unexpected habitats.
-Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \hl{[NWFSC trawl survey ref; OLE XXX]}.
+For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa and revealing unexpected habitats.
+Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \cite{keller2017}.
 This species has been of special interest, as its targeted commercial harvest resumed in the late 2010s after a decade of closure (harvest was less than 500 mt annually between 2003 and 2013 but has exceeded 10,000 mt annually since 2018) and no existing survey adequately captures its patterns of abundance \cite{adams2019a}.
@@
 
-The majority of eDNA shed into the environment dissipates from surface waters within hours \hl{(Brasseale et al., in review)[add this reference; RYAN or OLE XXX]}.
-Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most.
-Given that our sampling stations are spaced more than 30 km apart, the spatial patterns we report are likely to reflect genuine differences in fish abundance rather than artifacts of eDNA transport dynamics \cite{guri2023}.
-Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; \textit{Sebastes entomelas} occurs almost exclusively at the 50m depth bin; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig.
+The majority of eDNA shed into the environment dissipates from surface waters within hours \cite{andruszkiewiczallan2021}.
+Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most \cite{baetscher2024}.
+Given that our sampling area spans more than 1,200 km north to south and tens of km east to west, the spatial patterns we report are likely to reflect genuine differences in fish communities rather than artifacts of eDNA transport dynamics \cite{guri2023}.
+Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; Widow rockfish occurs almost exclusively at the 50m depth; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig.
 \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station.
-Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+
+Large-scale sampling combined with statistical smoothing mitigates many challenges associated with eDNA transport and degradation -- which decouple the eDNA observations from their particular source organisms -- allowing reliable detection of broad patterns in species distributions.
+Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature or other covariates \cite{jo2019b}.
+
+%Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
 S7}) and spatially smoothed maps (Fig.
@@
 
-Large-scale sampling combined with statistical smoothing mitigates many challenges associated with eDNA transport and degradation -- which decouple the eDNA observations from their particular source organisms -- allowing reliable detection of broad patterns in species distributions.
-Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature \cite{jo2019b}.
-
 %(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S6}) posterior predictive
@@
 \subsection*{Integrating DNA with traditional surveys}
-Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes.
-Here, we follow \cite{shelton2022} and \hl{Liu et al.
-(submitted)[add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
-Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision)[add this reference; RYAN or OLE XXX]}.
+Effective management of natural resources at large scales requires that we maximize the information content of existing sampling regimes.
+Here, we follow \cite{shelton2022}, leveraging the same water samples to provide information on an additional 12 species of ecosystem and management importance.
+Elsewhere, a subset of these samples has yielded information on the distribution of marine mammals \cite{valdiviacarrillo2025}.
 The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa.
@@
 We have shown an example of the substantial value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
-Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional trawl-based surveys.
+Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional acoustic or trawl-based surveys.
+Furthermore, eDNA survey provide data in units of eDNA concentration, not more familiar units of mass or abundance.
+Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
+We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
 
-Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
+%Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
 In reality, fish can move in between sampling location within the study area, which may bias abundance estimates  \cite{rota2009}.
@@
 
-Perhaps most notably, management decisions happen in units of mass or abundance, not in units of eDNA concentration.
-Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that are likely to vary among species.
-We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
-
 \section{Methods}
@@
 \subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
-Samples were collected in 10L Niskin bottles deployed on a CTD rosette at five depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
-Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey.
-2.5L water samples were collected from the Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
+We used samples collected in 10L Niskin bottles deployed on a CTD rosette at six depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
+Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey \cite{deblois2020}.
+2.5L water samples were collected from each Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
 A subset of the samples collected in \cite{shelton2022}, was used for this study.
@@
 \centering
-\includegraphics[width=16cm]{plots/ppt_figures/DAG.pdf} 
-\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
-Model compartment indicated in the same color (2-5) are run jointly.
-Latent parameters, state variables, data and subscripts are indicated in gray.
+\includegraphics[width=16cm]{plots/DAG-Ole.pdf} 
+\caption{Schematic for the statistical models to produce species distributions.
+Three sets of observations (orange) were estimated simultaneously to provide estimated eDNA concentrations twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
+Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
+Notation for parameters, subscripts, and observations are indicated in gray.
 See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
-\label{fig:DAG}
+\label{fig:DAG2}
 \end{figure*}
@@
 Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix} for details on metabarcoding methods).
+Importantly, we only analyzed one water sample collected from each sampling station using metabarcoding.
 
@@
 Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S4}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI_Appendix.pdf}{SI Appendix}).
+S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix}).
 Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
@@
 We selected 12 species (additional to Pacific hake) of commercial and ecological importance that co-occurred in the environmental samples and in the mock community for downstream analysis: Pacific herring (\textit{Clupea pallasii}), northern anchovy (\textit{Engraulis mordax}), northern smoothtongue (\textit{Leuroglossus stilbius}), Dover sole (\textit{Microstomus pacificus}), Pacific sardine (\textit{Sardinops sagax}), Pacific chub mackerel (\textit{Scomber japonicus}), widow rockfish (\textit{Sebastes entomelas}), northern lampfish (\textit{Stenobrachius leucopsarus}), California barracudina (\textit{Tactostoma macropus}), plainfin midshipman (\textit{Tarletonbeania crenularis}), eulachon (\textit{Thaleichthys pacificus}), Pacific jack mackerel (\textit{Trachurus symmetricus}).
-Because we used Pacific hake (\textit{Merluccius productus}) as a comparative reference species throughout (see below), we did not include the spatial patterns of that species in our analysis.
-After amplification and sequencing using the protocols outlined above, we used the mock communities to estimate species-specific amplification efficiencies and correct for amplification bias following \cite{shelton2022}.
+As Pacific hake (\textit{Merluccius productus}) has been examined elsewhere using qPCR \cite{shelton2022}, we did not include the spatial patterns of that species in our analysis.
+After amplification and sequencing using the protocols outlined above, we used the mock communities to estimate species-specific amplification efficiencies and correct for amplification bias following \cite{shelton2023}.
 
@@
 The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig.
-\ref{fig:DAG}).
+\ref{fig:DAG2}).
 We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
 This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig.
-\ref{fig:DAG}).
-This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification) as well as the effect of covariates (e.g., the depth of the ocean bottom) on hake DNA concentration (Fig.
-\ref{fig:DAG}).
-This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions coming from DNA metabarcoding (Fig.
-\ref{fig:DAG}).
+\ref{fig:DAG2}).
+This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (Fig.
+\ref{fig:DAG2}).
+This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (Fig.
+\ref{fig:DAG2}).
 
@@
 Furthermore, metabarcoding does not yield a consistent, species-specific detection probability (i.e., sensitivity).
-Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification, as well as on the total sequencing depth for that sample \cite{kelly2019}.
+Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification \cite{shelton2023}, as well as on the total sequencing depth for that sample \cite{kelly2019}.
 Higher sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
 
-We follow \cite{shelton2022} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR.
+We follow \cite{shelton2023} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR.
 We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species).
@@
 To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurance, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
 S5}).
 Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), and their omission appears not to have induced substantial bias in our estimates.
+S4}), and deserve further attention in future work.
 
@@
 The above statistical model (Fig.
-\ref{fig:DAG}) provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA (Fig.
-\ref{fig:DAG}, green box).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
+\ref{fig:DAG2}) provides predictions of species-specific DNA concentrations (D) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (Fig.
+\ref{fig:DAG2}, green box).
 
-All statistical analysis, figure preparation, and data curation were performed in R \cite{rcoreteam2024}.
+%Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
 
-\bibliography{pnas-sample.bib}
+Analyses, figure preparation, and data curation were performed in R \cite{rcoreteam2024} and the joint statistical model for qPCR and metabarcoding was coded in stan \cite{standevelopmentteam2023}.
+
+\bibliography{Bib.bib}
 

### 2025-02-14 Manuscript_plain_AOS.tex
update main text

@@
 1 & School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA \\
-2 & Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA \\
+2 & Conservation Biology Division, Northwest Fisheries Science Center,  NOAA Fisheries, \\
+ & \qquad 2715 Montlake Blvd E. Seattle, Washington, USA \\
 3 & Museo Nacional de Ciencias Naturales, Madrid, Spain \\
@@
 GG, OL, RPK, and AOS interpreted the results;\\
-All authors provided edits to the manuscript
-
-[Please check that this is correct and add/edit as needed XXX]
+All authors provided edits to the manuscript \\
+Please check that this is correct and add/edit as needed XXX
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}.
@@
 
-\section{Results}
+\section*{Results}
 We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
@@
 
-The eDNA patterns for northern lampfish (Stenobrachius leucopsarus) -- an abundant myctophid that migrates from 400-700m up to 20-200m at night \cite{moku2000,suntsov2008} -- are again consistent with the ecology of the species, having notably lower concentrations in surface samples.
+The eDNA patterns for northern lampfish (\textit{Stenobrachius leucopsarus}) -- an abundant myctophid that migrates from 400-700m up to 20-200m at night \cite{moku2000,suntsov2008} -- are again consistent with the ecology of the species, having notably lower concentrations in surface samples.
 Owing to the diel vertical migration of the northern lampfish, the signal of eDNA for this species is observed across depths.
@@
 \centering
-\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}  
+\includegraphics[width=10cm]{plots/2_Figure_2.jpg}  
 \caption{Correlation plot describing species co-abundance and co-occurance.
@@
 
-\section{Discussion}
+\section*{Discussion}
 We report species-specific DNA concentrations smoothed over ten degrees of latitude and across the continental shelf to 500m.
@@
 
-\section{Methods}
+\section*{Methods}
 We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
@@
 \caption{Schematic for the statistical models to produce species distributions.
-Three sets of observations (orange) were estimated simultaneously to provide estimated eDNA concentrations twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
+Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
 Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
 Notation for parameters, subscripts, and observations are indicated in gray.
-See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
+See \href{SI_Appendix.pdf}{SI Appendix} for full model description.}
 \label{fig:DAG2}
@@
 Our model generates estimates of DNA concentration (copies/L) for each species.
-Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species; we refer to hake as our “reference species”.
+Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species.
 Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species.
@@
 The above statistical model (Fig.
-\ref{fig:DAG2}) provides predictions of species-specific DNA concentrations (D) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+\ref{fig:DAG2}) provides predictions of species-specific DNA concentrations (\textit{D}) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
 To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (Fig.

### 2025-02-20 Priors table.tex

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{soul}
\bibliographystyle{plain}
\usepackage{amsmath}

\begin{document}

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
         & \textbf{Description} & \textbf{Prior} \\
&Data & \\
\hline
$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
$Z$ & qPCR amplification (yes=1; no=0) & - \\
$Y$ & qPCR cycle threshold (Ct) & - \\
$R$ & metabarcoding number of reads & - \\
$T$ & total number of reads for all S species & - \\
$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
$V$ & volume filtered and aliquote dilution offset & -\\
$I$ & aliquote dilution factor & -\\

&&\\
&State processes&\\
\hline
$E$ & unknown DNA concentration in copies/$\mu$L & - \\
$D$ & unknown DNA concentration in copies/L & - \\
$F$ & smoothed DNA concentration in copies/L & - \\
&&\\

&Parameters&\\
\hline
$\psi$& mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
$\epsilon_{(x,y,d)}$& depth specific random effect & $\mathcal{N}(0,\tau)$\\
$\tau$& standard deviation of depth specific random effect & $\mathcal{N}(0,0.2)$\\
$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
$\sigma_D$& standard deviation of biological replicate random effect  & $\mathcal{N}(0,3)$\\
$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
&and eDNA concentration (K and E) & \\
$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
&eDNA concentration (K and E) & \\
$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
&of Ct values and eDNA concentration (K and E) & \\
$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
&of Ct values and eDNA concentration (K and E) & \\
$\alpha$& species amplification efficiency & fixed\\
$\zeta$& ethanol wash effect & fixed\\
&&\\
&Subscripts&\\
\hline
$s$& species ($H$ = Hake) & -\\
$x$& northing & -\\
$y$& easting & -\\
$d$& depth & -\\
$i$& biological replicate & -\\
$r$& technical replicate & -\\
$p$& qPCR plate & -\\
$m$& mock community sample & -\\
$j$& qPCR standard aliquote sample & -\\

    \end{tabular}
    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
    \label{tab:filter_error}
\end{table}

\end{document}

mean_hake ~normal(2,2); //global mean hake concentration
log_D_sigma ~ normal(0,3); //variance on log_D_station_depth
log_D_station_depth
logit_phi_0
gamma_1 ~ std_normal();
gamma_0 ~ std_normal();
beta_std_curve_0
beta_std_curve_1
log_D_raw

### 2025-02-21 Priors table.tex
Updates from Overleaf

@@
 \hline
-$E$ & unknown DNA concentration in copies/$\mu$L & - \\
+$E_{s\neq H}$ & unknown DNA concentration in copies/$\mu$L & $\mathcal{N}(0,5)$\\
 $D$ & unknown DNA concentration in copies/L & - \\
@@
 \hline
-$\psi$& mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
-$\epsilon_{(x,y,d)}$& depth specific random effect & $\mathcal{N}(0,\tau)$\\
-$\tau$& standard deviation of depth specific random effect & $\mathcal{N}(0,0.2)$\\
-$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D$& standard deviation of biological replicate random effect  & $\mathcal{N}(0,3)$\\
+$\psi$& overall mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
+$\sigma_D= (0, +\infty)$& standard deviation of spatial and depth specific random effect & $\mathcal{N}(0,3)$\\
+
+$\delta$& biological replicate random effect & $\mathcal{N}(0,\tau)$\\
+$\tau = (0, +\infty)$& standard deviation of biological replicate random effect & $\mathcal{N}(0,0.2)$\\
+
 $\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
@@
 \end{document}
-
-mean_hake ~normal(2,2); //global mean hake concentration
-log_D_sigma ~ normal(0,3); //variance on log_D_station_depth
-log_D_station_depth
-logit_phi_0
-gamma_1 ~ std_normal();
-gamma_0 ~ std_normal();
-beta_std_curve_0
-beta_std_curve_1
-log_D_raw

### 2025-02-21 Appendix_plain_AOS.tex

--- 2025-02-14 Appendix_plain.tex
@@
 Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA $copies$ $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
-Let $\alpha_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
+Hereafter we refer to a combination of spatial location and depth as a ``station".
+Let $\psi_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
 We write these as additive on the log-scale,
@@
 This is intentional.
-The focus of this model is on the other species DNA concentration (not Pacific hake) and therefore we allow the estimated hake concentrations to follow the qPCR observations as much as possible.
+The focus of this model is on the other species DNA concentration (not Pacific hake) and therefore we allow the estimated hake concentrations at a to follow the qPCR observations as much as possible.
 
@@
 2022 Supp S1 for additional description).
+Here we fix $\zeta$ = XXX based on previous analyses (Shelton et al.
+2022).
 
 Finally, we add a random effect for the individual bottles sampled at each year-station-depth combination ($\delta_i(x,y,d)$).
-This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_i(x,y,d) \sim N(0,\tau)$, with $\tau$ indicating a depth-specific standard deviation among bottles.
+This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_i(x,y,d) \sim N(0,\sigma_D)$, with $\sigma_D$ indicating a depth-specific standard deviation among bottles.
 Because we have only two replicate bottles at each location, we impose a sum to zero constraint on the bottles collected from a single location (i.e.
@@
 The 2 is present in the first line of the equation because we use 2 $\mu L$ of sample in each qPCR reaction and so the expected DNA copies in reaction is $2e^{E_{ij}}$.
-We allow $\sigma_C(E_{ij})$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_C({E_{ij}}) = exp(\gamma_{0}+\gamma_{1}E_{ij})$ (see also Fig.
-XX).
+We allow $\sigma_C(E_{ij})$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_C({E_{ij}}) = exp(\gamma_{0}+\gamma_{1}E_{ij})$.
 
@@
 These standards span six orders of magnitude (1 to 100,000 copies $\mu L^{-1}$, each of 2$\mu L$) and determine the relationship between copy number and PCR cycle of detection.
-Let $K_{jp}$ be the known copy number for sample $j$ in PCR plate $p$, then,
+Let $K_{jp}$ be the known log-copy number for sample $j$ in PCR plate $p$, then,
 
 \begin{align}
-  Z_{jp} &\sim Bernoulli \left(1-exp(-2 K_{jp} \phi) \right)\\
-  Y_{jp} &\sim Normal \left(\beta_{0j}+\beta_{1j}\log{K_{jp}},\sigma_Y(K_{jp})\right) \quad  if \: Z_{jp} =1
+  Z_{jp} &\sim Bernoulli \left(1-exp(-2 e^{K_{jp}} \phi) \right)\\
+  Y_{jp} &\sim Normal \left(\beta_{0j}+\beta_{1j}{K_{jp}},\sigma_Y(K_{jp})\right) \quad  if \: Z_{jp} =1
 \end{align}
 
-Note that there are different intercept $\beta_{0p}$ and slope $\beta_{1p}$ parameters for each PCR plate to allow for among-plate variation in amplification.
-We model each calibration paratemeter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
+Note that there are different intercept ($\beta_{0p}$) and slope ($\beta_{1p}$) parameters for each PCR plate to allow for among-plate variation in amplification.
+We model each calibration parameter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
 $\beta_{0p} \sim N(\mu_{\beta_{0}},\sigma_{\beta_{0}})$).
@@
 Again, we are interested in determining the DNA concentration of each focal species in each sample.
-Unfortunately, metabarcoding data is compositional data (see *Glor et al.
-*, *Shelton et al.
-2023*), and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
+Unfortunately, metabarcoding data is compositional data \cite{glor2017, shelton2023}, and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
 Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
@@
 Importantly one of the $s$ species is Pacific hake and so connects to the qPCR model above.
-Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; *Aitchison XXXX*).
+Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; \textbf{Aitchison 1986}).
 We define a reference species in each sample, $R$, with the species used as a reference varying among sample.
-Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see Shelton et al.
-2023).
+Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see \cite{shelton2023}).
 We can then calculate $\nu_{si}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see *Mock Communities* below),
@@
 \begin{align} 
-  \nu_{si} &= \left( E_{si}(x,y,d) - E_{Ri}(x,y,d)\right) + N_{PCR} \left(\alpha_s - \alpha_R \right) 
+  \nu_{si} &= \left( E_{si}(x,y,d) - E_{Ri}(x,y,d)\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) 
 \end{align}
 
-To maintain identifiability, we assign one $\alpha_s = 0$.
-We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\mu_{si}$ using the softmax transformation and to the observed reads via a Dirichlet-Multinomial likelihood,
+We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\mu_{si}$ using the softmax transformation and to the observed reads via a Multinomial likelihood,
 
@@
 As with the qPCR analyses, in the absence of additional information estimates of $\alpha$ and $E$ cannot be statistically identified.
-We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but for the moe
+We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but instead of having unknown DNA concentrations, we have known inputs for each species $L_{sm}$ in mock community $m$.
+Then:
+
+\begin{align} 
+ \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) \\
+  \pi_{sm} &= \dfrac{e^{\nu_{sm}}}{\sum_{s=1}^S e^{\nu_{sm}}} \\
+  \boldsymbol{R}_{m} & \sim Multinomial\left(\boldsymbol{\pi}_{m},T_{m}\right) 
+\end{align}
+
+To maintain identifiability, we assign one $\alpha_s = 0$. (ADD. WHICH ONE IS ZERO?, WHICH IS THE REF SPECIES?). ADD PLOT of ALPHAS.
 
@@
 
+Ole's questions:
+Smoothed model is based on E (a single water bottle), not D.
+What was used in the making the smooth?  Posterior mean?
+
 \textbf{OWEN.  COULD YOU HELP HERE?}
 
-The above statistical model provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
+The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
 We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as:
-$$\ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)$$
+$$E_i(xyz) = \tau + \eta(s) + \varepsilon_{d}(s)$$
 where x,y,z index longitude, latitude, and depth, respectively, $\tau$ is scalar intercept, $\eta(s)$ is a spatial smooth shared among depths, and $\varepsilon_d(s)$ is a depth-specific spatial smooth.
@@
 \subsection*{Model convergance}
-OLE + GLED
+
+Gelman-Rubin statistics.  Predicted-observed, etc.
 
@@
 \clearpage
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{lll}
+         & \textbf{Description} & \textbf{Prior} \\
+&Data & \\
+\hline
+$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
+$Z$ & qPCR amplification (yes=1; no=0) & - \\
+$Y$ & qPCR cycle threshold (Ct) & - \\
+$R$ & metabarcoding number of reads & - \\
+$T$ & total number of reads for all S species & - \\
+$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
+$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
+$V$ & volume filtered and aliquote dilution offset & -\\
+$I$ & aliquote dilution factor & -\\
+
+&&\\
+&State processes&\\
+\hline
+$E$ & unknown DNA concentration in copies/$\mu$L & - \\
+$D$ & unknown DNA concentration in copies/L & - \\
+$F$ & smoothed DNA concentration in copies/L & - \\
+&&\\
+
+&Parameters&\\
+\hline
+$\psi$& mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\epsilon_{(x,y,d)}$& depth specific random effect & $\mathcal{N}(0,\tau)$\\
+$\tau$& standard deviation of depth specific random effect & $\mathcal{N}(0,0.2)$\\
+$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
+$\sigma_D$& standard deviation of biological replicate random effect  & $\mathcal{N}(0,3)$\\
+$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
+$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
+&and eDNA concentration (K and E) & \\
+$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
+&eDNA concentration (K and E) & \\
+$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\alpha$& species amplification efficiency & fixed\\
+$\zeta$& ethanol wash effect & fixed\\
+&&\\
+&Subscripts&\\
+\hline
+$s$& species ($H$ = Hake) & -\\
+$x$& northing & -\\
+$y$& easting & -\\
+$d$& depth & -\\
+$i$& biological replicate & -\\
+$r$& technical replicate & -\\
+$p$& qPCR plate & -\\
+$m$& mock community sample & -\\
+$j$& qPCR standard aliquote sample & -\\
+
+    \end{tabular}
+    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
+    \label{tab:filter_error}
+\end{table}
+\clearpage
+
 \bibliography{Bib.bib}

### 2025-02-21 Appendix_plain_AOS.tex
resolve conflict appendix

@@
 \clearpage
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{lll}
-         & \textbf{Description} & \textbf{Prior} \\
-&Data & \\
-\hline
-$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
-$Z$ & qPCR amplification (yes=1; no=0) & - \\
-$Y$ & qPCR cycle threshold (Ct) & - \\
-$R$ & metabarcoding number of reads & - \\
-$T$ & total number of reads for all S species & - \\
-$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
-$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
-$V$ & volume filtered and aliquote dilution offset & -\\
-$I$ & aliquote dilution factor & -\\
-
-&&\\
-&State processes&\\
-\hline
-$E$ & unknown DNA concentration in copies/$\mu$L & - \\
-$D$ & unknown DNA concentration in copies/L & - \\
-$F$ & smoothed DNA concentration in copies/L & - \\
-&&\\
-
-&Parameters&\\
-\hline
-$\psi$& mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
-$\epsilon_{(x,y,d)}$& depth specific random effect & $\mathcal{N}(0,\tau)$\\
-$\tau$& standard deviation of depth specific random effect & $\mathcal{N}(0,0.2)$\\
-$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D$& standard deviation of biological replicate random effect  & $\mathcal{N}(0,3)$\\
-$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
-$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
-&and eDNA concentration (K and E) & \\
-$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
-&eDNA concentration (K and E) & \\
-$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
-&of Ct values and eDNA concentration (K and E) & \\
-$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
-&of Ct values and eDNA concentration (K and E) & \\
-$\alpha$& species amplification efficiency & fixed\\
-$\zeta$& ethanol wash effect & fixed\\
-&&\\
-&Subscripts&\\
-\hline
-$s$& species ($H$ = Hake) & -\\
-$x$& northing & -\\
-$y$& easting & -\\
-$d$& depth & -\\
-$i$& biological replicate & -\\
-$r$& technical replicate & -\\
-$p$& qPCR plate & -\\
-$m$& mock community sample & -\\
-$j$& qPCR standard aliquote sample & -\\
-
-    \end{tabular}
-    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
-    \label{tab:filter_error}
-\end{table}
-\clearpage
-
 \bibliography{Bib.bib}

### 2025-02-24 DAG-Ole.tex
push ole dag

\documentclass[landscape]{article}
\usepackage[paperwidth=250mm, paperheight=200mm, margin=0cm]{geometry}
\usepackage{tikz}
\usepackage{amsmath, amsthm, amssymb, mathtools, thmtools}
\usepackage{xcolor}
\definecolor{myBlue}{RGB}{173, 216, 230}
\definecolor{myGrey}{RGB}{200, 200, 200}

\usetikzlibrary{positioning, shapes.geometric, arrows}

% Define styles
\tikzstyle{node} = [rectangle, rounded corners, minimum width=1.5cm, minimum height=1cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=orange!35] at (-4,-2) {
    $\begin{array}{c}
       % \mu_{ik} = \beta_{0p} + \beta_{1p}  K_{i} \\
        \sigma_{Yj} = \exp(\gamma_0 + \gamma_1  K_{j}) \\
        Z_{jpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{j}} \phi) \right) \\
        Y_{jpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{j} , \sigma_{Yj}) \quad \textnormal{if } Z_{jpr} = 1 \\
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standards };
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_st.south east) {};

\node (qPCR_en) [node, below=of qPCR_st, fill=orange!35]  at (0,-3) {
    $\begin{array}{c}
       % \mu_{ip} = \beta_{0p} + \beta_{1p} E_{Hi} \\
        \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})\\
         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hi}} \phi) \right) \\
         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hi} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1
    \end{array}$
};
\node [below=of qPCR_en, yshift=+1.5cm, font=\large\bfseries] {Hake qPCR Likelihood};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_en.south east) {};

\node (state1) [node, above=of qPCR_en, fill=myBlue] at (2,0){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        D_{H}(x,y,d) = \psi  + \epsilon({x,y,d}) \\
        \epsilon({x,y,d})\sim \mathrm{N}(0,\tau)
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=0cm, font=\large\bfseries] {Hake DNA (log conc.)};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2H) [node, above=of qPCR_en, fill=myBlue] at (2.5,-2.5){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
       E_{Hi}(x,y,d) = D_H({x,y,d}) +  \log{V_i} + \\
       		\qquad \qquad \qquad \qquad \log{I_i} + \mathbf{I}\zeta + \delta_i(x,y,d) \\
		\delta_i(x,y,d) \sim \mathrm{N}(0,\sigma_D) \qquad \qquad
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2s) [node, right=of state2H, fill=myBlue]at (4.5,0) { 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        E_{s \neq H i}(x,y,d) 
    \end{array}$
};
\node [right=of state2s, yshift=0.25cm, xshift=-1.25cm, font=\large\bfseries] {DNA (log conc.) };
\node [right=of state2s, yshift=-0.25cm, xshift=-1.1cm, font=\large\bfseries] {species \textit{s} in bottle \textit{i} };
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (metabarcoding) [node, below=of state2s, fill=orange!35]  at (8,-3) {
    $\begin{array}{c}
        \nu_{si} = (E_{si} - E_{Ri} ) + N_{PCR}  \alpha_s  \\  [8pt]
        \pi_{si} = \frac{exp({\nu_{si}})}{\sum_{s=1}^{S} exp({\nu_{si}})} \\ [8pt]
         \mathbf{R}_{i} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{i}, T_{i} \right) 
    \end{array}$
};
\node [below=of metabarcoding, yshift=1.5cm, xshift=0cm, font=\large\bfseries] {Metabarcoding Likelihood};
\node[anchor=south east] at ([xshift=4pt, yshift=-15pt] metabarcoding.south east) {};

\node (mock) [node,  right=of metabarcoding,fill=yellow!50] at (9,-2.75){
    $\begin{array}{c}
       \nu_{sm} = (L_{sm} - L_{Rm}) + N_{PCR}  \alpha_s     \\   [8pt]
       \pi_{sm} = \frac{exp({\nu_{sm}})}{\sum_{s=1}^{S} exp({\nu_{sm}})} \\ [8pt]
       \mathbf{R}_{m} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{m}, T_{m} \right) 
    \end{array}$
};
\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock-community Metabarcoding};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] mock.south east) {};

\node (spatial) [node, right=of state2s, fill=green!40] at (8,2){
    $\begin{array}{c}
        F_s(x,y,d) = s\left(E_{s\neq H}(x,y,d) \right)
    \end{array}$
};
\node [above=of spatial, yshift=-1.50cm, xshift=0cm, font=\large\bfseries] {Smooth DNA predictions (conc.) };
\node[anchor=south east] at ([xshift=19pt, yshift=-4pt] spatial.south east) {};

        % Arrows
        \draw [arrow] (qPCR_st) -- node[pos=0.1, below] {
        $\begin{array}{c}
            \phi, \\ \beta_0, \beta_1, \\ \gamma_0, \gamma_1
        \end{array}$} (qPCR_en);
        \draw [arrow] (state1) -- node[midway, left] {$$} (state2H);
        \draw [arrow] (state2H) -- node[midway, right] {} (qPCR_en);
        \draw [arrow] (state2s) -- node[midway, right] {}  (spatial);
        %\draw [arrow] (state1) -- node[midway, right] {}  (spatial);
        \draw [arrow] (mock) -- node[pos=0.7, above] {$\alpha_s$} (metabarcoding);
        \draw [arrow] (state2H) -- node[midway, right] {$$} (metabarcoding);
        \draw [arrow] (state2s) -- node[midway, right] {$$} (metabarcoding);

% Legend: Parameters
\node (parameters) [rectangle, draw, align=left, below=of metabarcoding, text width=9cm, fill=myGrey] at (3.5,-6){
    \textbf{Parameters} \\
    $\psi$ = mean hake eDNA concentration \\
    %$\delta0, \delta1$ = offset from mean hake \\
    $\delta$ = biological replicate random effect \\
    %$\omega$ = ethanol wash effect \\
    $V_i$ = volume filtered and aliquote dilution offset \\
   % $\mu$ = mean Ct values of qPCR \\
    $\phi$ =  qPCR amplification parameter\\
    $\sigma_{Y}$ = standard deviation of qPCR Ct values \\
    $\beta_{0p}, \beta_{1p}$ = intercept and slope defining mean qPCR - DNA concentration relationship\\
    $\gamma0, \gamma1$ = intercept and slope defining  standard deviation of qPCR - DNA concentration relationship \\
    $\alpha_s$ = amplification efficiency relative to reference species \\
    $\pi$ = predicted sequence proportions at the end of PCR \\
    $\nu$ = sequence proportions at the end of PCR (additive log-ratio)\\
    $\zeta$ = ethanol wash effect \\
    $\tau$ = standard deviation among biological replicates\\
    $\eta$ = spatial field \\
    $\varepsilon$ =  depth specific spatial field
};

% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_en, text width=6cm, fill=myGrey] at (-4.5,-6) {
    \textbf{Index} \\
    $s$ = species ($H$ = Hake) \\
    $x$ = northing \\
    $y$ = easting \\
    $d$ = depth \\
    $i$ = biological replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community sample \\
    $j$ = qPCR standard aliquote sample \\
    $R$ = metabarcoding reference species 
};

% Legend: Data
\node (data) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-6){
    \textbf{Observations } \\
    $Z$ = qPCR amplification (yes=1; no=0) \\
    $Y$ = qPCR Ct values \\
    $K$ = known DNA conc. (log(copies/$\mu$L))\\
    $L$ = known DNA conc. (log(copies/$\mu$L))\\
    $R$ = metabarcoding sequencing reads \\
    $T$ = total number sequencing reads \\
};

% Legend: Data
\node (States of Interest) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-10){
    \textbf{States } \\
    $D$ = DNA conc. at a station-depth (log(copies/$\mu$L)) \\
    $E$ = DNA conc. in a biological replicate (log(copies/$\mu$L)) \\
    $F$ = Smoothed DNA conc. (copies/$L$)\\
    };

    \end{tikzpicture}
    %\caption{Multifish DAG}
\end{figure}

\end{document}

### 2025-02-24 Manuscript_plain_AOS.tex

@@
 Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
-For example, \cite{diaz2020a} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
+For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
 Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time.

### 2025-02-24 Priors table.tex

@@
 \hline
-$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
+$K$ & known DNA concentration from qPCR standards in ln(copies/$\mu$L) & - \\
 $Z$ & qPCR amplification (yes=1; no=0) & - \\
@@
 $N_{PCR}$ & number of metabarcoding PCR cycles & - \\
-$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
+$L$ & known DNA concentration from mock commuinties in ln(copies/$\mu$L) & - \\
 $V$ & volume filtered and aliquote dilution offset & -\\
@@
 \hline
-$E_{s\neq H}$ & unknown DNA concentration in copies/$\mu$L & $\mathcal{N}(0,5)$\\
-$D$ & unknown DNA concentration in copies/L & - \\
-$F$ & smoothed DNA concentration in copies/L & - \\
+$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
+$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
+$F$ & smoothed DNA concentration in ln(copies/L) & - \\
 &&\\
@@
 \hline
-$\psi$& overall mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
 $\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D= (0, +\infty)$& standard deviation of spatial and depth specific random effect & $\mathcal{N}(0,3)$\\
+$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\
 
 $\delta$& biological replicate random effect & $\mathcal{N}(0,\tau)$\\
-$\tau = (0, +\infty)$& standard deviation of biological replicate random effect & $\mathcal{N}(0,0.2)$\\
+$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
 

### 2025-02-24 DAG-Ole.tex
Update DAG based on conversations on Friday.

@@
        % \mu_{ik} = \beta_{0p} + \beta_{1p}  K_{i} \\
-        \sigma_{Yj} = \exp(\gamma_0 + \gamma_1  K_{j}) \\
-        Z_{jpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{j}} \phi) \right) \\
-        Y_{jpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{j} , \sigma_{Yj}) \quad \textnormal{if } Z_{jpr} = 1 \\
+        Z_{qpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{q}} \phi) \right) \\
+        Y_{qpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{q} , \sigma_{Yq}) \quad \textnormal{if } Z_{qpr} = 1 \\
+        \sigma_{Yq} = \exp(\gamma_0 + \gamma_1  K_{q}) \\
     \end{array}$
@@
        % \mu_{ip} = \beta_{0p} + \beta_{1p} E_{Hi} \\
-        \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})\\
-         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hi}} \phi) \right) \\
-         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hi} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1
+         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hji}} \phi) \right) \\
+         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hji} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1 \\
+         \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})
     \end{array}$
@@
         %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
-        D_{H}(x,y,d) = \psi  + \epsilon({x,y,d}) \\
-        \epsilon({x,y,d})\sim \mathrm{N}(0,\tau)
+        D_{Hj} = \psi  + \epsilon_j \\
+        \epsilon_j \sim \mathrm{N}(0,\tau)
     \end{array}$
@@
 
-\node (state2H) [node, above=of qPCR_en, fill=myBlue] at (2.5,-2.5){ 
+\node (state2H) [node, above=of qPCR_en, fill=myBlue] at (2.5,-2){ 
 % + \delta0_{ijdbt} 
@@
         %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
-       E_{Hi}(x,y,d) = D_H({x,y,d}) +  \log{V_i} + \\
-       		\qquad \qquad \qquad \qquad \log{I_i} + \mathbf{I}\zeta + \delta_i(x,y,d) \\
-		\delta_i(x,y,d) \sim \mathrm{N}(0,\sigma_D) \qquad \qquad
+       E_{Hji} = D_{Hj} +  \log{V_i} +\log{I_i} + \mathbf{I}\zeta + \delta_{ji} \\
+       		%\qquad \qquad \qquad \qquad 
+		\delta_{ji} \sim \mathrm{N}(0,\sigma_D) \qquad \qquad
     \end{array}$
@@
 
-\node (state2s) [node, right=of state2H, fill=myBlue]at (4.5,0) { 
+\node (state2s) [node, right=of state2H, fill=myBlue]at (4.5,0.1) { 
 % + \delta0_{ijdbt} 
@@
         %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
-        E_{s \neq H i}(x,y,d) 
+        E_{s \neq Hji}
     \end{array}$
@@
     $\begin{array}{c}
-        \nu_{si} = (E_{si} - E_{Ri} ) + N_{PCR}  \alpha_s  \\  [8pt]
+        \nu_{si} = (E_{sji} - E_{Rji} ) + N_{PCR}  \alpha_s  \\  [8pt]
         \pi_{si} = \frac{exp({\nu_{si}})}{\sum_{s=1}^{S} exp({\nu_{si}})} \\ [8pt]
@@
 
-\node (spatial) [node, right=of state2s, fill=green!40] at (8,2){
+\node (spatial) [node, right=of state2s, fill=green!40] at (9,2){
     $\begin{array}{c}
-        F_s(x,y,d) = s\left(E_{s\neq H}(x,y,d) \right)
+        F_s(x,y,d) 
     \end{array}$
@@
         % Arrows
-        \draw [arrow] (qPCR_st) -- node[pos=0.1, below] {
-        $\begin{array}{c}
-            \phi, \\ \beta_0, \beta_1, \\ \gamma_0, \gamma_1
-        \end{array}$} (qPCR_en);
+        \draw [arrow] (qPCR_st) -- node[pos=0.1, below] {} (qPCR_en);
+        % \begin{array}{c}
+        %     \phi, \\ \beta_0, \beta_1, \\ \gamma_0, \gamma_1
+        % \end{array}$} (qPCR_en);
         \draw [arrow] (state1) -- node[midway, left] {$$} (state2H);
         \draw [arrow] (state2H) -- node[midway, right] {} (qPCR_en);
-        \draw [arrow] (state2s) -- node[midway, right] {}  (spatial);
+        \draw [dashed,arrow] (state2s) -- node[midway, right] {}  (spatial);
         %\draw [arrow] (state1) -- node[midway, right] {}  (spatial);
@@
     $s$ = species ($H$ = Hake) \\
+    $j$ = location (northing, easting, depth) \\
     $x$ = northing \\
@@
     $m$ = mock community sample \\
-    $j$ = qPCR standard aliquote sample \\
+    $q$ = qPCR standard aliquote sample \\
     $R$ = metabarcoding reference species 

### 2025-02-25 Priors table.tex

--- 2025-02-21 Priors table.tex
@@
 \hline
-$\psi$& overall mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\psi$& overall mean hake eDNA concentration (copies/L) & $\mathcal{N}(2,2)$\\
 $\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\

### 2025-02-25 Priors table.tex

--- 2025-02-24 Priors table.tex
@@
 \hline
-$K$ & known DNA concentration from qPCR standards in ln(copies/$\mu$L) & - \\
+$K$ & known DNA concentration from qPCR stsandards in ln(copies/$\mu$L) & - \\
 $Z$ & qPCR amplification (yes=1; no=0) & - \\

### 2025-02-25 Appendix_plain_AOS.tex
Updates from Overleaf

@@
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
-Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water.
+Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see Table S.\ref{tab:mock_table} for DNA quantification results).
 The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read.
@@
 
-\subsection*{Model convergance}
+\section*{Model convergance}
+Convergence and reliability of the Bayesian model were assessed through comprehensive diagnostics.
+All parameters (Table S.\ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (Fig.
+S.\ref{tab:diagnostics}).
+An exception was the parameter $\psi$ (mean hake), which showed marginally lower mixing ESS$_\psi$ = 149.7) suggesting mild autocorrelation but no critical failure of convergence.
 
-Gelman-Rubin statistics.  Predicted-observed, etc.
+No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (Fig.
+S.\ref{tab:diagnostics}).
+The posterior likelihood demonstrated convergence before the sampling phase began, with all chains exhibiting high mixing, confirming robust exploration of the parameter space (Fig.
+S.\ref{tab:diagnostics}).
+
+Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (Fig.
+S.\ref{tab:prior_sensitivity}).
+Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (Fig.
+S.\ref{tab:PPC}).
+
+Collectively, these diagnostics confirm the reliability and validity of the Bayesian model used here.
 
@@
 \caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
+\label{tab:PPC}
 \end{figure}
@@
 \caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.88\textwidth]{plots/11_Diagnostic_Fig_1.jpg}
+\caption{Bayesian model diagnostic plots including $\hat{R}$ and ESS for each parameter (A) and posterior likelihood, acceptance rate, tree depth, divergence transition, and step size for each iteration in each chain (including warm up phase; B).}
+\label{tab:diagnostics}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/12_Diagnostic_Fig_2.jpg}
+\caption{Prior sensitivity plots indicating the distribution of priors in red and posteriors in black with dots indicating the mean and ticks showing 95\% credible interval.}
+\label{tab:prior_sensitivity}
 \end{figure}
@@
     \end{tabular}
+    \label{tab:mock_table}
 \end{table}
 \clearpage
-\bibliography{Bib.bib}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{lll}
+         & \textbf{Description} & \textbf{Prior} \\
+&Data & \\
+\hline
+$K$ & known DNA concentration from qPCR stsandards in ln(copies/$\mu$L) & - \\
+$Z$ & qPCR amplification (yes=1; no=0) & - \\
+$Y$ & qPCR cycle threshold (Ct) & - \\
+$R$ & metabarcoding number of reads & - \\
+$T$ & total number of reads for all S species & - \\
+$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
+$L$ & known DNA concentration from mock commuinties in ln(copies/$\mu$L) & - \\
+$V$ & volume filtered and aliquote dilution offset & -\\
+$I$ & aliquote dilution factor & -\\
+
+&&\\
+&State processes&\\
+\hline
+$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
+$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
+$F$ & smoothed DNA concentration in copies/L & - \\
+&&\\
+
+&Parameters&\\
+\hline
+$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\tau)$\\
+$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
+
+$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
+$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\
+
+$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
+$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
+&and eDNA concentration (K and E) & \\
+$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
+&eDNA concentration (K and E) & \\
+$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (Table S.\ref{tab:alphas})\\
+$\zeta$& ethanol wash effect & fixed (-1.15)\\
+&&\\
+&Index&\\
+\hline
+$s$& species ($H$ = Hake) & -\\
+$j$& location (northing(x), easting(y), depth(d)))& -\\
+$i$& biological replicate & -\\
+$r$& technical replicate & -\\
+$p$& qPCR plate & -\\
+$m$& mock community sample & -\\
+$q$& qPCR standard aliquote sample & -\\
+$R$& metabarcoding reference species & -\\
+
+    \end{tabular}
+    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
+    \label{tab:priortable}
+\end{table}
+
+\clearpage
+\begin{table}\centering
+\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} here).}
+    \begin{tabular}{lc}
+\textbf{Species} & $\alpha$ \\
+\textit{Clupea pallasii} &   -0.00138 \\
+\textit{Engraulis mordax} &    0.00557 \\
+\textit{Leuroglossus stilbius} &   -0.00390 \\
+\textit{Microstomus pacificus} &    0.00194 \\
+\textit{Sardinops sagax} &    0.00705 \\
+\textit{Scomber japonicus} &    0.00434 \\
+\textit{Sebastes entomelas} &    0.00459 \\
+\textit{Stenobrachius leucopsarus} &    0.00970 \\
+\textit{Tactostoma macropus} &   -0.00091 \\
+\textit{Tarletonbeania crenularis} &    0.00144 \\
+\textit{Thaleichthys pacificus} &   -0.00506 \\
+\textit{Trachurus symmetricus} &   -0.00875 \\
+\textit{Merluccius productus} &    0.00000 \\
+    \end{tabular}
+        \label{tab:alphas}
+\end{table}
+\clearpage
+
+\bibliography{clean_bib.bib}
 \end{document}

### 2025-02-25 Appendix_plain_AOS.tex
Updates from Overleaf

@@
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
-Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see Table S.\ref{tab:mock_table} for DNA quantification results).
+Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see SI Appendix Table S.\ref{tab:mock_table} for DNA quantification results).
 The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read.
@@
 We developed a joint model for modeling DNA concentration in the coastal ocean using single-species qPCR data Pacific hake and metabarcoding data from the 12S MiFish Universal primer.
-This work builds off previous single-species eDNA models constructed for Pacific hake (Shelton et al.
-2022) and models for joining qPCR and metabarcoding data (Guri et al.
-2024).
+This work builds off previous single-species eDNA models constructed for Pacific hake \cite{shelton2022} and models for joining qPCR and metabarcoding data \cite{guri2024a}.
 Here qPCR observations provide information about DNA concentration for one species (Pacific hake) and the metabarcoding data provide information about species composition within a sample (i.e.
@@
 
-We developed a state-space framework for modeling hake DNA concentration in the coastal ocean building and modifying the work of Shelton et al.
-(2022).
+We developed a state-space framework for modeling hake DNA concentration in the coastal ocean building and modifying the work of \cite{shelton2022}.
 State-space models separate the true biological process from the methods used to observe the process.
@@
 Hereafter we refer to a combination of spatial location and depth as a ``station".
-Let $\psi_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
+Let $\psi$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
 We write these as additive on the log-scale,
@@
 \begin{align}
-  D_H(x,y,d) &= \psi_0 + \epsilon(x,y,d) \\
+  D_H(x,y,d) &= \psi + \epsilon(x,y,d) \\
   \epsilon(x,y,d) &\sim N(0,\sigma_D)
@@
 For water sample $i$: 1) $V_i$ is the proportion of 2.5 L filtered from Niskin $i$ (occasionally some seawater was spilled or not filtered; $V_i =1$ for the vast majority of samples); 2) $I_i$ is the known dilution used to on sample $i$ to eliminate PCR inhibition.
-PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see Shelton et al.
-2022 Supp S1 for additional description).
-Here we fix $\zeta$ = XXX based on previous analyses (Shelton et al.
-2022).
+PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see \cite{shelton2022} Supp S1 for additional description).
+Here we fix $\zeta$ = -1.15 based on previous analyses \cite{shelton2022}.
 
@@
 Again, we are interested in determining the DNA concentration of each focal species in each sample.
-Unfortunately, metabarcoding data is compositional data \cite{glor2017, shelton2023}, and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
+Unfortunately, metabarcoding data is compositional data \cite{gloor2017, shelton2023}, and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
 Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
-Here, we follow and extend the metabarcoding model developed by Shelton et al.
-(2023) and show how it links with the qPCR model above.
+Here, we follow and extend the metabarcoding model developed by \cite{shelton2023} and show how it links with the qPCR model above.
 
@@
 Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see \cite{shelton2023}).
-We can then calculate $\nu_{si}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see *Mock Communities* below),
+We can then calculate $\nu_{si}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see Mock Communities below),
 
@@
 Convergence and reliability of the Bayesian model were assessed through comprehensive diagnostics.
-All parameters (Table S.\ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (Fig.
+All parameters (SI Appendix Table S.\ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (SI Appendix Fig.
 S.\ref{tab:diagnostics}).
@@
 
-No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (Fig.
+No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (SI Appendix Fig.
 S.\ref{tab:diagnostics}).
@@
 
-Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (Fig.
+Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (SI Appendix Fig.
 S.\ref{tab:prior_sensitivity}).
-Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (Fig.
+Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (SI Appendix Fig.
 S.\ref{tab:PPC}).
@@
 
+\begin{figure*}[tbhp] 
+\centering
+\includegraphics[width=16cm]{plots/DAG-Ole.pdf} 
+\caption{Schematic for the statistical models to produce species distributions.
+Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
+Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
+Notation for parameters, subscripts, and observations are indicated in gray.}
+\label{fig:DAG2}
+\end{figure*}
+
 %%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.89\textwidth]{plots/4_Figure_4.jpg}
-\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.89\textwidth]{plots/4_Figure_4.jpg}
+%\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
 Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A).
@@
 Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
-\end{figure}
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
-\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
-\label{tab:PPC}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
 \includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
@@
 \label{tab:prior_sensitivity}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
+\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
+\label{tab:PPC}
 \end{figure}
@@
 &of Ct values and eDNA concentration (K and E) & \\
-$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (Table S.\ref{tab:alphas})\\
+$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (SI Appendix Table S.\ref{tab:alphas})\\
 $\zeta$& ethanol wash effect & fixed (-1.15)\\

### 2025-02-25 Manuscript_plain_AOS.tex

@@
 \ref{fig:fig1}).
-eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
+eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; Fig.
+\ref{fig:fig4}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
 Furthermore, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
@@
 Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat well.
-We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
+We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Fig.
+\ref{fig:fig1} and \ref{fig:fig4}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
 
@@
 Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
-The eDNA concentrations appeared to reflect this activity (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
+The eDNA concentrations appeared to reflect this activity (Fig.
+\ref{fig:fig4}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
 \ref{fig:fig1}), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
@@
 Also consistent with known habitats are commercially caught flatfish such as Dover sole (\textit{Microstomus pacificus}), which lives at 200m to 1200m along the Pacific coast \cite{drazen2007,brodziak2000,drazen2012}; individuals move progressively deeper as they grow \cite{vetter1994,hunter1990}.
-The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}\textit{A}) \cite{ono2016}.
+The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (Fig.
+\ref{fig:fig4}\textit{A}) \cite{ono2016}.
 
@@
 Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$) with both more common in the more southern (and somewhat warmer) waters (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
+\ref{fig:fig1} and Fig.
+\ref{fig:fig4}).
 Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig.
-\ref{fig:fig2}) consistent with more northerly distribution (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
+\ref{fig:fig2}) consistent with more northerly distribution (Fig.
+\ref{fig:fig4}).
 Together these finding suggest that eDNA is capturing both depth and latitudinal habitats associations.
 Deep-water species are distinct (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}) and not strongly correlated with those from other zones (Fig.
+\ref{fig:fig1} and Fig.
+\ref{fig:fig4}) and not strongly correlated with those from other zones (Fig.
 \ref{fig:fig2}).
 Deeper water tends to feature lower and more diffuse fish eDNA concentrations (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
+\ref{fig:fig1} and Fig.
+\ref{fig:fig4}).
 Lower overall eDNA concentrations and fewer deep water samples may make similarities among species distributions more difficult to detect at 300m and deeper.
@@
 In addition to providing species-specific patterns, eDNA metrics of abundance from individual species can be combined to reveal important aggregate ecological patterns.
-For the 12 species considered here, we show how species richness (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
+For the 12 species considered here, we show how species richness (see \href{SI_Appendix.pdf}{SI Appendix Fig.
+S3} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
 \ref{fig:fig3}\textit{A}).
@@
 \caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m).
-(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
+(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see Fig.
+\ref{fig:fig4} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
 \label{fig:fig3}
@@
 
-%\begin{figure*}[tbhp]  
-%\centering
-%\includegraphics[width=17cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
-%\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
+\begin{figure*}[tbhp]  
+\centering
+\includegraphics[width=17cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
 Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A).
@@
 Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
-%\label{fig:fig4}
-%\end{figure*}
+\label{fig:fig4}
+\end{figure*}
 
@@
 
+Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+S6}) and spatially smoothed maps (Fig.
+\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S1}) revealed minimal differences.
 Large-scale sampling combined with statistical smoothing mitigates many challenges associated with eDNA transport and degradation -- which decouple the eDNA observations from their particular source organisms -- allowing reliable detection of broad patterns in species distributions.
 Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature or other covariates \cite{jo2019b}.
-
-%Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}) and spatially smoothed maps (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound.
-This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.
-\hl{[Check the framing of the last two sentences XXX]}
-
-%(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S6}) posterior predictive
 
@@
 
-%Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
-In reality, fish can move in between sampling location within the study area, which may bias abundance estimates  \cite{rota2009}.
-By contrast, fish continuously shed DNA into their surroundings, creating a dynamic signal of presence.
-As the shedded DNA decays, disperses, and degrades over time, fresher DNA remains at higher concentrations while older DNA persists at lower concentrations \cite{thomsen2015,jerde2011}.
+Conventional fish surveys, such as trawling or mark-recapture methods, typically rely on a closure assumption assuming that fish populations are static during the sampling period \cite{bailey2014}, ignoring the reality of fish movement within the study area thus introducing potential bias into abundance estimates \cite{rota2009}.
+By contrast, eDNA analysis leverages the continuous shedding of DNA by fish, which decays over time with higher concentrations remaining fresh while older DNA persists at lower levels \cite{thomsen2015,jerde2011}.
 This cumulative, time-integrated signal allows eDNA surveys to circumvent the closure assumption and provide a more accurate representation of fish presence and activity.
@@
 \section*{Methods}
-We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
+We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
 
@@
 
-\begin{figure*}[tbhp] 
-\centering
-\includegraphics[width=16cm]{plots/DAG-Ole.pdf} 
-\caption{Schematic for the statistical models to produce species distributions.
-Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
-Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
-Notation for parameters, subscripts, and observations are indicated in gray.
-See \href{SI_Appendix.pdf}{SI Appendix} for full model description.}
-\label{fig:DAG2}
-\end{figure*}
-
 \subsection*{Metabarcoding: Environmental Samples and Mock Community}
 In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
-Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix} for details on metabarcoding methods).
+Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods} for details on metabarcoding methods).
 Importantly, we only analyzed one water sample collected from each sampling station using metabarcoding.
@@
 Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix}).
-Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
+S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
+Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
 
@@
 To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples.
-The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig.
-\ref{fig:DAG2}).
+The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}, with an overview of the approach (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
 We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
@@
 \ref{fig:DAG2}).
-This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (Fig.
-\ref{fig:DAG2}).
-This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (Fig.
-\ref{fig:DAG2}).
+This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
+This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
 
@@
 
-The above statistical model (Fig.
-\ref{fig:DAG2}) provides predictions of species-specific DNA concentrations (\textit{D}) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
-To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (Fig.
-\ref{fig:DAG2}, green box).
+The above statistical model (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2) provides predictions of species-specific DNA concentrations (\textit{D}) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2, green box).
 

### 2025-02-25 Appendix_plain_AOS.tex
Updates from Overleaf

@@
 $\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
-$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
+$\beta_0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
 &and eDNA concentration (K and E) & \\
-$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
+$\beta_1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
 &eDNA concentration (K and E) & \\
-$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
+$\gamma_0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
 &of Ct values and eDNA concentration (K and E) & \\
-$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
+$\gamma_1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
 &of Ct values and eDNA concentration (K and E) & \\

### 2025-02-27 Appendix_plain_v2.tex
push ole and owen edits to DAG and appendix.

--- 2025-02-14 Appendix_plain.tex
@@
 State-space models separate the true biological process from the methods used to observe the process.
-Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA $copies$ $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
-Let $\alpha_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
+Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA copies $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
+Hereafter we refer to a combination of spatial location and depth as a ``station" and index station using $j$ (thus, $D_{Hj} = D_H(x,y,d)$).
+Let $\psi$ be the intercept representing the mean hake DNA concentration and $\epsilon_j$ be a random effect that allows for the effect of that station.
 We write these as additive on the log-scale,
@@
 \begin{align}
-  D_H(x,y,d) &= \psi_0 + \epsilon(x,y,d) \\
-  \epsilon(x,y,d) &\sim N(0,\sigma_D)
+  D_{Hj} &= \psi + \epsilon_j \\
+  \epsilon_j &\sim  \mathrm{Normal}(0,\tau)
 \end{align}
-
 This model describes perhaps the simplest possible random effect model and notably does not leverage any information about the spatial arrangement of samples (i.e.
@@
 2022 Supp S1 for additional description).
+Here we fix $\zeta$ = \textbf{XXX} based on previous analyses (Shelton et al.
+2022).
 
-Finally, we add a random effect for the individual bottles sampled at each year-station-depth combination ($\delta_i(x,y,d)$).
-This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_i(x,y,d) \sim N(0,\tau)$, with $\tau$ indicating a depth-specific standard deviation among bottles.
+Finally, we add a random effect for the individual bottles sampled at each station ($\delta_{ij}$).
+This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_{ji} \sim  \mathrm{Normal}(0,\sigma_D)$, with $\sigma_D$ indicating a  standard deviation among bottles.
 Because we have only two replicate bottles at each location, we impose a sum to zero constraint on the bottles collected from a single location (i.e.
-for a given location, $\sum_i \delta_i(x,y,d) = 0$).
-Then, the log-concentration of hake DNA in a sample analyzed by qPCR, $E_{Hi}$ is,
+for a given location, $\sum_i \delta_{ji} = 0$).
+Then, the log-concentration of hake DNA in a sample analyzed by qPCR, $E_{Hji}$ is,
 
 \begin{align}
-  E_{Hi} &= D_H({x,y,d}) +  \log{V_i} + \log{I_i} + \mathbf{I}\zeta + \delta_i(x,y,d)
+  E_{Hji} &= D_{Hj} +  \log{V_i} + \log{I_i} + \mathbf{I}\zeta + \delta_{ji}
 \end{align}
-
 We connect the estimated concentrations to observations from qPCR using two likelihoods connected to the concentration $E$.
-First we determine whether amplification was detected in each qPCR replicate $r$ run on each plate $j$, $Z_{Hijr}$ and if amplification was observed we model the PCR cycle at which amplification was detected $Y_{Hijr}$; $Y_{Hijr}$ is observed as a continuous, positive value.
-DNA concentrations have a log-linear relationships with $Y_{Hijr}$; smaller values of $Y_{Hijr}$ are associated with higher DNA concentrations.
+First we determine whether amplification was detected in each qPCR replicate $r$ run on each plate $p$, $Z_{Hjipr}$ and if amplification was observed we model the PCR cycle at which amplification was detected $Y_{Hjipr}$; $Y_{Hjipr}$ is observed as a continuous, positive value.
+DNA concentrations have a log-linear relationships with $Y_{Hjipr}$; smaller values of $Y_{Hjipr}$ are associated with higher DNA concentrations.
 
@@
 PCR inhibition) that may reduce amplification further and therefore we expect the probability of amplification to be at most $1-e^{-X}$.
-As a result, we estimate an additional term, $\phi$ representing the fractional reduction in amplification efficiency due to other factors on qPCR plate $j$ ($0<\phi<1$).
-Then we have a pair of observation models for hake DNA, both of which are a function of DNA concentration, $E_{Hi}$
+As a result, we estimate an additional term, $\phi$, representing the fractional reduction in amplification efficiency due to other factors during qPCR ($0<\phi<1$).
+Then we have a pair of observation models for hake DNA, both of which are a function of DNA concentration, $E_{Hji}$,
 
 \begin{align}
-  Z_{Hijr} &\sim Bernoulli\left(1 - \exp(-2 e^{E_{Hij}}\phi) \right) \\ 
-  Y_{Hijr} &\sim Normal \left(\beta_{0j}+\beta_{1j}{E_{Hij}},\sigma_Y(E_{Hij}) \right) \qquad  if \: Z_{Hijr} = 1 
+  Z_{Hjipr} &\sim  \mathrm{Bernoulli} \left(1 - \exp(-2 e^{E_{Hji}}\phi) \right) \\ 
+  Y_{Hjipr} &\sim  \mathrm{Normal} \left(\beta_{0p}+\beta_{1p}{E_{Hji}},\sigma_{Yji} \right) \qquad   \mathrm{if} \: Z_{Hjipr} = 1 
 \end{align} 
-
-The 2 is present in the first line of the equation because we use 2 $\mu L$ of sample in each qPCR reaction and so the expected DNA copies in reaction is $2e^{E_{ij}}$.
-We allow $\sigma_C(E_{ij})$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_C({E_{ij}}) = exp(\gamma_{0}+\gamma_{1}E_{ij})$ (see also Fig.
-XX).
+The 2 is present in the first line of the equation because we use 2 $\mu$L of sample in each qPCR reaction and so the expected DNA copies in reaction is $2e^{E_{Hji}}$.
+We allow $\sigma_{Yji}$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_{Yji} = exp(\gamma_{0}+\gamma_{1}E_{Hji})$.
 
@@
 These standards span six orders of magnitude (1 to 100,000 copies $\mu L^{-1}$, each of 2$\mu L$) and determine the relationship between copy number and PCR cycle of detection.
-Let $K_{jp}$ be the known copy number for sample $j$ in PCR plate $p$, then,
+Let $K_{qp}$ be the known log-copy number for sample $q$ in PCR plate $p$, then,
 
 \begin{align}
-  Z_{jp} &\sim Bernoulli \left(1-exp(-2 K_{jp} \phi) \right)\\
-  Y_{jp} &\sim Normal \left(\beta_{0j}+\beta_{1j}\log{K_{jp}},\sigma_Y(K_{jp})\right) \quad  if \: Z_{jp} =1
+  Z_{qpr} &\sim  \mathrm{Bernoulli} \left(1-exp(-2 e^{K_{qp}} \phi) \right)\\
+  Y_{qpr} &\sim  \mathrm{Normal} \left(\beta_{0j}+\beta_{1j}{K_{qp}},\sigma_{Yqp}\right) \quad  \mathrm{if} \: Z_{qjp} =1
 \end{align}
-
-Note that there are different intercept $\beta_{0p}$ and slope $\beta_{1p}$ parameters for each PCR plate to allow for among-plate variation in amplification.
-We model each calibration paratemeter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
-$\beta_{0p} \sim N(\mu_{\beta_{0}},\sigma_{\beta_{0}})$).
+Note that there are different intercept ($\beta_{0p}$) and slope ($\beta_{1p}$) parameters for each PCR plate to allow for among-plate variation in amplification.
+We model each calibration parameter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
+$\beta_{0p} \sim \mathrm{Normal}(\mu_{\beta_{0}},\sigma_{\beta_{0}})$).
 
@@
 Again, we are interested in determining the DNA concentration of each focal species in each sample.
-Unfortunately, metabarcoding data is compositional data (see *Glor et al.
-*, *Shelton et al.
-2023*), and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
+Unfortunately, metabarcoding data is compositional data \cite{shelton2023}, and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
 Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
@@
 
-As above, we let $D_s(x,y,d)$ be the log-scale eDNA concentration but now for species $s$.
-We let $E_{si}(x,y,d)$ be the log-DNA concentration for each species in each bottle.
+As above, we let $D_{sj}$ be the log-scale eDNA concentration but now for species $s$.
+We let $E_{sji}$ be the log-DNA concentration for each species in each bottle.
 Importantly one of the $s$ species is Pacific hake and so connects to the qPCR model above.
-Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; *Aitchison XXXX*).
-We define a reference species in each sample, $R$, with the species used as a reference varying among sample.
-Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see Shelton et al.
-2023).
-We can then calculate $\nu_{si}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see *Mock Communities* below),
+Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; \textbf{Aitchison 1986}).
+We define a reference species in each sample, $R_i$, with the species used as a reference varying among sample.
+Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see \cite{shelton2023}).
+We can then calculate $\nu_{sji}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see \textit{Mock Communities} below),
 
 \begin{align} 
-  \nu_{si} &= \left( E_{si}(x,y,d) - E_{Ri}(x,y,d)\right) + N_{PCR} \left(\alpha_s - \alpha_R \right) 
+  \nu_{sji} &= \left( E_{sji} - E_{Rji}\right) + N_{PCR} \left(\alpha_s - \alpha_{R_i} \right) 
 \end{align}
-
-To maintain identifiability, we assign one $\alpha_s = 0$.
-We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\mu_{si}$ using the softmax transformation and to the observed reads via a Dirichlet-Multinomial likelihood,
+We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\pi_{sji}$ using the softmax transformation and to the observed reads via a multinomial likelihood,
 
 \begin{align} 
-  \pi_{si} &= \dfrac{e^{\nu_{si}}}{\sum_{s=1}^S e^{\nu_{si}}} \\
-  \boldsymbol{R}_{i} & \sim Multinomial\left(\boldsymbol{\pi}_{i},T_{i}\right) 
+  \pi_{sji} &= \dfrac{e^{\nu_{sji}}}{\sum_{s=1}^S e^{\nu_{sji}}} \\
+  \boldsymbol{W}_{ji} & \sim  \mathrm{Multinomial}\left(\boldsymbol{\pi}_{ji},T_{i}\right) 
 \end{align}
-
 where $T_{i}$ is the total number of reads observed across all species for that bottle.
@@
 As with the qPCR analyses, in the absence of additional information estimates of $\alpha$ and $E$ cannot be statistically identified.
-We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but for the moe
+We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but instead of having unknown DNA concentrations, we have known inputs for each species $L_{sm}$ in mock community $m$.
+
+\begin{align} 
+ \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{R_i} \right) \\
+  \pi_{sm} &= \dfrac{e^{\nu_{sm}}}{\sum_{s=1}^S e^{\nu_{sm}}} \\
+  \boldsymbol{W}_{m} & \sim  \mathrm{Multinomial}\left(\boldsymbol{\pi}_{m},T_{m}\right) 
+\end{align}
+To maintain identifiability, we assign one $\alpha_s = 0$.
 
@@
 
-\textbf{OWEN.  COULD YOU HELP HERE?}
+The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
+To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model, exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
+Because of the predominance of very low estimated DNA concentrations at a large number of site, we set $G_{s}(x,y,d) < 0.5$ equal to zero and then used a Tweedie likelihood in sdmTMB,
 
-The above statistical model provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as:
-$$\ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)$$
-where x,y,z index longitude, latitude, and depth, respectively, $\tau$ is scalar intercept, $\eta(s)$ is a spatial smooth shared among depths, and $\varepsilon_d(s)$ is a depth-specific spatial smooth.
-Both $\eta(s)$ and $\varepsilon_d(s)$ are a zero-centered Gaussian random field estimated using the stochastic partial differntial equation (SPDE) approximation (see \cite{anderson2022}).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model.
+\begin{align} 
+	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \varepsilon_{d}(x,y,d) \\
+	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
+\end{align}
+where x, y, and d index longitude, latitude, and depth, respectively.
+$\kappa$ is scalar intercept.
+Both $\eta(x,y)$ and $\varepsilon_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\varepsilon_d \sim  \mathrm{MVN}(0,\sum_{\varepsilon_d})$ (see \cite{anderson2022}).
+$\eta$ is a depth-invariant spatial field, while $\varepsilon_d$ is a spatial field specific to each depth category.
+In simple terms, each station-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific.
+$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\varepsilon_d$ modulates that estimate for different sampling depths.
+The $\varepsilon_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
+
+Estimation of large, multi-level spatiotemporal models can strain computational power, in part because of the need in spatial models to invert large covariance matrices between the data.
+We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (Lindgren et al.
+2024; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh (Lindgren et al.
+2011).
+The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\varepsilon_d$.
+Model estimates and the degree of spatial smoothing can be affected by the details of mesh construction, and we therefore produced a range of meshes with different complexities (i.e., varying numbers of triangles, triangle sizes, and therefore vertices) to test for each species.
+
+Each species was fitted individually using the equation above.
+We tested 18 candidate meshes for each species, increasing in complexity from 33 to 205 vertices, as well as allowing each model to include or exclude the scalar intercept $\kappa$.
+Each candidate model for each species was fit by maximum marginal likelihood, implemented in the sdmTMB package in R ( \cite{anderson2022} ).
+Model selection was carried out by first ensuring proper convergence, then choosing the model for each species that maximized the summed likelihood.
+
+We present results based on the predicted smooth mean concentration, $F_s(x,y,d)$, in all results.
 
 \subsection*{Model convergance}
-OLE + GLED
 

### 2025-02-27 Priors table.tex

@@
 $\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
-$\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
+$\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
 $\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\

### 2025-02-27 Manuscript_plain_AOS.tex
Updates from Overleaf

@@
 \section*{Keywords}
-eDNA $|$ Bayesian $|$ bio-monitoring $|$ quantification $|$ fish
+eDNA $|$ natural resources $|$ bio-monitoring $|$ quantification $|$ fish
 
@@
 Northern anchovy (\textit{Engraulis mordax}) and Pacific sardine (\textit{Sardinops sagax}) are the two iconic coastal pelagic species of the eastern North Pacfic \cite{steinbeck1945} and play a vital ecological role linking lower and upper trophic levels (e.g., Chinook salmon diets, sea lions, \cite{kaplan2019}); historically these supported large commercial fisheries particularly in California.
-Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2°N; Fig.
+Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2$^\circ$N; Fig.
 \ref{fig:fig1}).
-Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24°C for sardine vs.
-10–14°C for anchovy).
+Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24$^\circ$C for sardine vs.
+10–14$^\circ$C for anchovy).
 We found more Pacific sardine DNA in sub-surface waters as well (Fig.
 \ref{fig:fig1}).
-eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; Fig.
+eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41$^\circ$N (range 40$^\circ$–45$^\circ$), while anchovy eDNA abundance was greatest at 43.5$^\circ$N (range 42$^\circ$–48$^\circ$; Fig.
 \ref{fig:fig4}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
 Furthermore, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
-However, due to differences in stock definitions and survey areas (the anchovy stock assessment includes southern California waters (32.5°–36° N) not covered by the eDNA survey) these magnitudes are not directly comparable.
+However, due to differences in stock definitions and survey areas (the anchovy stock assessment includes southern California waters (32.5$^\circ$–36$^\circ$N) not covered by the eDNA survey) these magnitudes are not directly comparable.
 
@@
 Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common, yet under studied, such as the dragonfish (\textit{Tactostoma macropus}).
-Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
+Available literature suggests dragonfish occur roughly between 40$^\circ$N-50$^\circ$N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
 The eDNA concentrations appeared to reflect this activity (Fig.
@@
 
-Pairwise correlations of DNA concentration showed interesting patterns among species associated with depth (Fig.
+Pairwise correlations of DNA concentration showed meaningful patterns among species associated with depth (Fig.
 \ref{fig:fig2}).
@@
 \ref{fig:fig4}).
-Together these finding suggest that eDNA is capturing both depth and latitudinal habitats associations.
+Together these findings suggest that eDNA is capturing both depth and latitudinal habitat associations.
 Deep-water species are distinct (Fig.
@@
 \caption{Correlation plot describing species co-abundance and co-occurance.
-Spearman’s rho correlation coefficient are expressed in numbers and colors ranging from -1 (red) negatively correlated to 1 (blue) positively correlated.
+Spearman’s $\rho$ correlation coefficient are expressed in numbers and colors ranging from -1 (red) negatively correlated to 1 (blue) positively correlated.
 Species are ordered in known ecological groups (surface, midwater, and deepwater) to visually indicate the gradual change of species co-occurance.}
@@
 \centering
-\includegraphics[width=17cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=16cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
 \caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
@@
 We report species-specific DNA concentrations smoothed over ten degrees of latitude and across the continental shelf to 500m.
-We reveal distributional pattern for 12 species ranging from the iconic coastal pelagic species of the northeast Pacific to understudied-but-abundant deepwater species, and our results suggest clear pathways for the broader use of eDNA in both fisheries and ecosystem management.
+We reveal distributional patterns for 12 species ranging from the iconic coastal pelagic species of the northeast Pacific to understudied-but-abundant deepwater species, and our results suggest clear pathways for the broader use of eDNA in both fisheries and ecosystem management.
 
@@
 \subsection*{Environmental Smoothing}
-Our findings underline eDNA as a transformative tool for fisheries management, biodiversity conservation, and the study of environmental change.
+Our findings illustrate the power of eDNA as a transformative tool for fisheries management, biodiversity conservation, and the study of environmental change.
 However, the observable eDNA signal at any given location and time is an integrated product of DNA production and loss across time and space \cite{gold2023}: the vast majority of DNA detectable in a given sample is of recent origin (within the previous hours to days \cite{strickler2015,collins2018,machler2018}), and it may be subject to a degree of physical transport within its window of detectability.
@@
 Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey \cite{deblois2020}.
-2.5L water samples were collected from each Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
+2.5L water samples were collected from each Niskin and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
 A subset of the samples collected in \cite{shelton2022}, was used for this study.
@@
 The sequences were then BLASTed against the NCBI nucleotide (nt) database (access: August 2024) using BLASTn algorithm with a cut off at 97\% identity, assigning the least common ancestor of the top hit, resulting in species, genus, family, or higher rank classifications.
-Samples with fewer total number of reads than 1000 were filtered as indicator of poor reliability.
+Samples with fewer total number of reads than 1000 were filtered as an indicator of poor reliability.
 Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
@@
 We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
-This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig.
-\ref{fig:DAG2}).
+This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
 This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (\href{SI_Appendix.pdf}{SI Appendix} Fig.
@@
 Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification \cite{shelton2023}, as well as on the total sequencing depth for that sample \cite{kelly2019}.
-Higher sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
+Greater sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
 

### 2025-02-27 Appendix_plain_v1.tex
Updates from Overleaf

--- 2025-02-25 Appendix_plain_AOS.tex
@@
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
-Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see Table S.\ref{tab:mock_table} for DNA quantification results).
+Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water.
 The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read.
@@
 \begin{align} 
- \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) \\
+ \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) 
   \pi_{sm} &= \dfrac{e^{\nu_{sm}}}{\sum_{s=1}^S e^{\nu_{sm}}} \\
@@
 
-\section*{Model convergance}
-Convergence and reliability of the Bayesian model were assessed through comprehensive diagnostics.
-All parameters (Table S.\ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (Fig.
-S.\ref{tab:diagnostics}).
-An exception was the parameter $\psi$ (mean hake), which showed marginally lower mixing ESS$_\psi$ = 149.7) suggesting mild autocorrelation but no critical failure of convergence.
+\subsection*{Model convergance}
 
-No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (Fig.
-S.\ref{tab:diagnostics}).
-The posterior likelihood demonstrated convergence before the sampling phase began, with all chains exhibiting high mixing, confirming robust exploration of the parameter space (Fig.
-S.\ref{tab:diagnostics}).
-
-Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (Fig.
-S.\ref{tab:prior_sensitivity}).
-Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (Fig.
-S.\ref{tab:PPC}).
-
-Collectively, these diagnostics confirm the reliability and validity of the Bayesian model used here.
+Gelman-Rubin statistics.  Predicted-observed, etc.
 
@@
 \caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
-\label{tab:PPC}
 \end{figure}
@@
 \caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.88\textwidth]{plots/11_Diagnostic_Fig_1.jpg}
-\caption{Bayesian model diagnostic plots including $\hat{R}$ and ESS for each parameter (A) and posterior likelihood, acceptance rate, tree depth, divergence transition, and step size for each iteration in each chain (including warm up phase; B).}
-\label{tab:diagnostics}
-\end{figure}
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=0.98\textwidth]{plots/12_Diagnostic_Fig_2.jpg}
-\caption{Prior sensitivity plots indicating the distribution of priors in red and posteriors in black with dots indicating the mean and ticks showing 95\% credible interval.}
-\label{tab:prior_sensitivity}
 \end{figure}
@@
     \end{tabular}
-    \label{tab:mock_table}
 \end{table}
 \clearpage
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{lll}
-         & \textbf{Description} & \textbf{Prior} \\
-&Data & \\
-\hline
-$K$ & known DNA concentration from qPCR stsandards in ln(copies/$\mu$L) & - \\
-$Z$ & qPCR amplification (yes=1; no=0) & - \\
-$Y$ & qPCR cycle threshold (Ct) & - \\
-$R$ & metabarcoding number of reads & - \\
-$T$ & total number of reads for all S species & - \\
-$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
-$L$ & known DNA concentration from mock commuinties in ln(copies/$\mu$L) & - \\
-$V$ & volume filtered and aliquote dilution offset & -\\
-$I$ & aliquote dilution factor & -\\
-
-&&\\
-&State processes&\\
-\hline
-$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
-$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
-$F$ & smoothed DNA concentration in copies/L & - \\
-&&\\
-
-&Parameters&\\
-\hline
-$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
-$\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\tau)$\\
-$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
-
-$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\
-
-$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
-$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
-&and eDNA concentration (K and E) & \\
-$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
-&eDNA concentration (K and E) & \\
-$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
-&of Ct values and eDNA concentration (K and E) & \\
-$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
-&of Ct values and eDNA concentration (K and E) & \\
-$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (Table S.\ref{tab:alphas})\\
-$\zeta$& ethanol wash effect & fixed (-1.15)\\
-&&\\
-&Index&\\
-\hline
-$s$& species ($H$ = Hake) & -\\
-$j$& location (northing(x), easting(y), depth(d)))& -\\
-$i$& biological replicate & -\\
-$r$& technical replicate & -\\
-$p$& qPCR plate & -\\
-$m$& mock community sample & -\\
-$q$& qPCR standard aliquote sample & -\\
-$R$& metabarcoding reference species & -\\
-
-    \end{tabular}
-    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
-    \label{tab:priortable}
-\end{table}
-
-\clearpage
-\begin{table}\centering
-\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} here).}
-    \begin{tabular}{lc}
-\textbf{Species} & $\alpha$ \\
-\textit{Clupea pallasii} &   -0.00138 \\
-\textit{Engraulis mordax} &    0.00557 \\
-\textit{Leuroglossus stilbius} &   -0.00390 \\
-\textit{Microstomus pacificus} &    0.00194 \\
-\textit{Sardinops sagax} &    0.00705 \\
-\textit{Scomber japonicus} &    0.00434 \\
-\textit{Sebastes entomelas} &    0.00459 \\
-\textit{Stenobrachius leucopsarus} &    0.00970 \\
-\textit{Tactostoma macropus} &   -0.00091 \\
-\textit{Tarletonbeania crenularis} &    0.00144 \\
-\textit{Thaleichthys pacificus} &   -0.00506 \\
-\textit{Trachurus symmetricus} &   -0.00875 \\
-\textit{Merluccius productus} &    0.00000 \\
-    \end{tabular}
-        \label{tab:alphas}
-\end{table}
-\clearpage
-
-\bibliography{clean_bib.bib}
+\bibliography{Bib.bib}
 \end{document}

### 2025-02-27 Appendix_plain_v3.tex
Updates from Overleaf

--- 2025-02-27 Appendix_plain_v2.tex
@@
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
-Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water.
+Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see SI Appendix Table S.\ref{tab:mock_table} for DNA quantification results).
 The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read.
@@
 
+Convergence and reliability of the Bayesian model were assessed through comprehensive diagnostics.
+All parameters (SI Appendix Table S.\ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (SI Appendix Fig.
+S.\ref{tab:diagnostics}\textit{A}).
+An exception was the parameter $\psi$ (mean hake), which showed marginally lower mixing ESS$_\psi$ = 149.7) suggesting mild autocorrelation but no critical failure of convergence.
+
+No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (SI Appendix Fig.
+S.\ref{tab:diagnostics}\textit{B}).
+The posterior likelihood demonstrated convergence before the sampling phase began, with all chains exhibiting high mixing, confirming robust exploration of the parameter space (Fig.
+S.\ref{tab:diagnostics}\textit{B}).
+
+Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (SI Appendix Fig.
+S.\ref{tab:prior_sensitivity}).
+Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (SI Appendix Fig.
+S.\ref{tab:PPC}).
+
+Collectively, these diagnostics confirm the reliability and validity of the Bayesian model used here.
+
 %%% Each figure should be on its own page
@@
 \centering
-\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
-\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
+\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
+\caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
-\caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
+\includegraphics[width=0.88\textwidth]{plots/11_Diagnostic_Fig_1.jpg}
+\caption{Bayesian model diagnostic plots including $\hat{R}$ and ESS for each parameter (A) and posterior likelihood, acceptance rate, tree depth, divergence transition, and step size for each iteration in each chain (including warm up phase; B).}
+\label{tab:diagnostics}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/12_Diagnostic_Fig_2.jpg}
+\caption{Prior sensitivity plots indicating the distribution of priors in red and posteriors in black with dots indicating the mean and ticks showing 95\% credible interval.}
+\label{tab:prior_sensitivity}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
+\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
+\label{tab:PPC}
 \end{figure}
@@
     \end{tabular}
+    \label{tab:mock_table}
 \end{table}
@@
 \hline
-$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
+$K$ & known DNA concentration from qPCR stsandards in ln(copies/$\mu$L) & - \\
 $Z$ & qPCR amplification (yes=1; no=0) & - \\
@@
 \hline
-$E$ & unknown DNA concentration in copies/$\mu$L & - \\
-$D$ & unknown DNA concentration in copies/L & - \\
+$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
+$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
 $F$ & smoothed DNA concentration in copies/L & - \\
@@
 \hline
-$\psi$& mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
-$\epsilon_{(x,y,d)}$& depth specific random effect & $\mathcal{N}(0,\tau)$\\
-$\tau$& standard deviation of depth specific random effect & $\mathcal{N}(0,0.2)$\\
+$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\tau)$\\
+$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
+
 $\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D$& standard deviation of biological replicate random effect  & $\mathcal{N}(0,3)$\\
+$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\
+
 $\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
-$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
+$\beta_0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
 &and eDNA concentration (K and E) & \\
-$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
+$\beta_1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
 &eDNA concentration (K and E) & \\
-$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
+$\gamma_0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
 &of Ct values and eDNA concentration (K and E) & \\
-$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
+$\gamma_1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
 &of Ct values and eDNA concentration (K and E) & \\
-$\alpha$& species amplification efficiency & fixed\\
-$\zeta$& ethanol wash effect & fixed\\
+$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (SI Appendix Table S.\ref{tab:alphas})\\
+$\zeta$& ethanol wash effect & fixed (-1.15)\\
 &&\\
-&Subscripts&\\
+&Index&\\
 \hline
 $s$& species ($H$ = Hake) & -\\
-$x$& northing & -\\
-$y$& easting & -\\
-$d$& depth & -\\
+$j$& location (northing(x), easting(y), depth(d)))& -\\
 $i$& biological replicate & -\\
@@
 $m$& mock community sample & -\\
-$j$& qPCR standard aliquote sample & -\\
+$q$& qPCR standard aliquote sample & -\\
+$R$& metabarcoding reference species & -\\
 
@@
     \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
-    \label{tab:filter_error}
+    \label{tab:priortable}
+\end{table}
+
+\clearpage
+\begin{table}\centering
+\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} here).}
+    \begin{tabular}{lc}
+\textbf{Species} & $\alpha$ \\
+\textit{Clupea pallasii} &   -0.00138 \\
+\textit{Engraulis mordax} &    0.00557 \\
+\textit{Leuroglossus stilbius} &   -0.00390 \\
+\textit{Microstomus pacificus} &    0.00194 \\
+\textit{Sardinops sagax} &    0.00705 \\
+\textit{Scomber japonicus} &    0.00434 \\
+\textit{Sebastes entomelas} &    0.00459 \\
+\textit{Stenobrachius leucopsarus} &    0.00970 \\
+\textit{Tactostoma macropus} &   -0.00091 \\
+\textit{Tarletonbeania crenularis} &    0.00144 \\
+\textit{Thaleichthys pacificus} &   -0.00506 \\
+\textit{Trachurus symmetricus} &   -0.00875 \\
+\textit{Merluccius productus} &    0.00000 \\
+    \end{tabular}
+        \label{tab:alphas}
 \end{table}
@@
 
-\bibliography{Bib.bib}
+\bibliography{clean_bib.bib}
 \end{document}

### 2025-02-27 Manuscript_plain_v2.tex
Updates from Overleaf

--- 2025-02-27 Manuscript_plain_AOS.tex
@@
 Notably, our analyses identified biodiversity hotspots that align with previously documented regions of ecological significance, such as the Columbia River plume and the Heceta Bank.
+
+\section*{Acknowledgment}
+We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
+This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
+The authors gratefully acknowledge the survey expertise and support of NOAA NWFSC's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
+Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
 

### 2025-02-27 Appendix_plain_v3.tex
Updates from Overleaf

@@
 We developed a joint model for modeling DNA concentration in the coastal ocean using single-species qPCR data Pacific hake and metabarcoding data from the 12S MiFish Universal primer.
-This work builds off previous single-species eDNA models constructed for Pacific hake (Shelton et al.
-2022) and models for joining qPCR and metabarcoding data (Guri et al.
-2024).
+This work builds off previous single-species eDNA models constructed for Pacific hake (\cite{shelton2022}) and models for joining qPCR and metabarcoding data (\cite{guri2024a}).
 Here qPCR observations provide information about DNA concentration for one species (Pacific hake) and the metabarcoding data provide information about species composition within a sample (i.e.
@@
 For water sample $i$: 1) $V_i$ is the proportion of 2.5 L filtered from Niskin $i$ (occasionally some seawater was spilled or not filtered; $V_i =1$ for the vast majority of samples); 2) $I_i$ is the known dilution used to on sample $i$ to eliminate PCR inhibition.
-PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see Shelton et al.
-2022 Supp S1 for additional description).
-Here we fix $\zeta$ = \textbf{XXX} based on previous analyses (Shelton et al.
-2022).
+PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see \cite{shelton2022} Supp S1 for additional description).
+Here we fix $\zeta$ = \textbf{-1.15} based on previous analyses (\cite{shelton2022}).
 
@@
 Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
-Here, we follow and extend the metabarcoding model developed by Shelton et al.
-(2023) and show how it links with the qPCR model above.
+Here, we follow and extend the metabarcoding model developed by \cite{shelton2023} and show how it links with the qPCR model above.
 
@@
 The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model, exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
-Because of the predominance of very low estimated DNA concentrations at a large number of site, we set $G_{s}(x,y,d) < 0.5$ equal to zero and then used a Tweedie likelihood in sdmTMB,
+To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model (SI Appendix Figure S.\ref{fig:DAG_v3}), exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
+Because of the predominance of very low estimated DNA concentrations at a large number of site, we set $G_{s}(x,y,d) < 0.5$ equal to zero (SI Appendix Figure S.\ref{fig:presence_threshold}) and then used a Tweedie likelihood in sdmTMB,
 
 \begin{align} 
-	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \varepsilon_{d}(x,y,d) \\
+	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \psi_{d}(x,y,d) \\
 	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
@@
 $\kappa$ is scalar intercept.
-Both $\eta(x,y)$ and $\varepsilon_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\varepsilon_d \sim  \mathrm{MVN}(0,\sum_{\varepsilon_d})$ (see \cite{anderson2022}).
-$\eta$ is a depth-invariant spatial field, while $\varepsilon_d$ is a spatial field specific to each depth category.
-In simple terms, each station-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific.
-$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\varepsilon_d$ modulates that estimate for different sampling depths.
-The $\varepsilon_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
+Both $\eta(x,y)$ and $\psi_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\psi_d \sim  \mathrm{MVN}(0,\sum_{\psi_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\psi_d$ is a spatial field specific to each depth category.
+In simple terms, each location-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific.
+$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\psi_d$ modulates that estimate for different sampling depths.
+The $\psi_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
 
 Estimation of large, multi-level spatiotemporal models can strain computational power, in part because of the need in spatial models to invert large covariance matrices between the data.
-We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (Lindgren et al.
-2024; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh (Lindgren et al.
-2011).
-The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\varepsilon_d$.
+We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (\cite{lindgren2024}; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh (\cite{lindgren2011}).
+The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\psi_d$.
 Model estimates and the degree of spatial smoothing can be affected by the details of mesh construction, and we therefore produced a range of meshes with different complexities (i.e., varying numbers of triangles, triangle sizes, and therefore vertices) to test for each species.
@@
 We tested 18 candidate meshes for each species, increasing in complexity from 33 to 205 vertices, as well as allowing each model to include or exclude the scalar intercept $\kappa$.
-Each candidate model for each species was fit by maximum marginal likelihood, implemented in the sdmTMB package in R ( \cite{anderson2022} ).
+Each candidate model for each species was fit by maximum marginal likelihood, implemented in the sdmTMB package in R (\cite{anderson2022}).
 Model selection was carried out by first ensuring proper convergence, then choosing the model for each species that maximized the summed likelihood.
 
-We present results based on the predicted smooth mean concentration, $F_s(x,y,d)$, in all results.
+The results included in this article are based on the predicted smooth mean concentration, $F_s(x,y,d)$.
 
@@
 \centering
-\includegraphics[width=16cm]{plots/DAG-Ole.pdf} 
+\includegraphics[width=16cm]{plots/DAG_v3.pdf} 
 \caption{Schematic for the statistical models to produce species distributions.
@@
 Notation for parameters, subscripts, and observations are indicated in gray.}
-\label{fig:DAG2}
+\label{fig:DAG_v3}
 \end{figure*}
@@
 The dashed red line indicates the threshold model estimated concentration (copies/L) for defining presence and absence of sequenced (through metabarcoding) taxa}
+\label{fig:presence_threshold}
 \end{figure}

### 2025-02-27 DAG_v3.tex
Updates from Overleaf

--- 2025-02-24 DAG-Ole.tex
@@
 \documentclass[landscape]{article}
-\usepackage[paperwidth=250mm, paperheight=200mm, margin=0cm]{geometry}
+\usepackage[paperwidth=252mm, paperheight=170mm, margin=0cm]{geometry}
 \usepackage{tikz}
@@
        % \mu_{ik} = \beta_{0p} + \beta_{1p}  K_{i} \\
-        Z_{qpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{q}} \phi) \right) \\
-        Y_{qpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{q} , \sigma_{Yq}) \quad \textnormal{if } Z_{qpr} = 1 \\
-        \sigma_{Yq} = \exp(\gamma_0 + \gamma_1  K_{q}) \\
+        Z_{qpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{qp}} \phi) \right) \\
+        Y_{qpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{qp} , \sigma_{Yqp}) \quad \mathrm{if}\quad Z_{qpr} = 1 \\
+        \sigma_{Yqp} = \exp(\gamma_0 + \gamma_1  K_{qp}) \\
     \end{array}$
@@
        % \mu_{ip} = \beta_{0p} + \beta_{1p} E_{Hi} \\
-         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hji}} \phi) \right) \\
-         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hji} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1 \\
-         \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})
+         Z_{Hjipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hji}} \phi) \right) \\
+         Y_{Hjipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hji}, \sigma_{Yjip}) \quad \mathrm{if} \quad Z_{Hjipr} = 1 \\
+         \sigma_{Yjip} = \exp(\gamma_0 + \gamma_1 E_{Hji})
     \end{array}$
@@
     $\begin{array}{c}
-        \nu_{si} = (E_{sji} - E_{Rji} ) + N_{PCR}  \alpha_s  \\  [8pt]
-        \pi_{si} = \frac{exp({\nu_{si}})}{\sum_{s=1}^{S} exp({\nu_{si}})} \\ [8pt]
-         \mathbf{R}_{i} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{i}, T_{i} \right) 
+        \nu_{sji} = (E_{sji} - E_{R_iji} ) + N_{PCR} \left(\alpha_s - \alpha_{R_i} \right)  \\  [8pt]
+        \pi_{sji} = \frac{exp({\nu_{sji}})}{\sum_{s=1}^{S} exp({\nu_{sji}})} \\ [8pt]
+         \mathbf{W}_{ji} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{ji}, T_{i} \right) 
     \end{array}$
@@
     $\begin{array}{c}
-       \nu_{sm} = (L_{sm} - L_{Rm}) + N_{PCR}  \alpha_s     \\   [8pt]
+       \nu_{sm} = (L_{sm} - L_{Rm}) + N_{PCR} \left(\alpha_s - \alpha_{R} \right)     \\   [8pt]
        \pi_{sm} = \frac{exp({\nu_{sm}})}{\sum_{s=1}^{S} exp({\nu_{sm}})} \\ [8pt]
-       \mathbf{R}_{m} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{m}, T_{m} \right) 
+       \mathbf{W}_{m} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{m}, T_{m} \right) 
     \end{array}$
@@
 \node (parameters) [rectangle, draw, align=left, below=of metabarcoding, text width=9cm, fill=myGrey] at (3.5,-6){
-    \textbf{Parameters} \\
+    \textbf{Parameters and derived quantities} \\
     $\psi$ = mean hake eDNA concentration \\
@@
     $\beta_{0p}, \beta_{1p}$ = intercept and slope defining mean qPCR - DNA concentration relationship\\
-    $\gamma0, \gamma1$ = intercept and slope defining  standard deviation of qPCR - DNA concentration relationship \\
+    $\gamma_0, \gamma_1$ = intercept and slope defining  standard deviation of qPCR - DNA concentration relationship \\
     $\alpha_s$ = amplification efficiency relative to reference species \\
     $\pi$ = predicted sequence proportions at the end of PCR \\
-    $\nu$ = sequence proportions at the end of PCR (additive log-ratio)\\
     $\zeta$ = ethanol wash effect \\
     $\tau$ = standard deviation among biological replicates\\
-    $\eta$ = spatial field \\
-    $\varepsilon$ =  depth specific spatial field
 };
@@
     $L$ = known DNA conc. (log(copies/$\mu$L))\\
-    $R$ = metabarcoding sequencing reads \\
+    $W$ = metabarcoding sequencing reads \\
     $T$ = total number sequencing reads \\
@@
 % Legend: Data
-\node (States of Interest) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-10){
+\node (States of Interest) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-9.25){
     \textbf{States } \\

### 2025-02-27 Appendix_plain_v3.tex
Updates from Overleaf

@@
 \begin{align} 
-	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \psi_{d}(x,y,d) \\
+	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \omega_{d}(x,y,d) \\
 	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
@@
 $\kappa$ is scalar intercept.
-Both $\eta(x,y)$ and $\psi_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\psi_d \sim  \mathrm{MVN}(0,\sum_{\psi_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\psi_d$ is a spatial field specific to each depth category.
+Both $\eta(x,y)$ and $\omega_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\omega_d \sim  \mathrm{MVN}(0,\sum_{\omega_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\omega_d$ is a spatial field specific to each depth category.
 In simple terms, each location-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific.
-$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\psi_d$ modulates that estimate for different sampling depths.
-The $\psi_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
+$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\omega_d$ modulates that estimate for different sampling depths.
+The $\omega_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
 
@@
 We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (\cite{lindgren2024}; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh (\cite{lindgren2011}).
-The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\psi_d$.
+The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\omega_d$.
 Model estimates and the degree of spatial smoothing can be affected by the details of mesh construction, and we therefore produced a range of meshes with different complexities (i.e., varying numbers of triangles, triangle sizes, and therefore vertices) to test for each species.

### 2025-02-27 Appendix_plain_v3.tex
Updates from Overleaf

@@
 The $\omega_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
+Additionally to convert eDNA concentration from from copies/$\mu$L (reaction concentration) to copies/L of water filtered (ecological concentration) we multiply $F_s(x,y,d)$ by 20 which is the conversion value indicating the ratio of filtered volume and the volume of DNA template used in the PCR reaction.
 

### 2025-02-27 DAG_v3.tex
Updates from Overleaf

@@
 \documentclass[landscape]{article}
-\usepackage[paperwidth=252mm, paperheight=170mm, margin=0cm]{geometry}
+\usepackage[paperwidth=252mm, paperheight=180mm, margin=0cm]{geometry}
 \usepackage{tikz}
@@
         % \end{array}$} (qPCR_en);
-        \draw [arrow] (state1) -- node[midway, left] {$$} (state2H);
+        \draw [arrow] (state1) -- node[midway, left] {} (state2H);
         \draw [arrow] (state2H) -- node[midway, right] {} (qPCR_en);
@@
         \draw [arrow] (mock) -- node[pos=0.7, above] {$\alpha_s$} (metabarcoding);
-        \draw [arrow] (state2H) -- node[midway, right] {$$} (metabarcoding);
-        \draw [arrow] (state2s) -- node[midway, right] {$$} (metabarcoding);
+        \draw [arrow] (state2H) -- node[midway, right] {} (metabarcoding);
+        \draw [arrow] (state2s) -- node[midway, right] {} (metabarcoding);
 
@@
     \textbf{Parameters and derived quantities} \\
-    $\psi$ = mean hake eDNA concentration \\
+    $\psi$ = mean hake eDNA concentration (log(copies/$\mu$L) \\
     %$\delta0, \delta1$ = offset from mean hake \\
@@
     $T$ = total number sequencing reads \\
+    $V$ = volume filtered and aliquote dilution offset \\
+    $I$ = aliquote dilution factor \\
 };
@@
 % Legend: Data
-\node (States of Interest) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-9.25){
+\node (States of Interest) [rectangle, draw, align=left, below=of subscripts, text width=8cm, fill=myGrey] at (12.5,-10.25){
     \textbf{States } \\

### 2025-02-27 Manuscript_plain_v2.tex
Updates from Overleaf

@@
 
+\section*{Acknowledgment}
+We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
+This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
+The authors gratefully acknowledge the survey expertise and support of NOAA NWFSC's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
+Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
+
 \section*{Abstract}
@@
 
-\section*{Acknowledgment}
-We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
-This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
-The authors gratefully acknowledge the survey expertise and support of NOAA NWFSC's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
-Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.
-
 \section*{Introduction}
-Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}.
+Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023}.
 However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2022a}).
 The use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
-there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2024}, and coastal oceans \cite{maes2023, shelton2019}).
+there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2023}, and coastal oceans \cite{maes2023, shelton2019}).
 However, uses of quantitative eDNA observations in management applications have significantly lagged behind their use in occurrence and biodiversity applications.
@@
 qPCR and ddPCR) because these approaches directly measure DNA concentration.
-In contrast, multispecies analysis (metabarcoding) reflects information on many species of interest simultaneously, but at the costs of producing compositional estimates of proportional abundance \cite{gloor2017a, mclaren2019} that may require statistical correction for species-specific biases due to PCR amplification \cite{shelton2023}.
+In contrast, multispecies analysis (metabarcoding) reflects information on many species of interest simultaneously, but at the costs of producing compositional estimates of proportional abundance \cite{gloor2017, mclaren2019} that may require statistical correction for species-specific biases due to PCR amplification \cite{shelton2023}.
 Consequently, metabarcoding data alone cannot yield information about the abundance of DNA from environmental samples (e.g.
@@
 The majority of eDNA shed into the environment dissipates from surface waters within hours \cite{andruszkiewiczallan2021}.
-Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most \cite{baetscher2024}.
+Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most \cite{baetscher2023}.
 Given that our sampling area spans more than 1,200 km north to south and tens of km east to west, the spatial patterns we report are likely to reflect genuine differences in fish communities rather than artifacts of eDNA transport dynamics \cite{guri2023}.
@@
 
-\bibliography{Bib.bib}
+\bibliography{clean_bib.bib}
 

### 2025-03-06 Appendix_plain_v3.tex
Updates from Overleaf
Update manuscript and main text for submission.

@@
 Unfortunately, qPCR does not directly measure DNA concentration.
-Instead qPCR measures the PCR cycle at which a 2$\mu L$ aliquot of a extacted DNA was detected to fluoresce and compares that observation against the flourescence pattern of samples of known DNA concentration ("the standard curve").
+Instead qPCR measures the PCR cycle at which a 2$\mu$L aliquot of a extacted DNA was detected to fluoresce and compares that observation against the flourescence pattern of samples of known DNA concentration ("the standard curve").
 This provides an estimate of DNA concentration for each unknown sample.
@@
 PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see \cite{shelton2022} Supp S1 for additional description).
-Here we fix $\zeta$ = \textbf{-1.15} based on previous analyses (\cite{shelton2022}).
+Here we fix $\zeta$ = -1.15 based on previous analyses (\cite{shelton2022}).
 
@@
 Each qPCR plate has replicate samples with a known number of DNA copies.
-These standards span six orders of magnitude (1 to 100,000 copies $\mu L^{-1}$, each of 2$\mu L$) and determine the relationship between copy number and PCR cycle of detection.
+These standards span six orders of magnitude (1 to 100,000 copies $\mu \mathrm{L}^{-1}$, each of 2$\mu \mathrm{L}$) and determine the relationship between copy number and PCR cycle of detection.
 Let $K_{qp}$ be the known log-copy number for sample $q$ in PCR plate $p$, then,
@@
 Importantly one of the $s$ species is Pacific hake and so connects to the qPCR model above.
-Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; \textbf{Aitchison 1986}).
+Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; \cite{aitchison1986}).
 We define a reference species in each sample, $R_i$, with the species used as a reference varying among sample.
@@
 \end{align}
-where x, y, and d index longitude, latitude, and depth, respectively.
+where $x$, $y$, and $d$ index longitude, latitude, and depth, respectively.
 $\kappa$ is scalar intercept.
@@
 
-The results included in this article are based on the predicted smooth mean concentration, $F_s(x,y,d)$.
+All the results included in this article are based on the predicted smooth mean concentration, $F_s(x,y,d)$.
 
@@
 $\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\tau)$\\
-$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
+$\tau$& standard deviation of biological replicate random effect & $\mathcal{H-N}(0,0.2;0, +\infty)$\\
 
 $\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
-$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\
+$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{H-N}(0,3;0, +\infty)$\\
 
@@
     \end{tabular}
-    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
+    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions.
+$N$ and $H-N$ designate a normal and half-normal distribution, respectively.}
     \label{tab:priortable}
@@
 \begin{table}\centering
-\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} here).}
+\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} (Pacific hake) here).}
     \begin{tabular}{lc}

### 2025-03-06 Manuscript_plain_v3.tex
Update manuscript and main text for submission.

--- 2025-02-27 Manuscript_plain_v2.tex
@@
 PB, RPK, and GG did the bioinformatics;\\
-GG, OL, RPK, and eAOS analyzed the data;\\
+GG, OL, RPK, and AOS analyzed the data;\\
 GG, OL, RPK, and AOS interpreted the results;\\
 All authors provided edits to the manuscript \\
-Please check that this is correct and add/edit as needed XXX
 
@@
 These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously.
-Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological research.
+Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological interest.
 However, unlike single-species techniques, metabarcoding by itself can reflect only the proportions of target-species DNA present, not their absolute quantities.
@@
 
-While trawling and other traditional sampling methods have been the fundamental tools for directly measuring fish abundance and biomass, they are often impractical for species inhabiting inaccessible habitats, occurring at low densities, or fragile species easily injured during capture.
+Surveys of marine environments play a fundamental role in supporting fisheries and conservation of protected species but are often difficult and expensive endeavors.
+While trawling, acoustics, or other traditional sampling methods have long been used to directly measure species abundance and biomass, they can be impractical for species inhabiting inaccessible habitats, occurring at low densities, or for fragile species easily injured during capture.
 By contrast, eDNA sampling captures information on a broader range of taxa, including those that are difficult or impossible to study with traditional gear.
@@
 \section*{Results}
+
+Previous eDNA work has tended to feature more qualitative multispecies results, or else quantitative analysis of single species at small spatial scales.
+Here, we create a rigorous quantitative framework and use hundreds of water samples to make ecological inferences at the scale of management.
 We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
+\ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S1}, Pacific Herring and Jack Mackerel).
+For other high-biomass species otherwise lacking well-described distributions, we provide the first species distribution models (Fig.
+\ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S1}).
+We find that the estimated eDNA concentrations are correlated among species with similar traits and habitat preferences (Fig.
+\ref{fig:fig2}).
+Emergent patterns of species richness and estimated eDNA aggregation correspond with known ecological drivers, evidence that eDNA traces can reliably capture ecologically meaningful patterns (Fig.
+\ref{fig:fig3}).
+Particularly for forage fish species -- which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} -- molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
+
+%We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
 \ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
@@
 S1}.
+In particular, our analyses present detailed, depth-specific information on abundance that is difficult to obtain with traditional sampling methods.
 
@@
 
-\subsection*{Mid-water}
+\subsection*{Midwater}
 Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024, laidig2007} .
@@
 
-\subsection*{Benthic}
+\subsection*{Deepwater}
 Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common, yet under studied, such as the dragonfish (\textit{Tactostoma macropus}).
@@
 We also identify aggregations of richness and total eDNA near high-productivity regions of ecological and commercial value.
-Regions like Heceta Bank host dense aggregations of forage fish and are critical for both fishery management and for broader marine food-web stability.
+
+%Regions like Heceta Bank host dense aggregations of forage fish and are critical for both fishery management and for broader marine food-web stability.
 
@@
 
-Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
+Moreover, integrating retrospective eDNA analyses a changing climate can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
 For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
@@
 
-We have shown an example of the substantial value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
-Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional acoustic or trawl-based surveys.
-Furthermore, eDNA survey provide data in units of eDNA concentration, not more familiar units of mass or abundance.
+We have shown an example of the potential value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
 Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
 We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
+
+%Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional acoustic or trawl-based surveys.
+Furthermore, eDNA surveys provide data in units of eDNA concentration, not more familiar units of mass or abundance.
 

### 2025-03-06 Manuscript_plain_v3.tex
add acknowledgment

@@
 We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
-This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
+This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719) and support from the National Marine Fisheries Service's Genomic Strategic Initiative.
 The authors gratefully acknowledge the survey expertise and support of NOAA NWFSC's Pacific hake survey team and the personnel of the NOAA Ship Bell M.

### 2025-03-06 older versions/Appendix_plain_AOS_ORL.tex
older versions folder with old .tex.

--- 2025-02-27 Appendix_plain_v1.tex
@@
 Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA $copies$ $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
-Hereafter we refer to a combination of spatial location and depth as a ``station".
-Let $\psi_0$ be the intercept representing the mean hake DNA concentration and $\epsilon(x,y,d)$ be a random effect that allows for the effect of spatial location and water depth.
+Hereafter we refer to a combination of spatial location and depth as a ``station" and index station using $j$ (thus,$D_{Hj} = D_H(x,y,d)$ .
+Let $\psi_0$ be the intercept representing the mean hake DNA concentration and $\epsilon_j$ be a random effect that allows for the effect of that station.
 We write these as additive on the log-scale,
@@
 \begin{align}
-  D_H(x,y,d) &= \psi_0 + \epsilon(x,y,d) \\
-  \epsilon(x,y,d) &\sim N(0,\sigma_D)
+  D_{Hj} &= \psi_0 + \epsilon_j \\
+  \epsilon_j &\sim N(0,\sigma_D)
 \end{align}
@@
 \begin{align} 
- \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) 
+ \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{Ri} \right) \\
   \pi_{sm} &= \dfrac{e^{\nu_{sm}}}{\sum_{s=1}^S e^{\nu_{sm}}} \\
@@
 
-Ole's questions:
-Smoothed model is based on E (a single water bottle), not D.
-What was used in the making the smooth?  Posterior mean?
+The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
+We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as: $$E_i(xyz) = \tau + \eta(xy) + \varepsilon_{d}(xyz)$$ where x,y,z index longitude, latitude, and depth, respectively.
+We used posterior means of $E_i$ from the joint statistical model above as the dependent variable.
+$\tau$ is scalar intercept.
+Both $\eta(xy)$ and $\varepsilon_d(xyz)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim MVN(O,\sum_\eta)$ and $\varepsilon_d \sim MVN(O,\sum_{\varepsilon_d})$ (see \cite{anderson2022}).
+$\eta$ is a depth-invariant spatial field, while $\varepsilon_d$ is a spatial field specific to each depth category.
+In simple terms, each station-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept (that can be zero) and two spatial fields, one that is common across depths and one that is depth-specific.
+$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated eDNA, while $\varepsilon_d$ modulates that estimate for different sampling depths.
+The $\varepsilon_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
 
-\textbf{OWEN.  COULD YOU HELP HERE?}
+Estimation of large, multi-level spatiotemporal models can strain computational power, in part because of the need in spatial models to invert large covariance matrices between the data.
+We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (Lindgren et al.
+2024; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh (Lindgren et al.
+2011).
+The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\varepsilon_d$.
+Model estimates and the degree of spatial smoothing can be affected by the details of mesh construction, and we therefore produced a range of meshes with different complexities (i.e., varying numbers of triangles, triangle sizes, and therefore vertices) to test for each species.
 
-The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as:
-$$E_i(xyz) = \tau + \eta(s) + \varepsilon_{d}(s)$$
-where x,y,z index longitude, latitude, and depth, respectively, $\tau$ is scalar intercept, $\eta(s)$ is a spatial smooth shared among depths, and $\varepsilon_d(s)$ is a depth-specific spatial smooth.
-Both $\eta(s)$ and $\varepsilon_d(s)$ are a zero-centered Gaussian random field estimated using the stochastic partial differntial equation (SPDE) approximation (see \cite{anderson2022}).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model.
+Each species was fitted individually using the equation above.
+We tested 18 candidate meshes for each species, increasing in complexity from 33 to 205 vertices, as well as allowing each model to include or exclude the scalar intercept $\tau$.
+Each candidate model for each species was fit by maximum marginal likelihood, implemented in the sdmTMB package in R ( \cite{anderson2022} ).
+Model selection was carried out by first ensuring proper convergence, then choosing the model for each species that maximized the summed likelihood.
 

### 2025-03-06 older versions/Manuscript_plain_v1.tex
older versions folder with old .tex.

--- 2025-02-12 Manuscript_plain.tex
@@
 We have shown an example of the substantial value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
-Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional trawl-based surveys.
+Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional acoustic or trawl-based surveys.
 
-Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
+A significant difference between  management decisions rely on units of mass or abundance, not in units of eDNA concentration.
+Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
+We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
+
+%Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
 In reality, fish can move in between sampling location within the study area, which may bias abundance estimates  \cite{rota2009}.
@@
 This cumulative, time-integrated signal allows eDNA surveys to circumvent the closure assumption and provide a more accurate representation of fish presence and activity.
-
-Perhaps most notably, management decisions happen in units of mass or abundance, not in units of eDNA concentration.
-Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that are likely to vary among species.
-We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
 
@@
 
-\bibliography{pnas-sample.bib}
+\bibliography{Bib.bib}
 

### 2025-03-06 Main_text.tex
Updates from Overleaf

--- 2025-02-11 Main_text.tex
@@
 PB, RPK, and GG did the bioinformatics;
-GG, OL, RPK, and AOS analyzed the data;
+GG, OL, RPK, and eAOS analyzed the data;
 GG, OL, RPK, and AOS interpreted the results;
 GG, OL, RPK, and AOS wrote the manuscript
-\hl{[Please check that this is correct and add/edit as needed XXX]}
 }
@@
 Furthermore, eDNA abundance patterns aligned with stock assessments in indicating a higher abundance of northern anchovy; however, the magnitude of difference varied, with eDNA suggesting a 2.5-fold difference, while stock assessments indicated a 35-fold difference in abundance.
-\hl{[CITE stock assessment here; OLE XXX]}
 
 \subsection*{Mid-water}
-Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} \hl{[ADD SCUBA REF; OLE XXX]}.
+Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024}.
 Moreover, because it is a common bycatch species in the hake (\textit{Merluccius productus}) fishery, limits on bycatch of widow rockfish can also limit the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
@@
 For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa, revealing unexpected habitats.
-Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \hl{[NWFSC trawl survey ref; OLE XXX]}.
+Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys.
 This species has been of special interest, as its targeted commercial harvest resumed in the late 2010s after a decade of closure (harvest was less than 500 mt annually between 2003 and 2013 but has exceeded 10,000 mt annually since 2018) and no existing survey adequately captures its patterns of abundance \cite{adams2019a}.
@@
 
-The majority of eDNA shed into the environment dissipates from surface waters within hours \hl{(Brasseale et al., in review)[add this reference; RYAN or OLE XXX]}.
+The majority of eDNA shed into the environment dissipates from surface waters within hours.
 Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most.
@@
 This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.
-\hl{[Check the framing of the last two sentences XXX]}
 
@@
 Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes.
-Here, we follow \cite{shelton2022} and \hl{Liu et al.
-(submitted)[add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
-Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision)[add this reference; RYAN or OLE XXX]}.
+Here, we follow \cite{shelton2022}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
+Elsewhere, the same set of samples has yielded information on the distribution of marine mammals.
 The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa.
@@
 \centering
-\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=18cm]{plots/DAG_v3.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
 \caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
@@
 
-\bibsplit[9]
+\bibsplit[25]
 %Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
 % Bibliography
-\bibliography{pnas-sample}
+\bibliography{clean_bib}
 

### 2025-03-06 SI_Appendix.tex
Updates from Overleaf

@@
 \documentclass[9pt,twoside,lineno]{pnas-SI}
-% Use the lineno option to display guide line numbers if required.
 
@@
 \title{Quantitative, Multispecies Monitoring at a Continental Scale}
-\author{Gledis Guri, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Ana Ram\'on-Laca,\\
-Krista M. Nichols, Pedro F. P. Brandão-Dias, Abigail Wells and Andrew Olaf Shelton}
+\author{Gledis Guri*, Owen Liu, Ryan P. Kelly, Megan Shaffer, Kim Parsons, Ana Ram\'on-Laca, Krista M. Nichols, Pedro F. P. Brandão-Dias, Abigail Wells and Andrew Olaf Shelton}
 \correspondingauthor{Gledis Guri.\\E-mail: gguri@uw.edu}
@@
 
-%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
 \SItext
@@
 We then quantified the concentration of the 12S rRNA gene using MarVer1 primers (forward: CGTGCCAGCCACCGCG; reverse: GGGTATCTAATCCYAGTTTG \cite{valsecchi2020}), which perfectly matched the template to give unbiased estimates of concentration, using ddPCR (Bio-Rad, Inc., QX200 Droplet Digital PCR system).
-Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water.
+Each tissue of each species was quantified in a 22 $\mu$L reaction consisted of 2 $\mu$L of DNA template from genomic DNA , 11 $\mu$L of ddPCR EvaGreen (Bio-Rad), 0.22 $\mu$L of each forward and reverse primers (10 uM), and 0.56 $\mu$L of nuclease free water (see SI Appendix Table \ref{tab:mock_table} for DNA quantification results).
 The thermocycler reactions were run in C1000 Touch Thermal Cycler with 96-Deep Well Reaction Module (Bio-Rad) using the PCR program as follows: 2 min at 50°C for enzyme activation, 2 min at 95°C for initial denaturation, and 40 cycles of denaturation for 1 sec at 95°C and primer annealing and elongation for 30 sec min at 60°C, with a ramp rate of 2°C per s and heldat 4°C until droplets were read.
@@
 
-\subsection*{Bayesian model}
-OLE
+\subsection*{Statistical model}
 
-The above statistical model provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA , as:
-$$\ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)$$
-where x,y,z index longitude, latitude, and depth, respectively, $\tau$ is scalar intercept, $\eta(s)$ is a spatial smooth shared among depths, and $\varepsilon_d(s)$ is a depth-specific spatial smooth.
-Both $\eta(s)$ and $\varepsilon_d(s)$ are a zero-centered Gaussian random field estimated using the stochastic partial differntial equation (SPDE) approximation (see \cite{anderson2022}).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model.
+\subsubsection*{Overview}
+We developed a joint model for modeling DNA concentration in the coastal ocean using single-species qPCR data Pacific hake and metabarcoding data from the 12S MiFish Universal primer.
+This work builds off previous single-species eDNA models constructed for Pacific hake \cite{shelton2022} and models for joining qPCR and metabarcoding data \cite{guri2024a}.
+Here qPCR observations provide information about DNA concentration for one species (Pacific hake) and the metabarcoding data provide information about species composition within a sample (i.e.
+proportions that sum to one).
+In combination, these two data sources allow for estimation of DNA concentrations for all species observed by metabarcoding.
+\subsubsection*{qPCR model}
+
+We developed a state-space framework for modeling hake DNA concentration in the coastal ocean building and modifying the work of \cite{shelton2022}.
+State-space models separate the true biological process from the methods used to observe the process.
+Let $D_H(x,y,d)$ be the true, but unobserved concentration of hake DNA (log DNA copies $\mu L^{-1}$) present at spatial coordinates $\{x,y\}$ (eastings and northings, respectively, in km) and sample depth $d$ (meters).
+Hereafter we refer to a combination of spatial location and depth as a ``station" and index station using $j$ (thus, $D_{Hj} = D_H(x,y,d)$).
+Let $\psi$ be the intercept representing the mean hake DNA concentration and $\epsilon_j$ be a random effect that allows for the effect of that station.
+We write these as additive on the log-scale,
+
+\begin{align}
+  D_{Hj} &= \psi + \epsilon_j \\
+  \epsilon_j &\sim  \mathrm{Normal}(0,\tau)
+\end{align}
+This model describes perhaps the simplest possible random effect model and notably does not leverage any information about the spatial arrangement of samples (i.e.
+the random effect does not include a spatial correlation component).
+This is intentional.
+The focus of this model is on the other species DNA concentration (not Pacific hake) and therefore we allow the estimated hake concentrations at a to follow the qPCR observations as much as possible.
+
+Unfortunately, qPCR does not directly measure DNA concentration.
+Instead qPCR measures the PCR cycle at which a 2$\mu$L aliquot of a extacted DNA was detected to fluoresce and compares that observation against the flourescence pattern of samples of known DNA concentration ("the standard curve").
+This provides an estimate of DNA concentration for each unknown sample.
+Each water sample is analyzed using at least 3 independent qPCR replicates to account for laboratory and machine variability.
+In addition to qPCR variability among replicates, we also have to account for modifications that affect each water sample and occurred during water sampling and processing.
+First, we have three offsets that modify the true DNA concentration to affect what we observed in the qPCR.
+For water sample $i$: 1) $V_i$ is the proportion of 2.5 L filtered from Niskin $i$ (occasionally some seawater was spilled or not filtered; $V_i =1$ for the vast majority of samples); 2) $I_i$ is the known dilution used to on sample $i$ to eliminate PCR inhibition.
+PCR inhibition is most commonly observed in surface samples and is vanishingly rare for samples collected below 100m; 3) $\mathbf{I}\zeta$ is an estimated offset for an ethanol wash error ($\zeta$ is the estimated effect of the wash error and $\mathbf{I}$ is an indicator variable where $\mathbf{I}=1$ for affected samples and $\mathbf{I}=0$ otherwise; see \cite{shelton2022} Supp S1 for additional description).
+Here we fix $\zeta$ = -1.15 based on previous analyses \cite{shelton2022}.
+
+Finally, we add a random effect for the individual bottles sampled at each station ($\delta_{ij}$).
+This effect  describes the deviation of individual sampled bottles from the location mean, $\delta_{ji} \sim  \mathrm{Normal}(0,\sigma_D)$, with $\sigma_D$ indicating a  standard deviation among bottles.
+Because we have only two replicate bottles at each location, we impose a sum to zero constraint on the bottles collected from a single location (i.e.
+for a given location, $\sum_i \delta_{ji} = 0$).
+Then, the log-concentration of hake DNA in a sample analyzed by qPCR, $E_{Hji}$ is,
+
+\begin{align}
+  E_{Hji} &= D_{Hj} +  \log{V_i} + \log{I_i} + \mathbf{I}\zeta + \delta_{ji}
+\end{align}
+We connect the estimated concentrations to observations from qPCR using two likelihoods connected to the concentration $E$.
+First we determine whether amplification was detected in each qPCR replicate $r$ run on each plate $p$, $Z_{Hjipr}$ and if amplification was observed we model the PCR cycle at which amplification was detected $Y_{Hjipr}$; $Y_{Hjipr}$ is observed as a continuous, positive value.
+DNA concentrations have a log-linear relationships with $Y_{Hjipr}$; smaller values of $Y_{Hjipr}$ are associated with higher DNA concentrations.
+
+Because we are modeling the discrete number of DNA molecules that are present in the assayed sample, we know that the Poisson distribution provides an appropriate observation distribution for the number of molecules present in a given qPCR reaction.
+Assuming a Poisson distribution, conditional on a true mean number of DNA copies in the sample $X$, the probability of having exactly zero DNA copies in a qPCR reaction is $e^{-X}$ and the probability of having non-zero DNA copies is the complement, $1-e^{-X}$.
+If there are exactly zero molecules in a qPCR reaction, amplification and therefore detection of amplification will not occur.
+However, there are other factors (e.g.
+PCR inhibition) that may reduce amplification further and therefore we expect the probability of amplification to be at most $1-e^{-X}$.
+As a result, we estimate an additional term, $\phi$, representing the fractional reduction in amplification efficiency due to other factors during qPCR ($0<\phi<1$).
+Then we have a pair of observation models for hake DNA, both of which are a function of DNA concentration, $E_{Hji}$,
+
+\begin{align}
+  Z_{Hjipr} &\sim  \mathrm{Bernoulli} \left(1 - \exp(-2 e^{E_{Hji}}\phi) \right) \\ 
+  Y_{Hjipr} &\sim  \mathrm{Normal} \left(\beta_{0p}+\beta_{1p}{E_{Hji}},\sigma_{Yji} \right) \qquad   \mathrm{if} \: Z_{Hjipr} = 1 
+\end{align} 
+The 2 is present in the first line of the equation because we use 2 $\mu$L of sample in each qPCR reaction and so the expected DNA copies in reaction is $2e^{E_{Hji}}$.
+We allow $\sigma_{Yji}$ to vary as a log-linear function of DNA concentration to account for the fact that there is decreased variability in qPCR measures of $C_t$ at higher DNA concentrations: $\sigma_{Yji} = exp(\gamma_{0}+\gamma_{1}E_{Hji})$.
+
+By itself, the above model is unidentifiable because field samples do not provide information about the parameters that define relationship between the number of DNA copies and PCR cycle ($\beta_{0j}$, $\beta_{1j}$, and $\phi$).
+As is standard with qPCR analyses, we include standards of known concentration to estimate these parameters.
+Each qPCR plate has replicate samples with a known number of DNA copies.
+These standards span six orders of magnitude (1 to 100,000 copies $\mu \mathrm{L}^{-1}$, each of 2$\mu \mathrm{L}$) and determine the relationship between copy number and PCR cycle of detection.
+Let $K_{qp}$ be the known log-copy number for sample $q$ in PCR plate $p$, then,
+
+\begin{align}
+  Z_{qpr} &\sim  \mathrm{Bernoulli} \left(1-exp(-2 e^{K_{qp}} \phi) \right)\\
+  Y_{qpr} &\sim  \mathrm{Normal} \left(\beta_{0j}+\beta_{1j}{K_{qp}},\sigma_{Yqp}\right) \quad  \mathrm{if} \: Z_{qjp} =1
+\end{align}
+Note that there are different intercept ($\beta_{0p}$) and slope ($\beta_{1p}$) parameters for each PCR plate to allow for among-plate variation in amplification.
+We model each calibration parameter ($\beta_{0p},\beta_{1p})$ hierarchically using a normal distribution, with among plate mean and variance (i.e.
+$\beta_{0p} \sim \mathrm{Normal}(\mu_{\beta_{0}},\sigma_{\beta_{0}})$).
+
+\subsubsection*{Metabarcoding model}
+
+In addition to the qPCR observations, we used the MiFish primer and high throughput sequencing to detect and enumerate the sequences of other fish species in the same samples that were examined for hake DNA.
+Again, we are interested in determining the DNA concentration of each focal species in each sample.
+Unfortunately, metabarcoding data is compositional data \cite{shelton2023}, and so while it can inform the relative abundance of species within a sample, alone it cannot provide estimates of DNA concentration.
+Fortunately, the qPCR data from Pacific hake provides information about DNA concentration for one species and the metabarcoding data provides information about the relative abundance for tens of species, including hake.
+Here, we follow and extend the metabarcoding model developed by \cite{shelton2023} and show how it links with the qPCR model above.
+
+As above, we let $D_{sj}$ be the log-scale eDNA concentration but now for species $s$.
+We let $E_{sji}$ be the log-DNA concentration for each species in each bottle.
+Importantly one of the $s$ species is Pacific hake and so connects to the qPCR model above.
+Metabarcoding reads are compositional and so we model the DNA concentration relative to one another using the additive log-ratio transformation (ALR; \cite{aitchison1986}).
+We define a reference species in each sample, $R_i$, with the species used as a reference varying among sample.
+Let $\alpha_s$ be the per PCR cycle log amplification rate for species $s$ (see \cite{shelton2023}).
+We can then calculate $\nu_{sji}$, the log-ratio between species $s$ and the reference species after accounting for amplification bias (see \textit{Mock Communities} below),
+
+\begin{align} 
+  \nu_{sji} &= \left( E_{sji} - E_{Rji}\right) + N_{PCR} \left(\alpha_s - \alpha_{R_i} \right) 
+\end{align}
+We can then convert the $\nu$s to the predicted proportions of the metabarcoding reads $\pi_{sji}$ using the softmax transformation and to the observed reads via a multinomial likelihood,
+
+\begin{align} 
+  \pi_{sji} &= \dfrac{e^{\nu_{sji}}}{\sum_{s=1}^S e^{\nu_{sji}}} \\
+  \boldsymbol{W}_{ji} & \sim  \mathrm{Multinomial}\left(\boldsymbol{\pi}_{ji},T_{i}\right) 
+\end{align}
+where $T_{i}$ is the total number of reads observed across all species for that bottle.
+
+As with the qPCR analyses, in the absence of additional information estimates of $\alpha$ and $E$ cannot be statistically identified.
+We use the mock community data, which has known input DNA concentrations and  same basic model as for the metabarcoding observations but instead of having unknown DNA concentrations, we have known inputs for each species $L_{sm}$ in mock community $m$.
+
+\begin{align} 
+ \nu_{sm} &= \left( L_{sm} - L_{Rm}\right) + N_{PCR} \left(\alpha_s - \alpha_{R_i} \right) \\
+  \pi_{sm} &= \dfrac{e^{\nu_{sm}}}{\sum_{s=1}^S e^{\nu_{sm}}} \\
+  \boldsymbol{W}_{m} & \sim  \mathrm{Multinomial}\left(\boldsymbol{\pi}_{m},T_{m}\right) 
+\end{align}
+To maintain identifiability, we assign one $\alpha_s = 0$.
+
+\subsubsection*{Smoothing predictions}
+
+The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
+To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model (SI Appendix Figure \ref{fig:DAG_v3}), exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
+Because of the predominance of very low estimated DNA concentrations at a large number of site, we set $G_{s}(x,y,d) < 0.5$ equal to zero (SI Appendix Figure \ref{fig:presence_threshold}) and then used a Tweedie likelihood in sdmTMB,
+
+\begin{align} 
+	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \omega_{d}(x,y,d) \\
+	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
+\end{align}
+where $x$, $y$, and $d$ index longitude, latitude, and depth, respectively.
+$\kappa$ is scalar intercept.
+Both $\eta(x,y)$ and $\omega_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\omega_d \sim  \mathrm{MVN}(0,\sum_{\omega_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\omega_d$ is a spatial field specific to each depth category.
+In simple terms, each location-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific.
+$\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\omega_d$ modulates that estimate for different sampling depths.
+The $\omega_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.
+Additionally to convert eDNA concentration from from copies/$\mu$L (reaction concentration) to copies/L of water filtered (ecological concentration) we multiply $F_s(x,y,d)$ by 20 which is the conversion value indicating the ratio of filtered volume and the volume of DNA template used in the PCR reaction.
+
+Estimation of large, multi-level spatiotemporal models can strain computational power, in part because of the need in spatial models to invert large covariance matrices between the data.
+We use the Stochastic Partial Differential Equation (SPDE) model, implemented through the fmesher R package (\cite{lindgren2024}; 10.32614/CRAN.package.fmesher) to approximate the latent spatial process and its variance by using a triangulated mesh \cite{lindgren2011}.
+The SPDE approach drastically reduces computation time by using a sparse covariance matrix to approximate Gaussian Markov random fields like $\eta$ and $\omega_d$.
+Model estimates and the degree of spatial smoothing can be affected by the details of mesh construction, and we therefore produced a range of meshes with different complexities (i.e., varying numbers of triangles, triangle sizes, and therefore vertices) to test for each species.
+
+Each species was fitted individually using the equation above.
+We tested 18 candidate meshes for each species, increasing in complexity from 33 to 205 vertices, as well as allowing each model to include or exclude the scalar intercept $\kappa$.
+Each candidate model for each species was fit by maximum marginal likelihood, implemented in the sdmTMB package in R \cite{anderson2022}.
+Model selection was carried out by first ensuring proper convergence, then choosing the model for each species that maximized the summed likelihood.
+
+All the results included in this article are based on the predicted smooth mean concentration, $F_s(x,y,d)$.
 
 \subsection*{Model convergance}
-OLE + GLED
+
+Convergence and reliability of the Bayesian model were assessed through comprehensive diagnostics.
+All parameters (SI Appendix Table \ref{tab:priortable}) exhibited Gelman-Rubin convergence statistics ($\hat{R} < 1.05$) and effective sample sizes (ESS) exceeding 200 per parameter, indicating successful convergence and efficient mixing of the six independent chains (SI Appendix Fig.
+\ref{tab:diagnostics}\textit{A}).
+An exception was the parameter $\psi$ (mean hake), which showed marginally lower mixing ESS$_\psi$ = 149.7) suggesting mild autocorrelation but no critical failure of convergence.
+
+No divergent transitions were detected during sampling, and the maximum tree depth was not exceeded, indicating no issues with divergence or exploration limits (SI Appendix Fig.
+\ref{tab:diagnostics}\textit{B}).
+The posterior likelihood demonstrated convergence before the sampling phase began, with all chains exhibiting high mixing, confirming robust exploration of the parameter space (Fig.
+\ref{tab:diagnostics}\textit{B}).
+
+Prior sensitivity analyses revealed that posterior estimates differed from priors, demonstrating that the posteriors were appropriately updated based on the observed data rather than being heavily influenced by prior assumptions (SI Appendix Fig.
+\ref{tab:prior_sensitivity}).
+Additionally, the posterior predictive checks (PPC) demonstrated that the model reliably reproduced the observed data, supporting the validity of parameter estimates (SI Appendix Fig.
+\ref{tab:PPC}).
+
+Collectively, these diagnostics confirm the reliability and validity of the Bayesian model used here.
 
@@
 %%% Each figure should be on its own page
-\begin{figure}
+\begin{figure*}[tbhp] 
 \centering
-\includegraphics[width=0.89\textwidth]{plots/4_Figure_4.jpg}
-\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
-Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A).
-Probability density functions (PDF) of estimated eDNA concentrations across latitudinal gradients (B; square root transformed of x-axis to enhance low concentration peaks).
-Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
-\end{figure}
+\includegraphics[width=16cm]{plots/DAG_v3.pdf} 
+\caption{Schematic for the statistical models to produce species distributions.
+Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
+Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
+Notation for parameters, subscripts, and observations are indicated in gray.}
+\label{fig:DAG_v3}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=0.89\textwidth]{plots/10_Supplementary_Figure_6.jpg}
-\caption{Estimated species eDNA concentration (copies/L) for all the species included in this study concentration across 0, 50, 150, 300, and 500m depth samples, without spatial smoothing.}
+\includegraphics[width=0.88\textwidth]{plots/11_Diagnostic_Fig_1.jpg}
+\caption{Bayesian model diagnostic plots including $\hat{R}$ and ESS for each parameter (A) and posterior likelihood, acceptance rate, tree depth, divergence transition, and step size for each iteration in each chain (including warm up phase; B).}
+\label{tab:diagnostics}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/12_Diagnostic_Fig_2.jpg}
+\caption{Prior sensitivity plots indicating the distribution of priors in red and posteriors in black with dots indicating the mean and ticks showing 95\% credible interval.}
+\label{tab:prior_sensitivity}
+\end{figure}
+
+%%% Each figure should be on its own page
+\begin{figure}
+\centering
+\includegraphics[width=0.98\textwidth]{plots/9_Supplementary_Figure_5.jpg}
+\caption{Mean posterior predictions for three model compartments, qPCR continuous model (A), qPCR occupancy model (B), mock community model (C and D on species proportions and number of reads) and metabarcoding model (E)}
+\label{tab:PPC}
 \end{figure}
@@
     \end{tabular}
+    \label{tab:mock_table}
 \end{table}
 
-\bibliography{pnas-sample}
+\begin{table}[h]
+    \centering
+    \begin{tabular}{lll}
+         & \textbf{Description} & \textbf{Prior} \\
+&Data & \\
+\hline
+$K$ & known DNA concentration from qPCR stsandards in ln(copies/$\mu$L) & - \\
+$Z$ & qPCR amplification (yes=1; no=0) & - \\
+$Y$ & qPCR cycle threshold (Ct) & - \\
+$R$ & metabarcoding number of reads & - \\
+$T$ & total number of reads for all S species & - \\
+$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
+$L$ & known DNA concentration from mock commuinties in ln(copies/$\mu$L) & - \\
+$V$ & volume filtered and aliquote dilution offset & -\\
+$I$ & aliquote dilution factor & -\\
+
+&&\\
+&State processes&\\
+\hline
+$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
+$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
+$F$ & smoothed DNA concentration in copies/L & - \\
+&&\\
+
+&Parameters&\\
+\hline
+$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
+$\epsilon_{j}$& spatial and depth specific random effect & $\mathcal{N}(0,\tau)$\\
+$\tau$& standard deviation of biological replicate random effect & $\mathcal{H-N}(0,0.2;0, +\infty)$\\
+
+$\delta$& biological replicate random effect & $\mathcal{N}(0,\sigma_D)$\\
+$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{H-N}(0,3;0, +\infty)$\\
+
+$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
+$\beta_0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
+&and eDNA concentration (K and E) & \\
+$\beta_1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
+&eDNA concentration (K and E) & \\
+$\gamma_0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\gamma_1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
+&of Ct values and eDNA concentration (K and E) & \\
+$\alpha_s$& species amplification efficiency relative to reference species (R) & fixed (SI Appendix Table \ref{tab:alphas})\\
+$\zeta$& ethanol wash effect & fixed (-1.15)\\
+&&\\
+&Index&\\
+\hline
+$s$& species ($H$ = Hake) & -\\
+$j$& location (northing(x), easting(y), depth(d)))& -\\
+$i$& biological replicate & -\\
+$r$& technical replicate & -\\
+$p$& qPCR plate & -\\
+$m$& mock community sample & -\\
+$q$& qPCR standard aliquote sample & -\\
+$R$& metabarcoding reference species & -\\
+
+    \end{tabular}
+    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions.
+$N$ and $H-N$ designate a normal and half-normal distribution, respectively.}
+    \label{tab:priortable}
+\end{table}
+
+\begin{table}\centering
+\caption{Species specific metabarcoding amplification efficiency relative to reference species (R; \textit{Merluccius productus} (Pacific hake) here).}
+    \begin{tabular}{lc}
+\textbf{Species} & $\alpha$ \\
+\textit{Clupea pallasii} &   -0.00138 \\
+\textit{Engraulis mordax} &    0.00557 \\
+\textit{Leuroglossus stilbius} &   -0.00390 \\
+\textit{Microstomus pacificus} &    0.00194 \\
+\textit{Sardinops sagax} &    0.00705 \\
+\textit{Scomber japonicus} &    0.00434 \\
+\textit{Sebastes entomelas} &    0.00459 \\
+\textit{Stenobrachius leucopsarus} &    0.00970 \\
+\textit{Tactostoma macropus} &   -0.00091 \\
+\textit{Tarletonbeania crenularis} &    0.00144 \\
+\textit{Thaleichthys pacificus} &   -0.00506 \\
+\textit{Trachurus symmetricus} &   -0.00875 \\
+\textit{Merluccius productus} &    0.00000 \\
+    \end{tabular}
+        \label{tab:alphas}
+\end{table}
+
+\clearpage
+\bibliography{clean_bib}
 

### 2025-03-06 older versions/Main_text.tex
older versions folder with old .tex.
Updates from Overleaf

--- 2025-02-11 Main_text.tex
@@
 \documentclass[9pt,twocolumn,twoside]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-\usepackage{soul} % Delete this line before submission XXX
 \templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} %= Template for a one-column mathematics article
-% {pnasinvited} %= Template for a PNAS invited submission
 
@@
 \affil[a]{School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA}
-\affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}
+\affil[b]{Conservation Biology Division, Northwest Fisheries Science Center,  NOAA Fisheries, 2715 Montlake Blvd E. Seattle, Washington, USA}
 \affil[c]{Museo Nacional de Ciencias Naturales, Madrid, Spain}
@@
 GG, OL, RPK, and AOS interpreted the results;
-GG, OL, RPK, and AOS wrote the manuscript
-\hl{[Please check that this is correct and add/edit as needed XXX]}
+All authors provided edits to the manuscript.
 }
@@
 
-% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{eDNA $|$ Bayesian $|$ bio-monitoring $|$ quantification $|$ fish}
+\keywords{eDNA $|$ natural resources $|$ bio-monitoring $|$ quantification $|$ fish}
 
@@
 These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously.
-Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological research.
+Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological interest.
 However, unlike single-species techniques, metabarcoding by itself can reflect only the proportions of target-species DNA present, not their absolute quantities.
 Here, we combine qPCR and metabarcoding data derived from the same samples to map quantities of many fish species along the US West Coast in three dimensions, demonstrating a technique of practical importance for both management and ecology.
-We derive spatially explicit estimates of eDNA abundance for 12 common species of commercial and management importance and point the way to quantitative surveys of wild species using molecules alone.
+We derive spatially explicit estimates of eDNA abundance for 12 common species of ecological and management importance and point the way to quantitative surveys of wild species using molecules alone.
 We find that species distribution maps derived from eDNA effectively capture species niches and spatial distributions.
@@
 
-\firstpage[4]{4}
-Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023a}.
+\firstpage{4}
+Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023}.
 However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2022a}).
 The use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
-there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2024}, and coastal oceans \cite{maes2023, shelton2019}).
+there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2023}, and coastal oceans \cite{maes2023, shelton2019}).
 However, uses of quantitative eDNA observations in management applications have significantly lagged behind their use in occurrence and biodiversity applications.
@@
 
+Previous eDNA work has tended to feature more qualitative multispecies results, or else quantitative analysis of single species at small spatial scales.
+Here, we create a rigorous quantitative framework and use hundreds of water samples to make ecological inferences at the scale of management.
 We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
@@
 \ref{fig:fig3}).
-Particularly for forage fish species – which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} – molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
+Particularly for forage fish species -- which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} -- molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
 
@@
 S1}.
+In particular, our analyses present detailed, depth-specific information on abundance that is difficult to obtain with traditional sampling methods.
 
-\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
+\begin{figure*}[tbhp]
 \centering
@@
 Northern anchovy (\textit{Engraulis mordax}) and Pacific sardine (\textit{Sardinops sagax}) are the two iconic coastal pelagic species of the eastern North Pacfic \cite{steinbeck1945} and play a vital ecological role linking lower and upper trophic levels (e.g., Chinook salmon diets, sea lions, \cite{kaplan2019}); historically these supported large commercial fisheries particularly in California.
-Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2°N; Fig.
+Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2$^\circ$N; Fig.
 \ref{fig:fig1}).
-Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24°C for sardine vs.
-10–14°C for anchovy); Pacific sardine also typically inhabits slightly greater depths (25; Fig.
+Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24$^\circ$C for sardine vs.
+10–14$^\circ$C for anchovy).
+We found more Pacific sardine DNA in sub-surface waters as well (Fig.
 \ref{fig:fig1}).
-eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41°N (range 40°–45°), while anchovy eDNA abundance was greatest at 43.5°N (range 42°–48°; \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
-Furthermore, eDNA abundance patterns aligned with stock assessments in indicating a higher abundance of northern anchovy; however, the magnitude of difference varied, with eDNA suggesting a 2.5-fold difference, while stock assessments indicated a 35-fold difference in abundance.
-\hl{[CITE stock assessment here; OLE XXX]}
+eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41$^\circ$N (range 40$^\circ$–45$^\circ$), while anchovy eDNA abundance was greatest at 43.5$^\circ$N (range 42$^\circ$–48$^\circ$; Fig.
+\ref{fig:fig4}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
+Furthermore, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
+However, due to differences in stock definitions and survey areas (the anchovy stock assessment includes southern California waters (32.5$^\circ$–36$^\circ$N) not covered by the eDNA survey) these magnitudes are not directly comparable.
 
 \subsection*{Mid-water}
-Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024} \hl{[ADD SCUBA REF; OLE XXX]}.
-Moreover, because it is a common bycatch species in the hake (\textit{Merluccius productus}) fishery, limits on bycatch of widow rockfish can also limit the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
-Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat.
-We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Figs \ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
+Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024, laidig2007}.
+Moreover, because it is a common bycatch species in the Pacific hake (\textit{Merluccius productus}) fishery, concerns about bycatch of widow rockfish can also influence the much larger hake fishery (i.e., it is a “choke” species for hake; \cite{somers2018}).
+Despite its commercial importance, no targeted survey effort exists for widow rockfish, in part because bottom trawls do not sample its pelagic habitat well.
+We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Fig.
+\ref{fig:fig1} and \ref{fig:fig4}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
 
-The eDNA patterns for northern lampfish (Stenobrachius leucopsarus) -- an abundant myctophid that migrates from 400-700m up to 20-200m at night \cite{moku2000,suntsov2008} -- are again consistent with the ecology of the species, having notably lower concentrations in surface samples.
+The eDNA patterns for northern lampfish (\textit{Stenobrachius leucopsarus}) -- an abundant myctophid that migrates from 400-700m up to 20-200m at night \cite{moku2000,suntsov2008} -- are again consistent with the ecology of the species, having notably lower concentrations in surface samples.
 Owing to the diel vertical migration of the northern lampfish, the signal of eDNA for this species is observed across depths.
@@
 Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common, yet under studied, such as the dragonfish (\textit{Tactostoma macropus}).
-Available literature suggests dragonfish occur roughly between 40°N-50°N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
-The eDNA concentrations appeared to reflect this activity (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
+Available literature suggests dragonfish occur roughly between 40$^\circ$N-50$^\circ$N (adults at 300-900m, larvae at 0-60m in summer) \cite{kawaguchi1993} and perform diel vertical migrations to around 100 meters \cite{kawaguchi1993,willis1982}.
+The eDNA concentrations appeared to reflect this activity (Fig.
+\ref{fig:fig4}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
 \ref{fig:fig1}), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
@@
 Also consistent with known habitats are commercially caught flatfish such as Dover sole (\textit{Microstomus pacificus}), which lives at 200m to 1200m along the Pacific coast \cite{drazen2007,brodziak2000,drazen2012}; individuals move progressively deeper as they grow \cite{vetter1994,hunter1990}.
-The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected closer to the seabed (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}\textit{A}) \cite{ono2016}.
+The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (Fig.
+\ref{fig:fig4}\textit{A}) \cite{ono2016}.
 
@@
 
-Pairwise correlations of DNA concentration showed interesting patterns among species associated with depth (Fig.
+Pairwise correlations of DNA concentration showed meaningful patterns among species associated with depth (Fig.
 \ref{fig:fig2}).
-Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$).
-Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$).
-Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$) with both more common in the more southern (and somewhat warmer) waters (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
+Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$; Fig.
+\ref{fig:fig2}).
+Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$; Fig.
+\ref{fig:fig2}).
+Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$; Fig.
+\ref{fig:fig2}) with both more common in the more southern (and somewhat warmer) waters (Fig.
+\ref{fig:fig1} and \ref{fig:fig4}).
 Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig.
-\ref{fig:fig2}) consistent with more northerly distribution (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
-Together these finding suggest that eDNA is capturing both depth and latitudinal habitats associations.
+\ref{fig:fig2}) consistent with more northerly distribution (Fig.
+\ref{fig:fig4}).
+Together these findings suggest that eDNA is capturing both depth and latitudinal habitat associations.
 Deep-water species are distinct (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}) and not strongly correlated with those from other zones (Fig.
+\ref{fig:fig1} and \ref{fig:fig4}) and not strongly correlated with those from other zones (Fig.
 \ref{fig:fig2}).
 Deeper water tends to feature lower and more diffuse fish eDNA concentrations (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S2}).
+\ref{fig:fig1} and \ref{fig:fig4}).
 Lower overall eDNA concentrations and fewer deep water samples may make similarities among species distributions more difficult to detect at 300m and deeper.
 
-\begin{figure}[tbhp]  % Use figure* for a column-wide figure
+\begin{figure}[tbhp]
 \centering
-\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=8.5cm]{plots/2_Figure_2.jpg}
 \caption{Correlation plot describing species co-abundance and co-occurance.
-Spearman’s rho correlation coefficient are expressed in numbers and colors ranging from -1 (red) negatively correlated to 1 (blue) positively correlated.
+Spearman’s $\rho$ correlation coefficient are expressed in numbers and colors ranging from -1 (red) negatively correlated to 1 (blue) positively correlated.
 Species are ordered in known ecological groups (surface, midwater, and deepwater) to visually indicate the gradual change of species co-occurance.}
@@
 \caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m).
-(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+(B) Species richness represented as the cumulative number of species present (species with higher than $\sim$1.5 copies/L (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S2} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
@@
 
-%\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
-%\centering
-%\includegraphics[width=17cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
-%\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
+\begin{figure*}[tbhp] 
+\centering
+\includegraphics[width=17cm]{plots/4_Figure_4.jpg}
+\caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
 Mean eDNA concentration of vertical distribution is shown with dots and 99\% upper quantile with ticks (A).
@@
 We also identify aggregations of richness and total eDNA near high-productivity regions of ecological and commercial value.
-Regions like Heceta Bank host dense aggregations of forage fish and are critical for both fishery management and for broader marine food-web stability.
 
 For well-studied taxa, existing information helps contextualize eDNA signals, providing insights into habitat preferences, seasonal patterns, or temperature tolerances.
-For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa, revealing unexpected habitats.
-Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \hl{[NWFSC trawl survey ref; OLE XXX]}.
+For poorly studied species, eDNA serves as an exploratory tool, detecting elusive taxa and revealing unexpected habitats.
+Widow rockfish, for example, appear almost exclusively in pelagic habitats around 50m deep, which are sampled poorly by traditional bottom-trawl surveys \cite{keller2017}.
 This species has been of special interest, as its targeted commercial harvest resumed in the late 2010s after a decade of closure (harvest was less than 500 mt annually between 2003 and 2013 but has exceeded 10,000 mt annually since 2018) and no existing survey adequately captures its patterns of abundance \cite{adams2019a}.
@@
 \subsection*{Environmental Smoothing}
-Our findings underline eDNA as a transformative tool for fisheries management, biodiversity conservation, and the study of environmental change.
-However, the observable eDNA signal at any given location and time is an integrated product of DNA production and loss across time and space \cite{gold2023}: the vast majority of DNA detectable in a given sample is of recent origin (within the previous hours to days \cite{strickler2015,collins2018,machler2018}), and it may be subject to a degree of physical transport within its window of detectability.
+Our findings illustrate the power of eDNA as a transformative tool for fisheries management, biodiversity conservation, and the study of environmental change.
+However, the observable eDNA signal at any given location and time is an integrated product of DNA production and loss across time and space \cite{gold2023}: the vast majority of DNA detectable in a given sample is of recent origin (within the previous hours to days \cite{strickler2015, collins2018, machler2018}), and it may be subject to a degree of physical transport within its window of detectability.
 Accordingly, the vast majority of DNA will be relatively proximate to its source in both space and time, and the distribution of DNA from a given species at any given time will be smoother and more extensive when compared to the distribution of discrete source individuals.
@@
 
-The majority of eDNA shed into the environment dissipates from surface waters within hours \hl{(Brasseale et al., in review)[add this reference; RYAN or OLE XXX]}.
-Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most.
-Given that our sampling stations are spaced more than 30 km apart, the spatial patterns we report are likely to reflect genuine differences in fish abundance rather than artifacts of eDNA transport dynamics \cite{guri2023}.
-Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; \textit{Sebastes entomelas} occurs almost exclusively at the 50m depth bin; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig.
+The majority of eDNA shed into the environment dissipates from surface waters within hours \cite{andruszkiewiczallan2021}.
+Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most \cite{baetscher2023}.
+Given that our sampling area spans more than 1,200 km north to south and tens of km east to west, the spatial patterns we report are likely to reflect genuine differences in fish communities rather than artifacts of eDNA transport dynamics \cite{guri2023}.
+Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; Widow rockfish occurs almost exclusively at the 50m depth; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig.
 \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station.
+
 Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig.
-S7}) and spatially smoothed maps (Fig.
+S6}) and spatially smoothed maps (Fig.
 \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound.
-This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism.
-\hl{[Check the framing of the last two sentences XXX]}
-
+S1}) revealed minimal differences.
 Large-scale sampling combined with statistical smoothing mitigates many challenges associated with eDNA transport and degradation -- which decouple the eDNA observations from their particular source organisms -- allowing reliable detection of broad patterns in species distributions.
-Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature \cite{jo2019b}.
-
-%(see \href{SI_Appendix.pdf}{SI Appendix, Fig. S6}) posterior predictive
+Further methodological improvements may incorporate information about the ecology of eDNA itself, such as the degree to which observations of eDNA concentrations depend upon water temperature or other covariates \cite{jo2019b}.
 
 \subsection*{Integrating DNA with traditional surveys}
-Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes.
-Here, we follow \cite{shelton2022} and \hl{Liu et al.
-(submitted)[add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance.
-Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision)[add this reference; RYAN or OLE XXX]}.
+Effective management of natural resources at large scales requires that we maximize the information content of existing sampling regimes.
+Here, we follow \cite{shelton2022}, leveraging the same water samples to provide information on an additional 12 species of ecosystem and management importance.
+Elsewhere, a subset of these samples has yielded information on the distribution of marine mammals \cite{valdiviacarrillo2025}.
 The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa.
@@
 
-Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
-For example, \cite{diaz2020a} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
+Moreover, integrating retrospective eDNA analyses a changing climate can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
+For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
 Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time.
@@
 
-We have shown an example of the substantial value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
-Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional trawl-based surveys.
+We have shown an example of the potential value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024, stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
+Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
+We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
 
-Conventional fish surveys, such as trawling or mark-recapture studies, typically rely on a closure assumption, treating a fish population as if it were static and confined during the sampling window \cite{bailey2014}.
-In reality, fish can move in between sampling location within the study area, which may bias abundance estimates  \cite{rota2009}.
-By contrast, fish continuously shed DNA into their surroundings, creating a dynamic signal of presence.
-As the shedded DNA decays, disperses, and degrades over time, fresher DNA remains at higher concentrations while older DNA persists at lower concentrations \cite{thomsen2015,jerde2011}.
+Conventional fish surveys, such as trawling or mark-recapture methods, typically rely on a closure assumption assuming that fish populations are static during the sampling period \cite{bailey2014}, ignoring the reality of fish movement within the study area thus introducing potential bias into abundance estimates \cite{rota2009}.
+By contrast, eDNA analysis leverages the continuous shedding of DNA by fish, which decays over time with higher concentrations remaining fresh while older DNA persists at lower levels \cite{thomsen2015, jerde2011}.
 This cumulative, time-integrated signal allows eDNA surveys to circumvent the closure assumption and provide a more accurate representation of fish presence and activity.
 
-Perhaps most notably, management decisions happen in units of mass or abundance, not in units of eDNA concentration.
-Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that are likely to vary among species.
-We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
+\matmethods{We summarize all methodological components here and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.}
 
-\matmethods{We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix}.
+\showmatmethods{}
 
 \subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
-Samples were collected in 10L Niskin bottles deployed on a CTD rosette at five depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
-Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey.
-2.5L water samples were collected from the Niskins and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
+We used samples collected in 10L Niskin bottles deployed on a CTD rosette at six depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
+Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey \cite{deblois2020}.
+2.5L water samples were collected from each Niskin and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
 A subset of the samples collected in \cite{shelton2022}, was used for this study.
@@
 In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
-Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix} for details on metabarcoding methods).
+Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods} for details on metabarcoding methods).
+Importantly, we only analyzed one water sample collected from each sampling station using metabarcoding.
 
@@
 The sequences were then BLASTed against the NCBI nucleotide (nt) database (access: August 2024) using BLASTn algorithm with a cut off at 97\% identity, assigning the least common ancestor of the top hit, resulting in species, genus, family, or higher rank classifications.
-Samples with fewer total number of reads than 1000 were filtered as indicator of poor reliability.
+Samples with fewer total number of reads than 1000 were filtered as an indicator of poor reliability.
 Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S4}) due to presence detection of Pacific hake, the reference and link species for this study (see \href{SI_Appendix.pdf}{SI Appendix}).
-Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix}.
-
-\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
-\centering
-\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
-\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species.
-Model compartment indicated in the same color (2-5) are run jointly.
-Latent parameters, state variables, data and subscripts are indicated in gray.
-See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
-\label{fig:DAG}
-\end{figure*}
+S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
+Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
 
@@
 To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples.
-The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix}, but we provide an overview of the approach here (Fig.
-\ref{fig:DAG}).
+The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}, with an overview of the approach (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
 We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
-This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig.
-\ref{fig:DAG}).
-This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification) as well as the effect of covariates (e.g., the depth of the ocean bottom) on hake DNA concentration (Fig.
-\ref{fig:DAG}).
-This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions coming from DNA metabarcoding (Fig.
-\ref{fig:DAG}).
+This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
+This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
+This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2).
 
@@
 Furthermore, metabarcoding does not yield a consistent, species-specific detection probability (i.e., sensitivity).
-Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification, as well as on the total sequencing depth for that sample \cite{kelly2019}.
-Higher sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
+Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification \cite{shelton2023}, as well as on the total sequencing depth for that sample \cite{kelly2019}.
+Greater sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
 
-We follow \cite{shelton2022} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR.
+We follow \cite{shelton2023} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR.
 We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species).
@@
 Our model generates estimates of DNA concentration (copies/L) for each species.
-Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species; we refer to hake as our “reference species”.
+Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species.
 Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species.
 To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurance, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
 S5}).
 Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), and their omission appears not to have induced substantial bias in our estimates.
+S4}), and deserve further attention in future work.
 
@@
 
-The above statistical model (Fig.
-\ref{fig:DAG}) provides predictions of species-specific DNA concentrations (D) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-We used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA (Fig.
-\ref{fig:DAG}, green box).
-Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
+The above statistical model (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2) provides predictions of species-specific DNA concentrations (\textit{D}) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (\href{SI_Appendix.pdf}{SI Appendix} Fig.
+S2, green box).
 
-All statistical analysis, figure preparation, and data curation were performed in R \cite{rcoreteam2024}.
-
-}
-
-\showmatmethods{} % Display the Materials and Methods section
+Analyses, figure preparation, and data curation were performed in R \cite{rcoreteam2024} and the joint statistical model for qPCR and metabarcoding was coded in stan \cite{standevelopmentteam2023}.
 
 \acknow{We thank Linda Park for her foresight and perseverance, Katherine Maslenikov (University of Washington Burke Museum Ichthyology Collection) for providing tissue samples for the mock community analysis, and Olivia Scott for validation and Sanger sequencing of the mock communities.
-This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719).
-The authors gratefully acknowledge the survey expertise and support of NOAA's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
+This material is based upon research supported by the Office of Naval Research under Award Number (N00014-22-1-2719) and support from the National Marine Fisheries Service's Genomic Strategic Initiative.
+The authors gratefully acknowledge the survey expertise and support of NOAA NWFSC's Pacific hake survey team and the personnel of the NOAA Ship Bell M.
 Shimada for supporting the collection of eDNA samples on the west coast hake survey 2019 - 2023.
@@
 
-\showacknow{} % Display the acknowledgments section
+\showacknow{}
 
-\bibsplit[9]
-%Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
+\bibsplit[12]
+
 % Bibliography
-\bibliography{pnas-sample}
+\bibliography{clean_bib}
 

### 2025-04-02 Appendix_plain_v3.tex
update

@@
 Starting with Illumina-demultiplexed sequencing files, we first removed primer and adapter sequences using Cutadapt v4.9, applying the default mismatch tolerance parameter (-e 0.1) \cite{martin2011}.
-Next, we performed quality filtering using the filterAndTrim function from the dada2 R package, retaining its default parameters while specifying the optional truncLen argument to match the minimum MiFish amplicon length.
-After filtering, we denoised the sequences using dada2 with default settings \cite{callahan2016}, which resulted in a list of Amplicon Sequence Variants (ASVs) across all samples.
+Next, we performed quality filtering using the filterAndTrim function from the DADA2 R package, retaining its default parameters while specifying the optional truncLen argument to match the minimum MiFish amplicon length.
+After filtering, we denoised the sequences using DADA2 with default settings \cite{callahan2016}, which resulted in a list of Amplicon Sequence Variants (ASVs) across all samples.
 Each resulting ASV was then assigned a unique 28-character hash code using the sha-1 algorithm with the R function map\_chr.
@@
 The above statistical model provides predictions of species-specific DNA concentrations ($E$) at each depth sampled (d = 0, 50, 150, 300 or 500 m).
-To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model (SI Appendix Figure S.\ref{fig:DAG_v3}), exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
+To generate smooth distribution models, we use posterior means of $E_s(x,y,d)$ from the joint statistical model, exponentiated it to move to concentration from log concentration, $G_{s}(x,y,d) = exp(E_{s}(x,y,d))$, and used the sdmTMB package \cite{anderson2022} to generate distribution models for the concentration of eDNA.
 Because of the predominance of very low estimated DNA concentrations at a large number of site, we set $G_{s}(x,y,d) < 0.5$ equal to zero (SI Appendix Figure S.\ref{fig:presence_threshold}) and then used a Tweedie likelihood in sdmTMB,
@@
 
-\begin{figure*}[tbhp] 
-\centering
-\includegraphics[width=16cm]{plots/DAG_v3.pdf} 
-\caption{Schematic for the statistical models to produce species distributions.
+%\begin{figure*}[tbhp] 
+%\centering
+%\includegraphics[width=16cm]{plots/DAG_v3.pdf} 
+%\caption{Schematic for the statistical models to produce species distributions.
 Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
@@
 Notation for parameters, subscripts, and observations are indicated in gray.}
-\label{fig:DAG_v3}
-\end{figure*}
+%\label{fig:DAG_v3}
+%\end{figure*}
 

### 2025-04-02 Manuscript_plain_v3.tex
update

@@
 
-\section*{Significance statement}
-Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales.
+%\section*{Significance statement}
+%Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales.
 We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States.
@@
 We subsequently smooth the estimated eDNA quantities to generate species distribution maps \cite{liu2023}.
-We present results for 12 fish species to illustrate the approach, but note that it immediately generalizes to any ecological community of interest.
+We present quantitative results for 12 fish species using eDNA, comparing them with literature-based species distributions to illustrate the approach, but note that it immediately generalizes to any ecological community of interest.
 
@@
 
+\section*{Methods}
+
+\subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
+We used samples collected in 10L Niskin bottles deployed on a CTD rosette at six depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
+Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey \cite{deblois2020}.
+2.5L water samples were collected from each Niskin and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
+A subset of the samples (n = 554) collected in \cite{shelton2022}, was used for metabarcoding analysis this study in addition to qPCR samples already published for hake quantification \cite{shelton2022} (n= 1,794).
+
+\subsection*{Metabarcoding: Environmental Samples and Mock Community}
+In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
+Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods} for details on metabarcoding methods).
+Importantly, we only analyzed one water sample collected from each sampling station using metabarcoding.
+
+Sequences were de-multiplexed and the adapters removed by Illumina sequencing.
+Primers were removed with Cutadapt v4.9 \cite{martin2011}.
+We then used DADA2 (default parameters, \cite{callahan2016}) to denoise sequences, remove chimeras, and generate amplicon sequence variants (ASVs).
+The sequences were then BLASTed against the NCBI nucleotide (nt) database (access: August 2024) using BLASTn algorithm with a cut off at 97\% identity, assigning the least common ancestor of the top hit, resulting in species, genus, family, or higher rank classifications.
+Samples with fewer total number of reads than 1000 were filtered as an indicator of poor reliability.
+Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
+Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
+
+To calibrate metabarcoding observations and account for amplification bias \cite{shelton2023,gold2023}, we generated multiple mock communities comprising a total of 39 fish species.
+Vouchered DNA extracts or tissues were obtained from either the University of Washington Fish Collection at the Burke Museum or the NOAA Northwest Fisheries Science Center collections.
+After quantifying the concentration of mitochondrial DNA template in each DNA extract, we constructed a total of eight mock communities of varying species compositions (see \href{SI_Appendix.pdf}{SI Appendix, Table S1}).
+We selected 12 species (additional to Pacific hake) of commercial and ecological importance that co-occurred in the environmental samples and in the mock community for downstream analysis: Pacific herring (\textit{Clupea pallasii}), northern anchovy (\textit{Engraulis mordax}), northern smoothtongue (\textit{Leuroglossus stilbius}), Dover sole (\textit{Microstomus pacificus}), Pacific sardine (\textit{Sardinops sagax}), Pacific chub mackerel (\textit{Scomber japonicus}), widow rockfish (\textit{Sebastes entomelas}), northern lampfish (\textit{Stenobrachius leucopsarus}), California barracudina (\textit{Tactostoma macropus}), plainfin midshipman (\textit{Tarletonbeania crenularis}), eulachon (\textit{Thaleichthys pacificus}), Pacific jack mackerel (\textit{Trachurus symmetricus}).
+After amplification and sequencing using the protocols outlined above, we used the mock communities (which were run in the same PCR conditions as the environmental samples) to estimate species-specific amplification efficiencies and correct for amplification bias following \cite{shelton2023}.
+
+\subsection*{Statistical analysis}
+To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples.
+This model is based on the QM R package (www.github.com/gledguri/QM), with some modifications and advancements.
+The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}, with an overview of the approach (Fig.
+\ref{fig:DAG_v3}).
+We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
+This model takes observations of the PCR cycle at which amplification was detected in each field sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig.
+\ref{fig:DAG_v3}).
+This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (Fig.
+\ref{fig:DAG_v3}).
+This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (Fig.
+\ref{fig:DAG_v3}).
+
+\begin{figure*}[tbhp] 
+\centering
+\includegraphics[width=16cm]{plots/DAG_v3.pdf} 
+\caption{Schematic for the statistical models to produce species distributions.
+Three sets of observations (orange) were used simultaneously to provide estimates of eDNA concentration for  twelve fish species (blue) from which we derive smoothed predictions of DNA concentration (green).
+Mock communities for metabarcoding (yellow) were estimated externally and provided estimated $\alpha_s$ that were used in the joint model.
+Notation for parameters, subscripts, and observations are indicated in gray.}
+\label{fig:DAG_v3}
+\end{figure*}
+
+Unlike single-species observation methods, metabarcoding data is compositional data.
+This means that each sample only provides information about relative species abundances rather than absolute DNA concentrations.
+Furthermore, metabarcoding does not yield a consistent, species-specific detection probability (i.e., sensitivity).
+Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification \cite{shelton2023}, as well as on the total sequencing depth for that sample \cite{kelly2019}.
+Greater sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
+
+We follow \cite{shelton2023} and construct a compositional model for the metabarcoding data from 371 of the water samples (after quality filtering) analyzed for Pacific hake qPCR.
+We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species).
+In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample.
+But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig.
+S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
+Our model generates estimates of DNA concentration (copies/L) for each species.
+Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species.
+Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species.
+To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
+S5}).
+Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
+S4}), and deserve further attention in future work.
+
+We focused on species that were relatively common in the raw dataset – those well above the lower detection limits under our sampling and sequencing conditions – ensuring greater confidence in their consistent detectability.
+Although many more species were detected in the raw data, the majority were much rarer and thus less reliably observed.
+
+The above statistical model (Fig.
+\ref{fig:DAG_v3}) provides predictions of species-specific DNA concentrations at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+We used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (Fig.
+\ref{fig:DAG_v3}, green box).
+As Pacific hake (\textit{Merluccius productus}) has been examined elsewhere using qPCR \cite{shelton2022}, we did not include the spatial patterns of that species in our analysis.
+
+Analyses, figure preparation, and data curation were performed in R \cite{rcoreteam2024} and the joint statistical model for qPCR and metabarcoding was coded in stan \cite{standevelopmentteam2023}.
+For a simplified version of this model, the QM R package can be utilized (www.github.com/gledguri/QM).
 \section*{Results}
@@
 Here, we create a rigorous quantitative framework and use hundreds of water samples to make ecological inferences at the scale of management.
-We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
+In this study, we find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
 \ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
@@
 \ref{fig:fig3}).
-Particularly for forage fish species -- which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} -- molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
+Particularly for forage fish species -- which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} -- molecular methods offer atractable and quantitative index of abundance over large spatial scales.
 
@@
 \includegraphics[width=16cm]{plots/1_Figure_1.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
-\caption{Estimated species eDNA concentration (A) across 0, 50, 150, 300, and 500m depth samples and known depth distribution of those species from literature (B).}
+\caption{Estimated species (n=6; for the remaining species see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S1}) eDNA concentration (A) across 0, 50, 150, 300, and 500m depth samples and known depth distribution of those species from literature (B).}
 \label{fig:fig1}
@@
 
-\subsection*{Surface}
+\subsection*{Surface species}
 Northern anchovy (\textit{Engraulis mordax}) and Pacific sardine (\textit{Sardinops sagax}) are the two iconic coastal pelagic species of the eastern North Pacfic \cite{steinbeck1945} and play a vital ecological role linking lower and upper trophic levels (e.g., Chinook salmon diets, sea lions, \cite{kaplan2019}); historically these supported large commercial fisheries particularly in California.
-Our eDNA estimates consistently reflect the expected three-dimensional distributions of each species, with high concentrations in surface waters for both species at northern latitudes \cite{litz2008,parnel2008} particularly near the Columbia River outflow (46.2$^\circ$N; Fig.
+While larvae and juveniles typically occupy the upper 50m, adults move to 100m during the day to avoid predators.
+Our eDNA estimates consistently reflect the expected three-dimensional distributions of known surface dwelling species, with high concentrations in surface waters for both species\cite{litz2008,parnel2008}, and an accumulation pocket near the Columbia River outflow (46.2$^\circ$N; Fig.
 \ref{fig:fig1}).
@@
 10–14$^\circ$C for anchovy).
-We found more Pacific sardine DNA in sub-surface waters as well (Fig.
-\ref{fig:fig1}).
 eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41$^\circ$N (range 40$^\circ$–45$^\circ$), while anchovy eDNA abundance was greatest at 43.5$^\circ$N (range 42$^\circ$–48$^\circ$; Fig.
 \ref{fig:fig4}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
-Furthermore, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
+When comparing the magnitude difference between the two species, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
 However, due to differences in stock definitions and survey areas (the anchovy stock assessment includes southern California waters (32.5$^\circ$–36$^\circ$N) not covered by the eDNA survey) these magnitudes are not directly comparable.
 
-\subsection*{Midwater}
+\subsection*{Midwater species}
 Widow rockfish (\textit{Sebastes entomelas}) is common between 100-350m between northern Baja California and southern Alaska, with pelagic larvae and juveniles generally detected in shallower habitats ($<$ 150m depth) \cite{reynolds2001,bosley2014} including in nearshore kelp forests \cite{quigley2024, laidig2007} .
@@
 We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Fig.
-\ref{fig:fig1} and \ref{fig:fig4}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults are commonly found \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
+\ref{fig:fig1} and \ref{fig:fig4}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
 
@@
 
-\subsection*{Deepwater}
+\subsection*{Deepwater species}
 Among deeper-water fishes, eDNA revealed spatial and potentially ontogenetic migration patterns for species that are common, yet under studied, such as the dragonfish (\textit{Tactostoma macropus}).
@@
 \ref{fig:fig4}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
-\ref{fig:fig1}), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
+\ref{fig:fig1}; near Columbia River discharge), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
 However, no other quantitative estimates of abundance or detailed distribution patterns for dragonfish exist in the current literature.
@@
 The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (Fig.
-\ref{fig:fig4}\textit{A}) \cite{ono2016}.
+\ref{fig:fig1}) \cite{ono2016}.
 
@@
 \ref{fig:fig2}).
-Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$).
-Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$).
-Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$) with both more common in the more southern (and somewhat warmer) waters (Fig.
+Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$; Fig.
+\ref{fig:fig2}).
+Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$; Fig.
+\ref{fig:fig2}).
+Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$; Fig.
+\ref{fig:fig2}) with both more common in the more southern (and somewhat warmer) waters (Fig.
 \ref{fig:fig1} and Fig.
@@
 In part, this pattern is likely to be an artifact of the focal species included in this study -- chosen for their relevance to fisheries management as well as for their frequency in the observations (see Methods) -- but the richness-depth gradient also reflects the expectation that habitat and nutrient availability is greatest near the surface \cite{smith2002,hickey2005}.
-Surface waters also host the eggs, larvae, and juvenile stages of many meso- and bathypelagic species \cite{parnel2008}, some of which we detect, and hence we both expect and observe higher richness in the upper layer of the water column \cite{kim2011}.
+Surface waters also host the eggs, larvae, and juvenile stages of many meso- and bathypelagic species \cite{parnel2008}, some of which we have included here, and hence we both expect and observe higher richness in the upper layer of the water column \cite{kim2011}.
 
@@
 The total fish eDNA accumulation (Fig.
-\ref{fig:fig3}\textit{B}) suggests the highest aggregation of forage fish DNA near Heceta Bank (44 degrees N; overlapping anchovy and sardine distributions are the two largest contributors to eDNA concentration) in addition to other commercialy important mesopelagic fish (\textit{Sebastes entomelas} and \textit{Tarletonbeania crenularis}) a center of abundance for commercial fisheries \cite{bosley2014,Tissot2008TwentyYO}, and elsewhere nearshore in areas of coastal upwelling.
+\ref{fig:fig3}\textit{B}) suggests the highest aggregation of forage fish DNA near Heceta Bank (44 degrees N; overlapping anchovy and sardine distributions are the two largest contributors to the observed peak of accumulated eDNA concentration) in addition to other commercialy important mesopelagic fish (\textit{Sebastes entomelas} and \textit{Tarletonbeania crenularis}) a center of abundance for commercial fisheries \cite{bosley2014,Tissot2008TwentyYO}, and elsewhere nearshore in areas of coastal upwelling.
 
@@
 
-\section*{Methods}
-We summarize all methodological components and provide details for all sampling, laboratory, and statistical methods in the \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
-
-\subsection*{Sample Collection, Processing, and Quantitative PCR (qPCR)}
-We used samples collected in 10L Niskin bottles deployed on a CTD rosette at six depths (0, 50, 150, 300, 500 m) from stations along the West Coast of the US aboard the ship \textit{Bell M.
-Shimada} during Summer 2019 on the Joint U.S.–Canada Integrated Ecosystem and Pacific Hake Acoustic-Trawl Survey \cite{deblois2020}.
-2.5L water samples were collected from each Niskin and filtered immediately on 47 mm diameter mixed cellulose-ester sterile filters with a 1 $\mu$m pore size, subsequently extracted using a phase-lock phenol-chloroform protocol, and analyzed in triplicate using a qPCR assay, multiplexed for Pacific hake, sea lamprey, and eulachon, as described in \cite{ramon-laca2021} and \cite{shelton2022}.
-A subset of the samples collected in \cite{shelton2022}, was used for this study.
-
-\subsection*{Metabarcoding: Environmental Samples and Mock Community}
-In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
-Total PCR reactions including PCR blanks (n = 7) and positive controls (n = 7), were amplified using MiFish-U universal primers \cite{miya2015} (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods} for details on metabarcoding methods).
-Importantly, we only analyzed one water sample collected from each sampling station using metabarcoding.
-
-Sequences were de-multiplexed and the adapters removed by Illumina sequencing.
-Primers were removed with Cutadapt v4.9 \cite{martin2011}.
-We then used DADA2 (default parameters, \cite{callahan2016}) to denoise sequences, remove chimeras, and generate amplicon sequence variants (ASVs).
-The sequences were then BLASTed against the NCBI nucleotide (nt) database (access: August 2024) using BLASTn algorithm with a cut off at 97\% identity, assigning the least common ancestor of the top hit, resulting in species, genus, family, or higher rank classifications.
-Samples with fewer total number of reads than 1000 were filtered as an indicator of poor reliability.
-Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
-Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
-
-To calibrate metabarcoding observations and account for amplification bias \cite{shelton2023,gold2023}, we generated multiple mock communities comprising a total of 39 fish species.
-Vouchered DNA extracts or tissues were obtained from either the University of Washington Fish Collection at the Burke Museum or the NOAA Northwest Fisheries Science Center collections.
-After quantifying the concentration of mitochondrial DNA template in each DNA extract, we constructed a total of eight mock communities of varying species compositions (see \href{SI_Appendix.pdf}{SI Appendix, Table S1}).
-We selected 12 species (additional to Pacific hake) of commercial and ecological importance that co-occurred in the environmental samples and in the mock community for downstream analysis: Pacific herring (\textit{Clupea pallasii}), northern anchovy (\textit{Engraulis mordax}), northern smoothtongue (\textit{Leuroglossus stilbius}), Dover sole (\textit{Microstomus pacificus}), Pacific sardine (\textit{Sardinops sagax}), Pacific chub mackerel (\textit{Scomber japonicus}), widow rockfish (\textit{Sebastes entomelas}), northern lampfish (\textit{Stenobrachius leucopsarus}), California barracudina (\textit{Tactostoma macropus}), plainfin midshipman (\textit{Tarletonbeania crenularis}), eulachon (\textit{Thaleichthys pacificus}), Pacific jack mackerel (\textit{Trachurus symmetricus}).
-As Pacific hake (\textit{Merluccius productus}) has been examined elsewhere using qPCR \cite{shelton2022}, we did not include the spatial patterns of that species in our analysis.
-After amplification and sequencing using the protocols outlined above, we used the mock communities to estimate species-specific amplification efficiencies and correct for amplification bias following \cite{shelton2023}.
-
-\subsection*{Statistical analysis}
-To generate estimates of DNA concentration, we constructed a joint statistical model for qPCR and Mifish metabarcoding observations from the same samples.
-The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}, with an overview of the approach (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2).
-We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
-This model takes observations of the PCR cycle at which amplification was detected in each water sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2).
-This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2).
-This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2).
-
-Unlike single-species observation methods, metabarcoding data is compositional data.
-This means that each sample only provides information about relative species abundances rather than absolute DNA concentrations.
-Furthermore, metabarcoding does not yield a consistent, species-specific detection probability (i.e., sensitivity).
-Instead, the likelihood of detecting a particular species depends heavily on its competition with other sequences during PCR amplification \cite{shelton2023}, as well as on the total sequencing depth for that sample \cite{kelly2019}.
-Greater sequencing depth generally increases the probability of identifying rarer species, conditional on their DNA being collected and it being amplifiable with the chosen PCR primers.
-
-We follow \cite{shelton2023} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR.
-We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species).
-In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample.
-But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
-Our model generates estimates of DNA concentration (copies/L) for each species.
-Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species.
-Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species.
-To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S5}).
-Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), and deserve further attention in future work.
-
-We focused on species that were relatively common in the raw dataset – those well above the lower detection limits under our sampling and sequencing conditions – ensuring greater confidence in their consistent detectability.
-Although many more species were detected in the raw data, the majority were much rarer and thus less reliably observed.
-
-The above statistical model (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2) provides predictions of species-specific DNA concentrations (\textit{D}) at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
-To generate smooth maps for each species, we used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (\href{SI_Appendix.pdf}{SI Appendix} Fig.
-S2, green box).
-
 %Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
-
-Analyses, figure preparation, and data curation were performed in R \cite{rcoreteam2024} and the joint statistical model for qPCR and metabarcoding was coded in stan \cite{standevelopmentteam2023}.
 

### 2025-04-03 Manuscript_plain_v3.tex
update

@@
 \documentclass{article}
+\usepackage{graphicx}
 \usepackage[margin=1in]{geometry}
@@
 \usepackage{lineno}
+%\renewcommand\thelinenumber{\large\arabic{linenumber}}
+\renewcommand\thelinenumber{\fontsize{9}{12}\selectfont\arabic{linenumber}} 
+\bibliographystyle{unsrt}
+
 \linenumbers
-\bibliographystyle{plain}
 
@@
  & \qquad 2715 Montlake Blvd E. Seattle, Washington, USA \\
-3 & Museo Nacional de Ciencias Naturales, Madrid, Spain \\
+3 & Museo Nacional de Ciencias Naturales CSIC, Madrid, Spain \\
 \hline
@@
 Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural-resources management at a broad geographic scale.
-These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously.
+These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or ddPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously.
 Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological interest.
@@
 \section*{Introduction}
-Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., Richness, \cite{muenzel2024}), and genetic structure \cite{andres2023}.
-However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022,guri2024a,stoeckle2022a}).
+Environmental DNA (eDNA, residual genetic material sampled from water, soil, or air) is an increasingly common tool for non-invasively sampling marine and aquatic ecological communities, valuable for estimating species occurrence \cite{veilleux2021}, biodiversity (i.e., richness, \cite{muenzel2024}), and genetic structure \cite{andres2023}.
+However, eDNA has not been widely used for quantitative abundance estimation, upon which many practical management and conservation applications rely (but see \cite{shelton2022, guri2024a, stoeckle2022a}).
 The use of eDNA for quantification is supported both conceptually -- where there is more of a species, more cells are shed in the environment, and thus there is more DNA -- and empirically (e.g.
-there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b,ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2023}, and coastal oceans \cite{maes2023, shelton2019}).
+there are strong linkages between single-species eDNA measures of in aquaria \cite{jo2019b, ledger2024}, rivers and streams \cite{pont2024}, estuaries and nearshore habitats \cite{dibattista2022, baetscher2023}, and coastal oceans \cite{maes2023, shelton2019}).
 However, uses of quantitative eDNA observations in management applications have significantly lagged behind their use in occurrence and biodiversity applications.
@@
 
-\subsection*{Metabarcoding: Environmental Samples and Mock Community}
+\subsection*{MetabarcodFig. S3ing: Environmental Samples and Mock Community}
 In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
@@
 
-Sequences were de-multiplexed and the adapters removed by Illumina sequencing.
+Sequences were de-multiplexed and the adapters removed by Illumina processing software tools.
 Primers were removed with Cutadapt v4.9 \cite{martin2011}.
@@
 Samples with fewer total number of reads than 1000 were filtered as an indicator of poor reliability.
-Of the initial environmental samples, 535 passed the quality filtering and and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S4}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
+Of the initial environmental samples, 535 passed the quality filtering and 371 were selected for downstream analysis (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S3}) due to presence detection of Pacific hake in both qPCR and metabarcoding data streams (see \href{SI_Appendix.pdf}{SI Appendix Extended Methods}).
 Details on bioinformatic pipeline are available in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}.
@@
 In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample.
-But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
+But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+S3}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}.
 Our model generates estimates of DNA concentration (copies/L) for each species.
@@
 To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration.
-However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S5}).
-Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig.
-S4}), and deserve further attention in future work.
+However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurrence, like Pacific sardine or northern anchovy in our case (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+S4}).
+Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix, Fig.
+S3}), and deserve further attention in future work.
 
@@
 
-%We find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
-\ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S1}, Pacific Herring and Jack Mackerel).
-For other high-biomass species otherwise lacking well-described distributions, we provide the first species distribution models (Fig.
-\ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
-S1}).
-We find that the estimated eDNA concentrations are correlated among species with similar traits and habitat preferences (Fig.
-\ref{fig:fig2}).
-Emergent patterns of species richness and estimated eDNA aggregation correspond with known ecological drivers, evidence that eDNA traces can reliably capture ecologically meaningful patterns (Fig.
-\ref{fig:fig3}).
-Particularly for forage fish species – which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} – molecular methods offer a tractable and quantitative index of abundance over large spatial scales.
-
 Below we highlight results for six species (Fig.
@@
 \centering
-\includegraphics[width=16cm]{plots/1_Figure_1.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=16cm]{plots/1_Figure_1.jpg} 
 \caption{Estimated species (n=6; for the remaining species see \href{SI_Appendix.pdf}{SI Appendix, Fig.
@@
 In addition to providing species-specific patterns, eDNA metrics of abundance from individual species can be combined to reveal important aggregate ecological patterns.
-For the 12 species considered here, we show how species richness (see \href{SI_Appendix.pdf}{SI Appendix Fig.
-S3} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
+For the 12 species considered here, we show how species richness (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S2} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
 \ref{fig:fig3}\textit{A}).
@@
 \centering
-\includegraphics[width=16cm]{plots/4_Figure_4.jpg}  % Adjust width to 11.4 cm or 17.8 cm as per the template
+\includegraphics[width=16cm]{plots/4_Figure_4.jpg} 
 \caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
@@
 We also identify aggregations of richness and total eDNA near high-productivity regions of ecological and commercial value.
-
-%Regions like Heceta Bank host dense aggregations of forage fish and are critical for both fishery management and for broader marine food-web stability.
 
@@
 Moreover, integrating retrospective eDNA analyses a changing climate can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts.
-For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material.
+For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extend the utility of archived material.
 Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time.
@@
 We have shown an example of the potential value of multi-species eDNA analysis for fisheries and other natural-resource management questions \cite{ledger2024,stoeckle2024}, however, molecular patterns alone are unlikely to drive quantitative natural-resources decisions, in part because eDNA generally does not provide auxiliary information important for managing populations (e.g., information on size- or age structure, fecundity, sex, or disease).
-Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}, and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
+Relating observed eDNA concentrations to species mass requires data external to the eDNA signal \cite{guri2024a}; and in particular, requires that we understand proportionalities between eDNA concentration and biomass that may vary among species.
 We therefore see eDNA analysis as strengthening and extending existing surveys, with different data streams complementing one another.
 
-%Moreover, the eDNA sampling we describe here depends upon survey vessels that can reach and sample from deep water – survey vessels that exist to carry out traditional acoustic or trawl-based surveys.
-Furthermore, eDNA surveys provide data in units of eDNA concentration, not more familiar units of mass or abundance.
-
-Conventional fish surveys, such as trawling or mark-recapture methods, typically rely on a closure assumption assuming that fish populations are static during the sampling period \cite{bailey2014}, ignoring the reality of fish movement within the study area thus introducing potential bias into abundance estimates \cite{rota2009}.
+Conventional fish surveys, such as trawling or mark-recapture methods, typically rely on a closure assumption - assuming static fish population during the sampling period \cite{bailey2014}, which ignores the reality of fish movement within the study area thus introducing potential bias into abundance estimates \cite{rota2009}.
 By contrast, eDNA analysis leverages the continuous shedding of DNA by fish, which decays over time with higher concentrations remaining fresh while older DNA persists at lower levels \cite{thomsen2015,jerde2011}.
@@
 
-%Each species was fitted individually while we created and selected the best mesh for approximating spatial Gaussian random fields as part of the model (\href{SI_Appendix.pdf}{SI Appendix}).
-
-\bibliography{clean_bib.bib}
+\bibliography{clean_bib}
 

### 2025-04-07 Manuscript_plain_v3.tex
update

@@
 \section*{Author Contributions}
-AOS, KMN, RPK, MS, and KP designed the research;\\
-MRS, AR-L and AW did the laboratory work; \\
-PB, RPK, and GG did the bioinformatics;\\
-GG, OL, RPK, and AOS analyzed the data;\\
-GG, OL, RPK, and AOS interpreted the results;\\
+AOS, KMN, RPK, MS, and KP designed the research;
+MRS, AR-L and AW did the laboratory work; 
+PB, RPK, and GG did the bioinformatics;
+GG, OL, RPK, and AOS analyzed the data;
+GG, OL, RPK, and AOS interpreted the results;
 All authors provided edits to the manuscript \\
@@
 
-\subsection*{MetabarcodFig. S3ing: Environmental Samples and Mock Community}
+\subsection*{Metabarcoding: Environmental Samples and Mock Community}
 In total, 554 environmental samples were randomized and divided into 7 sequencing libraries for an Illumina MiSeq (v3 600 cycle kit).
@@
 The full statistical details are described in \href{SI_Appendix.pdf}{SI Appendix Extended Methods}, with an overview of the approach (Fig.
-\ref{fig:DAG_v3}).
+\ref{fig:fig1}).
 We used qPCR observations of hake DNA concentration for all water samples available (1,794 unique water samples were assayed and using a total of 5,394 qPCR reactions).
 This model takes observations of the PCR cycle at which amplification was detected in each field sample and estimates a hake DNA concentration for each water bottle collected at each sampling station-depth combination (Fig.
-\ref{fig:DAG_v3}).
+\ref{fig:fig1}).
 This statistical model accounts for sample-specific effects (e.g., dilution of the sample before PCR amplification), the effect of replicate water samples (a random effect), to provide estimates of hake DNA concentration in from each water sample collected at each location and depth (Fig.
-\ref{fig:DAG_v3}).
+\ref{fig:fig1}).
 This qPCR component of the joint model provides a quantification of DNA concentration for one species (Pacific hake) that can be connected to the multispecies proportions provided by DNA metabarcoding (Fig.
-\ref{fig:DAG_v3}).
+\ref{fig:fig1}).
 
-\begin{figure*}[tbhp] 
+\begin{figure}[tbhp] 
 \centering
@@
 Notation for parameters, subscripts, and observations are indicated in gray.}
-\label{fig:DAG_v3}
-\end{figure*}
+\label{fig:fig1}
+\end{figure}
 
@@
 The above statistical model (Fig.
-\ref{fig:DAG_v3}) provides predictions of species-specific DNA concentrations at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
+\ref{fig:fig1}) provides predictions of species-specific DNA concentrations at each location and depth sampled ($d$ = 0, 50, 150, 300, or 500 m).
 We used the sdmTMB package \cite{anderson2022} to generate smooth maps of DNA concentration for each species independently (Fig.
-\ref{fig:DAG_v3}, green box).
+\ref{fig:fig1}, green box).
 As Pacific hake (\textit{Merluccius productus}) has been examined elsewhere using qPCR \cite{shelton2022}, we did not include the spatial patterns of that species in our analysis.
@@
 In this study, we find spatial patterns of DNA concentrations reflect the distributions of commercially important coastal-pelagic fish species in three dimensions and across ten degrees of latitude in a single year of sampling in 2019 (Fig.
-\ref{fig:fig1}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
+\ref{fig:fig2}: Northern Anchovy, Pacific Sardine; \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}, Pacific Herring and Jack Mackerel).
 For other high-biomass species otherwise lacking well-described distributions, we provide the first species distribution models (Fig.
-\ref{fig:fig1}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
+\ref{fig:fig2}, \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}).
 We find that the estimated eDNA concentrations are correlated among species with similar traits and habitat preferences (Fig.
-\ref{fig:fig2}).
+\ref{fig:fig3}).
 Emergent patterns of species richness and estimated eDNA aggregation correspond with known ecological drivers, evidence that eDNA traces can reliably capture ecologically meaningful patterns (Fig.
-\ref{fig:fig3}).
+\ref{fig:fig4}).
 Particularly for forage fish species -- which are critical high-biomass links in marine food webs, but difficult to survey with traditional methods \cite{Pikitch2012} -- molecular methods offer atractable and quantitative index of abundance over large spatial scales.
@@
 Below we highlight results for six species (Fig.
-\ref{fig:fig1}) representing a range of habitat associations and general species information; distribution maps for all the remaining species are presented in \href{SI_Appendix.pdf}{SI Appendix, Fig.
+\ref{fig:fig2}) representing a range of habitat associations and general species information; distribution maps for all the remaining species are presented in \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}.
@@
 
-\begin{figure*}[tbhp]
+\begin{figure}[tbhp]
 \centering
-\includegraphics[width=16cm]{plots/1_Figure_1.jpg} 
+\includegraphics[width=0.9\textwidth, height=1.17\textwidth]{plots/1_Figure_1.pdf} 
 \caption{Estimated species (n=6; for the remaining species see \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}) eDNA concentration (A) across 0, 50, 150, 300, and 500m depth samples and known depth distribution of those species from literature (B).}
-\label{fig:fig1}
-\end{figure*}
+\label{fig:fig2}
+\end{figure}
 
@@
 Our eDNA estimates consistently reflect the expected three-dimensional distributions of known surface dwelling species, with high concentrations in surface waters for both species\cite{litz2008,parnel2008}, and an accumulation pocket near the Columbia River outflow (46.2$^\circ$N; Fig.
-\ref{fig:fig1}).
+\ref{fig:fig2}).
 Sardine favors slightly warmer temperatures and more southerly waters than anchovy (SST range from literature: 13–24$^\circ$C for sardine vs.
@@
 eDNA patterns tracked these differences with sardine eDNA in greatest abundance at 41$^\circ$N (range 40$^\circ$–45$^\circ$), while anchovy eDNA abundance was greatest at 43.5$^\circ$N (range 42$^\circ$–48$^\circ$; Fig.
-\ref{fig:fig4}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
+\ref{fig:fig5}) such that eDNA captured subtle yet meaningful differences in habitat preferences.
 When comparing the magnitude difference between the two species, eDNA abundance patterns suggest a 2.5-fold higher abundance for anchovy than sardine which agree in direction with stock assessments for anchovy and sardine (an approximately 35-fold difference; \cite{kuriyama2022, kuriyama2022b}).
@@
 We find widow rockfish eDNA strongly and nearly exclusively associated with 50m sample depths (Fig.
-\ref{fig:fig1} and \ref{fig:fig4}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
+\ref{fig:fig2} and \ref{fig:fig5}), consistent with the pelagic larvae and juveniles reported in the literature and somewhat shallower than adults \cite{reynolds2001,bosley2014,quigley2024}, potentially reflecting the species’ nocturnal feeding schooling behavior \cite{reynolds2001,bosley2014}.
 
@@
 The eDNA concentrations appeared to reflect this activity (Fig.
-\ref{fig:fig4}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
-\ref{fig:fig1}; near Columbia River discharge), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
+\ref{fig:fig5}), with the predominant surface signal at northern latitudes likely reflecting larvae and juveniles (Fig.
+\ref{fig:fig2}; near Columbia River discharge), and two deeper hotspots at 300m and 500m separated by about eight degrees of latitude.
 However, no other quantitative estimates of abundance or detailed distribution patterns for dragonfish exist in the current literature.
@@
 The eDNA concentration estimates are consistent with this pattern, showing high concentrations mostly in water samples collected close to the seabed (Fig.
-\ref{fig:fig1}) \cite{ono2016}.
+\ref{fig:fig2}) \cite{ono2016}.
 
@@
 Pairwise correlations of DNA concentration showed meaningful patterns among species associated with depth (Fig.
-\ref{fig:fig2}).
+\ref{fig:fig3}).
 Surface-associated species were nearly all positively correlated with each other, with some correlations particularly large (Pearson's $\rho > 0.50$; Fig.
-\ref{fig:fig2}).
+\ref{fig:fig3}).
 Correlations were weak among the six midwater and deepwater species (all $\rho < |0.40|$; Fig.
-\ref{fig:fig2}).
+\ref{fig:fig3}).
 Within the surface species, further subdivision was also apparent, with chub mackerel (\textit{Scomber japonicus}) and sardine (\textit{Sardinops sagax}) showing very high correlation ($\rho = 0.76$; Fig.
-\ref{fig:fig2}) with both more common in the more southern (and somewhat warmer) waters (Fig.
-\ref{fig:fig1} and Fig.
-\ref{fig:fig4}).
+\ref{fig:fig3}) with both more common in the more southern (and somewhat warmer) waters (Fig.
+\ref{fig:fig2} and Fig.
+\ref{fig:fig5}).
 Concentrations of the remaining surface-dwelling species (Northern anchovy, eulachon, jack mackerel, and Pacific herring) had modest correlations with midwater species (Fig.
-\ref{fig:fig2}) consistent with more northerly distribution (Fig.
-\ref{fig:fig4}).
+\ref{fig:fig3}) consistent with more northerly distribution (Fig.
+\ref{fig:fig5}).
 Together these findings suggest that eDNA is capturing both depth and latitudinal habitat associations.
 Deep-water species are distinct (Fig.
-\ref{fig:fig1} and Fig.
-\ref{fig:fig4}) and not strongly correlated with those from other zones (Fig.
-\ref{fig:fig2}).
+\ref{fig:fig2} and Fig.
+\ref{fig:fig5}) and not strongly correlated with those from other zones (Fig.
+\ref{fig:fig3}).
 Deeper water tends to feature lower and more diffuse fish eDNA concentrations (Fig.
-\ref{fig:fig1} and Fig.
-\ref{fig:fig4}).
+\ref{fig:fig2} and Fig.
+\ref{fig:fig5}).
 Lower overall eDNA concentrations and fewer deep water samples may make similarities among species distributions more difficult to detect at 300m and deeper.
@@
 \centering
-\includegraphics[width=10cm]{plots/2_Figure_2.jpg} 
+%\includegraphics[width=10cm, height = 13cm]{plots/2_Figure_2.jpg} 
+\includegraphics[width=0.6\textwidth, height=0.6\textwidth]{plots/2_Figure_2.pdf} 
 \caption{Correlation plot describing species co-abundance and co-occurance.
@@
 Species are ordered in known ecological groups (surface, midwater, and deepwater) to visually indicate the gradual change of species co-occurance.}
-\label{fig:fig2}
+\label{fig:fig3}
 \end{figure}
@@
 S2} for defining species presence) from eDNA varied substantially and predictably with depth, with the highest richness observed at the surface (Fig.
-\ref{fig:fig3}\textit{A}).
+\ref{fig:fig4}\textit{A}).
 In part, this pattern is likely to be an artifact of the focal species included in this study -- chosen for their relevance to fisheries management as well as for their frequency in the observations (see Methods) -- but the richness-depth gradient also reflects the expectation that habitat and nutrient availability is greatest near the surface \cite{smith2002,hickey2005}.
@@
 Species richness also corresponds strongly with the influence of Columbia River plume (Fig.
-\ref{fig:fig3}\textit{A}), which extends well beyond the coastline \cite{hickey2003}.
+\ref{fig:fig4}\textit{A}), which extends well beyond the coastline \cite{hickey2003}.
 The nitrate-rich river water interacts with ocean currents and intensifies coastal upwelling, enhancing both primary and secondary productivity \cite{hickey2010} and likely influencing the patterns of species richness we observe (Fig.
-\ref{fig:fig3}\textit{A}) \cite{tolimieri2015}.
+\ref{fig:fig4}\textit{A}) \cite{tolimieri2015}.
 The total fish eDNA accumulation (Fig.
-\ref{fig:fig3}\textit{B}) suggests the highest aggregation of forage fish DNA near Heceta Bank (44 degrees N; overlapping anchovy and sardine distributions are the two largest contributors to the observed peak of accumulated eDNA concentration) in addition to other commercialy important mesopelagic fish (\textit{Sebastes entomelas} and \textit{Tarletonbeania crenularis}) a center of abundance for commercial fisheries \cite{bosley2014,Tissot2008TwentyYO}, and elsewhere nearshore in areas of coastal upwelling.
+\ref{fig:fig4}\textit{B}) suggests the highest aggregation of forage fish DNA near Heceta Bank (44 degrees N; overlapping anchovy and sardine distributions are the two largest contributors to the observed peak of accumulated eDNA concentration) in addition to other commercialy important mesopelagic fish (\textit{Sebastes entomelas} and \textit{Tarletonbeania crenularis}) a center of abundance for commercial fisheries \cite{bosley2014,Tissot2008TwentyYO}, and elsewhere nearshore in areas of coastal upwelling.
 
-\begin{figure*}[tbhp] 
+\begin{figure}[tbhp] 
 \centering
-\includegraphics[width=13cm]{plots/3_Figure_3.jpg} 
+\includegraphics[width=13cm]{plots/3_Figure_3.pdf} 
 \caption{Spatial and vertical distribution of species richness (A) and environmental DNA (B) concentration across depths (0, 50, 150, 300, and 500 m).
-(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see Fig.
-\ref{fig:fig4} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
-\label{fig:fig3}
-\end{figure*}
+(B) Species richness represented as the cumulative number of species present (species with higher than 1.5 copies/L (see \href{SI_Appendix.pdf}{SI Appendix, Fig.
+S2} for defining presence) on smoothed estimated eDNA concentrations) and acumulated eDNA concentration representing the cumulative sum of eDNA (copies/L) across all species.}
+\label{fig:fig4}
+\end{figure}
 
-\begin{figure*}[tbhp] 
+\begin{figure}[tbhp] 
 \centering
-\includegraphics[width=16cm]{plots/4_Figure_4.jpg} 
+\includegraphics[width=16cm]{plots/4_Figure_4.pdf} 
 \caption{Vertical (A) and latitudinal (B) distribution of eDNA concentrations grouped in ecological groups (surface south, surface north, midwater, and deepwater).
@@
 Key geographic features, such as the Columbia River and Heceta Bank, are marked to indicate alignment of species distributions.}
-\label{fig:fig4}
-\end{figure*}
+\label{fig:fig5}
+\end{figure}
 
@@
 Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; Widow rockfish occurs almost exclusively at the 50m depth; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig.
-\ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station.
+\ref{fig:fig4}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station.
 
@@
 S6}) and spatially smoothed maps (Fig.
-\ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
+\ref{fig:fig2} and \href{SI_Appendix.pdf}{SI Appendix, Fig.
 S1}) revealed minimal differences.
